### 2021-02-08 cover-letter/cover-letter.Rmd
Initial submission to the R Journal

---
title: ""
output: pdf_document
bibliography: lg.bib
indent: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##

\pagenumbering{gobble}

\vspace{1cm}

\begin{flushright}
Håkon Otneim \\
Department of Business and Management Science \\
NHH Norwegian School of Economics \\
Helleveien 30, 5045 Bergen
Norway
\end{flushright}

\noindent 23 September 2019

\vspace{1cm}

### Dear Editor.

I wish to submit an original research article entitled *lg: An R-package for local Gaussian approximations* for consideration by the *R Journal*. I confirm that this work is original, and that it has not been published elsewhere, nor is it currently under consideration for publication elsewhere.

There have in recent years been published a large number of papers developing statistical methodology using local Gaussian approximations. The papers appear in very good statistics journals such as *Journal of Econometrics*, *Journal of Business and Economic Statistics*, *Journal of Empirical Finance*, *Statistics and Computing* and *Journal of Time Series Analysis*, and they develop various aspects and applications of the concept. Simply stated, one may approximate a general joint probability density function $f(x)$ locally using a multivariate normal distribution:

$$f(x) = f(x, \mu(x), \Sigma(x)),$$
where the local parameter functions $\mu(x)$ and $\Sigma(x)$ are estimated from data by means of a local likelihood procedure. The off-diagonal elements in the local correlation matrix $\Sigma(x)$ have been termed *local Gaussian correlations* (LGC) and are taken as measures of nonlinear dependence between stochastic variables. This construction is applied to e.g. density estimation and independence testing, both for independent observations and in time series analysis.

The first paper in this series [@tjos:huft:2013] was quickly followed by an R-package **localgauss** and a companion paper [@bere:klep:tjos:2014] that implements and describes the basic estimation problem, but this package has not been updated with new functionality as new methods have been published. Quite on the contrary: due to incompatibility issues with the **ggplot2**-package, the authors of **localgauss** removed the plotting capabilities from their package, while authors of new statistical methods using the LGC (of whom I am one) have kept their own separate R-scripts to be shared upon request.

Two factors prompted me to write the new R package **lg** for applications of the LGC, as well as a companion paper:

1. I, my colleagues and my co-authors receive many requests for code for implementation of the statistical methods that we have published. This is not sustainable in the long run, because there exist several versions of the same scripts, and also because several authors have left the academic world, and more may do so in the future.
2. I am not aware of new major methodological developments using the LGC that is currently under development. The literature cited in the submitted paper represents a mature body of work, carried out by several researchers, and published in good journals. My R package provides a unified framework to carry out all LGC-based statistical methods, except some work on spectral analysis that will be published in a separate package by another author, and it will likely stay relevant and be cited as long as the LGC, in its various shapes and forms, is applied in empirical problems.

Besides describing the use and principles of the **lg**-package, the paper that I now submit for publication provides an up-to-date survey of important developments in the LGC framework that will be useful to a broader audience than R programmers alone, which is why I believe that the paper is appropriate for publication by the *R Journal*.

Code for easy replication of all examples in the paper has been submitted as a separate supplement to the article.

I have no conflicts of interest to disclose.

Please address all correspondence concerning this manuscript to me at \texttt{hakon.otneim@nhh.no}.

Thank you for your consideration of the manuscript.

\vspace{1cm}

\noindent Sincerely,

\vspace{.5cm}

\noindent Håkon Otneim

## References

### 2021-02-08 otneim.tex
Initial submission to the R Journal

% !TeX root = RJwrapper.tex
\title{\CRANpkg{lg}: An R package for Local Gaussian Approximations}
\author{by H{\aa}kon Otneim}

\maketitle

\abstract{
The package \CRANpkg{lg} for the R programming language provides implementations of recent methodological advances on applications of the local Gaussian correlation. This includes estimation of the local Gaussian correlation itself, multivariate density estimation, conditional density estimation, various tests for independence and conditional independence, as well as a graphical module for creating dependence maps. This paper describes the \CRANpkg{lg} package, its principles and its practical use.
}

\section{Introduction}

\citet{tjos:huft:2013} propose the \emph{local Gaussian correlation} (LGC) as a new measure of statistical dependence between two stochastic variables $X_1$ and $X_2$, which has the following important property yet unrivaled in the literature: It can separate between positive and negative nonlinear dependence while still reducing to the ordinary Pearson's correlation coefficient if $X_1$ and $X_2$ are jointly normally distributed. The R-package \CRANpkg{localgauss} \citep{bere:klep:tjos:2014} provides two important functions in this context; one that calculates the sample LGC based on observed values of $(X_1,X_2)$, and one that uses the estimated LGC to perform a local test of independence between $X_1$ and $X_2$ as described in detail by \citet{bere:tjos:2014}.

We have lately seen a number of new applications of the LGC that the \CRANpkg{localgauss}-package does not support, however. \citet{stov:tjos:huft:2014} use the LGC to test for financial contagion across markets during crises. \citet{otne:tjos:2017} present a procedure for estimating multivariate density functions via the LGC, which \citet{otne:tjos:2018} modify in order to compute estimates of conditional density functions. \citet{laca:tjos:2017} present a test for serial independence within a time series, which \citet{laca:tjos:2018} extend in order to include a test for cross-correlation between two time series. Finally, \citet{otne:tjos:2019} develop the local Gaussian \emph{partial} correlation (LGPC) as a measure of conditional dependence, and a corresponding test for conditional independence.

This paper describes the \CRANpkg{lg}-package \citep{otne:2019}, which provides a unified framework to implement all these methods, as well as a tool for visualizing the LGC and LGPC as dependence maps (and conditional dependence maps). \citet{jord:tjos:2017} use the LGC in spectral analysis of time series, but those methods have their own computational ecosystem in the \pkg{localgaussSpec}-package, \citet{jord:2018}.

In Section \ref{chap:background} we provide a brief introduction to the LGC as well as the methods and applications referred to above. In Section \ref{chap:lgobject} we describe the core function in the \CRANpkg{lg}-package and move on to demonstrate the implementation of various applications in Section \ref{chap:inference}. We conclude this paper in Section \ref{chap:graphics} by demonstrating the graphical capabilities of the \CRANpkg{lg}-package.

\section{Statistical background} 
\label{chap:background}

Consider a random vector $X$ having the unknown probability density function $f_X(x)$. Based on a random sample $X_1,\ldots, X_n$ we wish to produce an estimate $\widehat f_X(x)$ of $f_X(x)$, to which end the statistical literature provides an abundance of methods. One may, for example, assume that $f_X \in F_{\theta}$, where $F_{\theta} = \{f(x;\theta), \theta \in \Theta\}$ is a family of density functions indexed by some parameter $\theta$ living in a parameter space $\Theta$, and estimate $\theta$ by means of the maximum likelihood method. On the other hand, we may estimate $f_X(\cdot)$ directly from the data using the classical kernel density estimator 
$$\widehat f_X(x) = \frac{1}{nb}\sum_{i=1}^n K\left(\frac{X_i - x}{b}\right),$$
where $K$ is a symmetric density function (the kernel) and $b$ is a tuning parameter (the bandwidth) controlling the smoothness of $\widehat f_X(\cdot)$. This requires very little prior assumptions. There is a massive literature on density estimation that include countless variations of the parametric and nonparametric approaches, as well as many other clever and creative techniques.

\citet{hjor:jone:1996} provide one such idea. They consider a parametric family $F_{\theta}$, but instead of searching for a single parameter value $\theta_0$ for which $f_X(x) = f(x;\theta_0)$ (or approximately so), they rather assert that different members of $F_{\theta}$ may approximate $f_X$, locally, in different parts of its domain. In other words, they seek to estimate a parameter \emph{function} $\theta_0(x)$ for which $f_X(x) = f(x;\theta_0(x))$ (or approximately so), and do this by pointwise maximizing a \emph{local likelihood function}:
\begin{align}
\widehat\theta(x) 
&= \arg\max_{\theta \in \Theta} L_n(\theta, x) \nonumber\\ 
&= \arg\max_{\theta \in \Theta} \frac{1}{nb}\sum_{i=1}^n K\left(\frac{X_i - x}{b}\right)\log f(X_i; \theta) - \int \frac{1}{b}K\left(\frac{y - x}{b}\right)f(y;\theta)\,\textrm{d}y,
\label{eq:loclik}
\end{align}
where, again, $K$ is a symmetric density function and $b$ is a bandwidth parameter that controls the smoothness of the estimate.

\citet{tjos:huft:2013} consider the bivariate case $X = (X_1, X_2)$ and take $F_{\theta}$ to be the family of bivariate normal distribution consisting of densities on the form

\begin{align}
f(x;\theta) 
&= \psi(x_1, x_2;\mu_1, \mu_2, \sigma_1, \sigma_2, \rho)  \nonumber\\ 
&= \frac{1}{2\pi \sigma_1 \sigma_2\sqrt{1-\rho^2}} \nonumber \\
& \qquad \times\exp \left\{-\frac{1}{2(1-\rho^2)}\left(\frac{(x_1-\mu_1)^2}{\sigma_1^2}-2\rho\frac{(x_1-\mu_1)(x_2-\mu_2)}{\sigma_1 \sigma_2}+\frac{(x_2-\mu_2)^2}{\sigma_2^2}\right)\right\}.
\label{eq:gaussian}
\end{align}
Using a sample $\{X_{1i}, X_{2i}\}$, $i = 1, \ldots, n$, they estimate 
$$\widehat\theta(x) = (\widehat\mu_1(x), \widehat\mu_2(x), \widehat\sigma_1(x), \widehat\sigma_2(x), \widehat\rho(x))$$
and take special interest in the estimated correlation function $\widehat\rho(x)$ (i.e. the LGC) because it serves as an attractive local measure of statistical dependence between $X_1$ and $X_2$. They show that the LGC reveals many types of nonlinear statistical dependence not seen by the ordinary (global) Pearson's correlation coefficient. Furthermore, the LGC distinguishes between positive and negative dependence and reduces to Pearson's $\rho$ if $X_1$ and $X_2$ are jointly normal. We refer to \cite{tjos:huft:2013} for a detailed treatment of the theoretical foundations of the LGC as well as several examples, and rather present two simple illustrations at this point in order to demonstrate the concept.

\begin{figure}[t]
    \centering
    \begin{subfigure}[t]{0.48\textwidth}
      \includegraphics[width=\textwidth]{gaussian-example}
      \caption{Estimated local Gaussian correlations between two  jointly normal variables having correlation equal to 0.5, based on 1000 observations.}
    \end{subfigure}
    \begin{subfigure}[t]{0.48\textwidth}
      \includegraphics[width=\textwidth]{returns-example}
      \caption{Local Gaussian correlation between daily returns from the CAC40 and FTSE100 stock indices, based on 1000 consecutive trading days.}
    \end{subfigure}
    \caption{Two dependence maps}
    \label{fig:initial-example}
  \end{figure}

In Figure \ref{fig:initial-example} we have plotted the estimated LGC for two bivariate data sets; 1000 simulated observations from a binormal distribution having correlation equal to 0.5, and the daily return on the CAC40 and FTSE100 stock indices on 1000 consecutive trading days starting on May 5th 2014. In the first panel, we see that the estimated local correlation coincides with the global Pearson's $\rho$ except for the estimation error. In the second panel we see clearly that the local correlation, and thus, as we will see, the dependence, is stronger in the lower left and upper right tails than in central parts of the distribution. This phenomenon is of course well known in the financial literature, but can now also be easily measured and quantitatively interpreted by means of the LGC. One may obtain these particular estimates from the older \CRANpkg{localgauss}-package (as well as the \CRANpkg{lg}-package of course), but the plotting routine that was used to produce these figures is included in the \CRANpkg{lg}-package and will be described in more detail in Section \ref{chap:graphics}.

Taking the LGC as a measure of dependence opens up for a number of possibilities to construct statistical tests. \citet{bere:tjos:2014} derive from the representation (\ref{eq:gaussian}) that $\rho(x) \equiv 0$ implies that $X_1$ and $X_2$ are independent (the converse statement holds if $\mu_i(x_1, x_2) = \mu_i(x_i)$ and $\sigma_i(x_1, x_2) = \sigma_i(x_i)$ for $i=1,2$, and in general if the observations have been suitably transformed according to a procedure presented later in this section). It follows then that departures from $\widehat\rho(x) \equiv 0$ may be taken as evidence against $X_1$ and $X_2$ being statistically independent. \citet{bere:tjos:2014} formalize this notion by testing whether $\rho(x) \equiv 0$ for all $x\in S \subset \mathbb{R}^2$ using the test statistic

\begin{equation}
T_{n,b} = \int_S h\left(\widehat \rho(x) \right) \,\textrm{d}F_n(x)
\label{eq:test-statistic1}
\end{equation}
for some function $h$, for example $h(x) = x^2$ or $h(x) = |x|$. Critical values may be obtained by permutations of the data under the null hypothesis. We demonstrate the implementation of this test using the \CRANpkg{lg}-package in Section \ref{chap:indtests}.

Consider next the stationary time series $\{X_t\}$. The autocorrelation $\rho_k = \rho(X_t, X_{t-k})$ is a fundamental descriptive and analytical concept, but is, again, only capable to completely capture \emph{linear} serial dependence relationships. \citet{laca:tjos:2017} seek to remedy this by rather calculating the LGC to measure the \emph{local} dependence between $X_t$ and $X_{t-k}$, which in turn leads to a test for serial independence in the natural way. In fact, this work is mainly a theoretical exercise in order to accommodate dependent data. Testing for serial independence between $X_t$ and $X_{t-k}$ using observations $\{X_t, X_{t-k}\}_{t = 1}^T$ leads to the same test statistic and bootstrap procedure as the test for independence between $X_1$ and $X_2$ that we described above, and we present its implementation in Section \ref{chap:indtests} as well.

\citet{laca:tjos:2018} extend this problem to test for serial cross dependence between two time series $\{X_t, Y_t\}$ by measuring the LGC between $X_t$ and $Y_{t-k}$. Indeed, departures from $\widehat\rho(x,y) \equiv 0$ is again taken as evidence against independence, and the test statistic (\ref{eq:test-statistic1}) provides an aggregate measure of just that in the specified region $S$. This time, however, we can not just obtain permutations of the data in order to produce replicates of the test statistic under the null hypothesis. \citet{laca:tjos:2018} accomplish this by rather suggesting two block bootstrap procedures using fixed and random block sizes. This test is implemented in the \CRANpkg{lg}-package, as we demonstrate in Section \ref{chap:indtests}.

\citet{stov:tjos:huft:2014} measure and test for financial contagion using the LGC. They define contagion by "a significant increase in cross-market linkages after a shock to one country" \citep{forb:rigo:2002} and employ the LGC to quantify such linkage. Contrary to the independence tests above, \citet{stov:tjos:huft:2014} estimate the LGC on a grid $\{x_1, x_2\}_k$, $k=1,\ldots,K$ along the diagonal $D=\{(x_1,x_2): x_1 = x_2\}$ \emph{before} and \emph{after} some critical event in the financial markets, i.e. in the crisis (C) and non-crisis (NC) periods. They compare the two estimates using the test statistic

$$T_{n,b}^D = \sum_{k=1}^K \left\{\widehat\rho_C(x_k, x_k) - \widehat\rho_{NC}(x_k, x_k) \right\}w(x_k,x_k),$$

where $w(\cdot, \cdot)$ is a weight function that serve the same purpose as the integration area $S$ in (\ref{eq:test-statistic1}). In this case, a standard bootstrap will suffice in order to produce approximate replicates of $T_{n,b}^D$ under the null hypothesis of no financial contagion, and we demonstrate the implementation of this test using the \CRANpkg{lg}-package in Section \ref{chap:cont}.

Although \citet{hjor:jone:1996} provide a general framework for locally parametric density estimation using using any parametric family, having any number of variables, it is evident that the method may struggle in multivariate applications much in the same way as the kernel density estimator does. This is a consequence of the curse of dimensionality, the effect of which is sought remedied by an algorithm provided by \citet{otne:tjos:2017}. The idea is to locally fit the $p$-variate normal distribution $\psi(\mu, \Sigma)$, where $\mu$ is the vector of $p$ expectations, and $\Sigma$ is the $p\times p$ covariance matrix (to which the correlation matrix $R$ corresponds), but under the following structural simplifications:

\begin{align}
\mu_i(x) &= \mu_i(x_1, \ldots, x_p) \stackrel{\textrm{def}}{=} \mu_(x_i) \label{eq:simp-mu}\\
\sigma_i(x) &= \sigma_i(x_1, \ldots, x_p)  \stackrel{\textrm{def}}{=} \sigma_i(x_i) \label{eq:simp-sig} \\
\rho_{ij}(x) &= \rho_{ij}(x_1, \ldots, x_p) \stackrel{\textrm{def}}{=} \rho_{ij}(x_i, x_j). \label{eq:simp-rho}
\end{align}
These local parameters are estimated by univariate marginal locally normal fits (eqs. \ref{eq:simp-mu} and \ref{eq:simp-sig}), and pairwise bivariate locally normal fits (eq. \ref{eq:simp-rho}), and in the latter case the estimates $\widehat\mu_i(x_i)$, $\widehat\mu_j(x_j)$, $\widehat\sigma_i(x_i)$ and $\widehat\sigma_j(x_j)$ are kept fixed.

Denote by $F_i(x_i)$, $i = 1,\ldots, p$ the marginal distribution functions of the stochastic vector $X$, and by $\widehat F_i(x_i) = n^{-1}\sum_{i=1}^n 1(X_i \leq x_i)$ their empirical counterparts. \citet{otne:tjos:2017} argue that estimating the density $f_Z(z)$ of the vector $Z = \{\Phi^{-1}(F_i(X_i))\}_{i=1,\ldots,p}$, in practice approximated by

\begin{equation}
\widehat Z = \{\Phi^{-1}(\widehat F_i(X_i))\}_{i=1,\ldots,p}, 
\label{eq:trans}
\end{equation}
where $\Psi(\cdot)$ is the univariate standard normal cdf, will lead to better density estimates in many situations. In that case, one may simplify even further and fix

\begin{equation}
\mu_i(z_i) \stackrel{\textrm{def}}{=} 0 \,\, \textrm{ and } \sigma_i(z_i) \stackrel{\textrm{def}}{=} 1, \,\, i = 1,\ldots,p,
\label{eq:simp}
\end{equation}
so that the only parameter functions left to estimate are the pairwise local Gaussian correlations $R(z) = \{\rho_{ij}(z_i, z_j)\}_{i<j}$, where we use the notation $Z$, $z_i$ and $z_j$ to signify that the estimation is performed on the (approximate) standard normal scale (or $Z$-scale). One then estimates the joint density $f_Z(z)$ of $Z$ as
\begin{equation}
\widehat f_Z(z) = \psi(z; \mu = 0, \sigma = 1, R = \widehat R(z)),
\label{eq:transformed-density}
\end{equation}
and then substitute $f_Z$ for $\widehat f_Z$ in the following relation obtained by \citet{otne:tjos:2017} in order to estimate the unknown density $f_X$:
\begin{equation}
f(x) = f_{Z}\big(\Phi^{-1}(F_1(x_1)), \ldots, \Phi^{-1}(F_p(x_p))\big) \times \prod_{i=1}^p \frac{f_i(x_i)}{\phi\big(\Phi^{-1}(F_i(x_i))\big)},
\label{eq:backtrans}
\end{equation}
where $\phi(\cdot)$ is the standard normal pdf. This estimator is implemented the \CRANpkg{lg}-package as described in Section \ref{chap:density}.

One particular feature enjoyed by the jointly normally distributed vector $X$ is that, for any partitioning $\left(X^{(1)}, X^{(2)}\right)$, the conditional distribution of $X^{(1)}|X^{(2)} = x^{(2)}$ is also normal. In fact, if $X \sim \mathcal{N}(\mu,\Sigma)$, and $\mu$ and $\Sigma$ is partitioned according to $\left(X^{(1)}, X^{(2)}\right)$ as
$$\mu = \begin{pmatrix} \mu_1 \\\mu_2 \end{pmatrix} \,\, \textrm{ and } \,\, \Sigma = \begin{pmatrix} \Sigma_{11} & \Sigma_{12} \\ \Sigma_{21} & \Sigma_{22} \end{pmatrix},$$
then $X^{(1)}|X^{(2)} = x^{(2)} \sim \mathcal{N}(\mu^*, \Sigma^*)$, where

\begin{align}
\mu^* &= \mu_1 + \Sigma_{12}\Sigma_{22}^{-1}\left(x^{(2)} - \mu_2\right) \label{eq:cond-mu} \\
\Sigma^* &= \Sigma_{11} - \Sigma_{12}\Sigma_{22}^{-1}\Sigma_{21}, \label{eq:cond-sig}
\end{align}
see e.g. \citet[chapter 4]{john:wich:2007}. \citet{otne:tjos:2018} demonstrate that this property may be translated into a corresponding local argument without modification. If the joint density $f_X(\cdot)$ is locally Gaussian, i.e. if

$$f_X(x) = \psi(x, \mu(x), \Sigma(x)),$$

then $X^{(1)}|X^{(2)} = x^{(2)}$ is also locally Gaussian with local parameters given by equations (\ref{eq:cond-mu}) and (\ref{eq:cond-sig}), except that all quantities are $x$-dependent. If we use the transformation technique described above together with simplification (\ref{eq:simp}), the local versions of equations (\ref{eq:cond-mu}) and (\ref{eq:cond-sig}) simplify to

\begin{align}
\mu^*(z) &= R_{12}(z)R_{22}(z)^{-1}z^{(2)},  \label{eq:cond-mu-trans}\\
\Sigma^*(z) &= R_{11}(z) - R_{12}(z)R_{22}(z)^{-1}R_{21}(z), \label{eq:cond-sig-trans}
\end{align}
where we, again, switch to $z$-notation in order to make it clear that these quantities are estimated on the standard normal $z$-scale. An estimator $\widehat f_{X^{(1)}|X^{(2)}}(\cdot|\cdot)$ of the conditional density $f_{X^{(1)}|X^{(2)}}(\cdot|\cdot)$ follows immediately from an expression corresponding to (\ref{eq:backtrans}), and the \CRANpkg{lg}-package provides functions for implementing this estimator in R as described in Section \ref{chap:density}.

Finally, we refer to \citet{otne:tjos:2019}, who take the local version of the conditional covariance matrix (\ref{eq:cond-sig}) (or (\ref{eq:cond-sig-trans}) in the transformed case) as a measure for conditional dependence, and thus as an instrument to test for conditional independence. Specifically, consider the stochastic vector $X = \left(X^{(1)}, X^{(2)}, X^{(3)}\right)$, where $X^{(1)}$ and $X^{(2)}$ are scalars and $X^{(3)}$ may be a vector. $X^{(1)}$ is conditionally independent from $X^{(2)}$ given $X^{(3)}$, written $X^{(1)} \perp X^{(2)} \,|\,X^{(3)}$ if the stochastic variables $X^{(1)}\,|\,X^{(3)}$ and $X^{(2)}\,|\,X^{(3)}$ are independent, or, equivalently, if the joint conditional density of $X^{(1)}$ and $X^{(2)}$ given $X^{(3)}$ can be written as the product

\begin{equation}
f_{X^{(1)},X^{(2)}|X^{(3)}}\left(x^{(1)}, x^{(2)}|x^{(3)}\right) = f_{X^{(1)}|X^{(3)}}\left(x^{(1)}|x^{(3)}\right) \times f_{X^{(2)}|X^{(3)}}\left(x^{(2)}|x^{(3)}\right).
\label{eq:cond-ind}
\end{equation}
In this case, denote by $\alpha(z)$ the off-diagonal element in the $2\times2$ local correlation matrix $R^*(z)$ (which derives directly from $\Sigma^*(z)$ as given in (\ref{eq:cond-sig-trans})). If $X$ has a local Gaussian distribution, the conditional independence (\ref{eq:cond-ind}) is equivalent to $\alpha(z) \equiv 0$, departures from which we take as evidence against conditional dependence between $X^{(1)}$ and $X^{(2)}$ given $X^{(3)}$, and this discrepancy is quantified using the test functional

\begin{equation}
T^{CI}_{n, b} = \int h(\widehat\alpha(z)) \, \textrm{d}F_n(z).
\label{eq:ci-test-statistic}
\end{equation}
\citet{otne:tjos:2019} describe a bootstrap procedure for generating replicates of $T_{n,b}^{CI}$ under the null hypothesis. In Section \ref{chap:partial} we demonstrate how the \CRANpkg{lg}-package  may be used to extract estimates of the local partial correlation and perform tests for conditional independence according to this scheme.

\section{The \texttt{lg}-object} 
\label{chap:lgobject}

The local Gaussian correlation may be used to perform a number of statistical analyses, as is evident from the preceding section. The practitioner must first, however, make three quite specific modelling choices; namely (i) to choose an estimation method, i.e. the level of simplification in multivariate applications, (ii) to determine whether the data should be transformed towards marginal standard normality before estimating the LGC, and (iii) to choose a set of bandwidths, or at least a method for calculating bandwidths. The architecture of the \pkg{lg}-packages requires the user to make these choices \emph{before} endeavoring further into specific applications by imposing a strict, two step procedure:

\begin{itemize}
\item[1.] Create an lg-object.
\item[2.] Apply relevant analysis functions to the lg-object.
\end{itemize}

In the following we assume that one has a \emph{data set} \code{x} loaded into the R workspace, which must be an $n\times p$ matrix (one column per variable, one row per observation), possibly including \code{NA}`s which will be excluded, or a data frame having the same dimensions. The fundamental syntax for creating an lg-object is \code{lg\_object <- lg\_main(x)}, and we will in this section explain how the modelling decisions (i)-(iii) can be encoded into the lg-object.

\subsection{Selecting the estimation method}
\label{chap:est-method}

Given a data set \code{x} having $n$ rows and $p \geq 2$ columns, the user must choose between four distinct estimation methods, and specify this choice by using the argument \code{est\_method} to the \code{lg\_main()}-function. We look at the built-in bivariate data set `faithful` and load the \CRANpkg{lg}-package in order to demonstrate the implementation:

\begin{example}
R> x <- faithful
R> library(lg)
\end{example}

\textbf{1. A full locally Gaussian fit for bivariate data.} If $p = 2$, we may fit the bivariate normal $\psi(x, \theta)$ locally to $f(x)$, and by a "full local fit" we mean that we jointly estimate the five parameters
$$\theta(x) = \big(\mu_1(x_1, x_2), \mu_2(X_1, x_2), \sigma_1(x_1, x_2), \sigma_2(x_1,x_2),\rho(x_1, x_2)\big)$$
by optimizing the local likelihood function (\ref{eq:loclik}) in the grid point $x = (x_1, x_2)$. To use this estimation method in the subsequent analysis, specify \code{est\_method = "5par"} in the call to \code{lg\_main()}:

\begin{example}
R> lg_object <- lg_main(x, est_method = "5par")
\end{example}
The resulting \code{lg\_object} is a list of class \code{lg}, and we may confirm that the assignment has been carried out correctly by inspecting its \code{est\_method}-element:
\begin{example}
R> lg_object$est_method

[1] "5par"
\end{example}

Note that the full locally Gaussian fit for raw data is not available if the number of variables $p$ is greater than 2. The \code{lg\_main()}-function will check for this and throw an error if $p>2$ and \code{est\_method = "5par"}.

\textbf{2. A simplified locally Gaussian fit for multivariate data.} As described in the preceding section, we may construct a simplified estimation procedure for calculating the LGC in two steps, which in principle works for any number of dimensions (including $p=2$):

\begin{itemize}
\item[1.] Calculate $\mu_i(x_i)$ and $\sigma_i(x_i)$, $i = 1,\ldots,p$ by fitting the univariate normal distribution locally to each \emph{marginal} density $f_i(x_i)$ of $f(x)$.
\item[2.] Keep $\widehat\mu_i(x)$ and $\widehat\sigma_i(x_i)$, $i=1,\ldots,p$ from step 1 fixed when estimating $\rho_{ij}(x_i, x_j)$, $1\leq i<j\leq p$ by fitting the bivariate normal distribution to each \emph{pair} of variables $X_i$ and $X_j$.
\end{itemize}
To use this method, create the lg-object by running the following line:

\begin{example}
R> lg_object2 <- lg_main(x, est_method = "5par_marginals_fixed")
\end{example}

\textbf{3. A simplified locally Gaussian fit for marginally standard normal data.} This estimation method is applicable for marginally standard normal data, or data that have been transformed to approximate marginal standard normality by e.g. the transformation (\ref{eq:trans}). In that case, we fix the marginal expectation functions and standard deviation functions to the constant values $0$ and $1$ respectively, rather then estimating them. To use this estimation method, create the lg-object by running

\begin{example}
R> lg_object3 <- lg_main(x, est_method = "1par")
\end{example}

Note that the function call above will issue a warning if the option for transforming the data to marginal standard normality is not at the same time set to \code{TRUE}, see the next sub-section for details.

\textbf{4. A full locally Gaussian fit for trivariate data.} If $p = 3$ \emph{and} we choose to transform the data to marginal standard normality (see the next sub-section), the transformed density $f_Z(\cdot)$ in (\ref{eq:transformed-density}) may be estimated by jointly estimating the three local correlations $\rho_{12}(z_1, z_2, z_3)$, $\rho_{13}(z_1, z_2, z_3)$ and $\rho_{23}(z_1, z_2, z_3)$. This estimation method was introduced recently by \citet{otne:tjos:2019} in order to increase power of their conditional independence test, but it can be used in any application described in this paper that consider trivariate data. To use this estimation method, create the lg-object by running

\begin{example}
R> lg_object4 <- lg_main(x, est_method = "trivariate")
\end{example}

\subsection{Data transformation}
\label{data-transformation}

\begin{figure}[t]
    \centering
    \begin{subfigure}[t]{0.48\textwidth}
      \includegraphics[width=\textwidth]{faithful}
      \caption{The original data.}
    \end{subfigure}
    \begin{subfigure}[t]{0.48\textwidth}
      \includegraphics[width=\textwidth]{faithful-transformed}
      \caption{The same data transformed to marginal standard normality.}
    \end{subfigure}
    \caption{The same data on different scales.}
    \label{fig:faithful}
  \end{figure}

Next, the user must determine if the local Gaussian correlation should be estimated directly on the raw data or on the marginally normal pseudo observations \eqref{eq:trans}. This is carried out by using the logical \code{transform\_to\_marginal\_normality}-argument in \code{lg\_main}, for example:

\begin{example}
R> lg_object <- lg_main(x, transform_to_marginal_normality = TRUE)
\end{example}

The resulting \code{lg\_object} now includes the element \code{transform\_to\_marginal\_normality} set according to the input, and if this is \code{TRUE}, it also includes the \code{transformed\_data} and a function \code{trans\_new()} that may be used later to apply the same transformation to e.g. grid points. If the transformation option is set to \code{FALSE}, the \code{transformed\_data} element contains the input data \code{x}, and \code{trans\_new()} is nothing more than the identity mapping for points in \(\mathbb{R}^p\). See Figure \ref{fig:faithful} for two plots that demonstrate the effect of the data transformation on the example data.

\subsection{Bandwidth selection}\label{chap:bw}

Finally, the user must specify a set of bandwidths, or a method for calculating them. Given that the different estimation methods described in Section \ref{chap:est-method} require different sets of bandwidths (i.e joint, marginal and/or pairwise), the easiest approach for the user is to leave the selection and formatting of the bandwidths to the \code{lg\_main()}-function.

The bandwidth plays a slightly different role in local likelihood estimation than elsewhere in the nonparametric literature. It controls the \emph{level of localization} and thus only indirectly the \emph{smoothness} of the estimates. Indeed, if we concentrate on the univariate case for the moment and assume that the (single) bandwidth \(b\) is small, we see from the local likelihood function \eqref{eq:loclik} that only the very few observations closest to a fixed grid point \(x_0\) will have significant weight when determining the
local parameter estimates \(\widehat\theta_0(x)\) at that point. Moving on to another nearby point \(x_1\) may then lead to a fairly different estimate \(\widehat\theta(x_1)\) because the set of observations having weight in this point is very different. This may again lead to rougher parameter estimates \(\widehat\theta(x)\) and in turn also to rougher density estimates \(f(x,\widehat\theta(x))\).

If the bandwidth \(b\) grows large, on the other hand, all observations receive similar weights, and furthermore: the local likelihood function \eqref{eq:loclik} becomes approximately proportional to the ordinary (global) likelihood function \(L_n(\theta) = \sum_{i=1}^n\log f(X_i;\theta)\). In other words, the local parameter estimates \(\widehat \theta(x)\) are smoothed towards the constant maximum likelihood estimates \(\widehat\theta_{ML}\), and the estimated density \(f(x;\widehat\theta(x))\) towards the maximum likelihood estimate \(f(x;\widehat\theta_{ML})\). This means that the bandwidth may be chosen to reflect the goodness-of-fit of \(f(x;\theta)\) to the true density \(f(x)\).

In the multivariate applications referred to in this paper, the bandwidth \(b\) in \eqref{eq:loclik} is a diagonal matrix, and \(1/b\) is
naturally taken to represent its inverse.

We have in practice seen two automatic bandwidth selectors employed in the applications referred to in Section \ref{chap:background}: a
cross-validation procedure, which is slow, but accurate with respect to density estimation, and a plug-in bandwidth that is quick, but cruder. We use the argument
\code{bw\_method} to the \code{lg\_main()}-function in order to choose between the two.

\textbf{1. Choosing bandwidths by cross-validation.} The functional

\[CV(b) = -\frac{1}{n}\sum_{i=1}^n \log f\Big(X_i; \widehat\theta^{(-i)}(X_i)\Big),\]
where \(\widehat\theta^{(-i)}(x)\) is the parameter estimate obtained after deleting observation \(X_i\) from the data, is proportional to a quantity that estimates the Kullback-Leibler distance between \(f(\cdot, \widehat\theta(\cdot))\) and the true density \(f(\cdot)\), see \citet{bere:tjos:2014}. The cross-validated bandwidth \(b_{CV}\) is hence given by \[b_{CV} = \arg\min_b \,\, CV(b).\] If we, for example, wish to use the simplified estimation procedure on the transformed data, we need bandwidths for the marginal estimates of the local means and standard deviations, as well as a \(2\times2\) diagonal bandwidth matrix for each pair of variables. This is accomplished by the following call to \code{lg\_main()}:

\begin{example}
R> lg_object <- lg_main(x,
R+                      est_method = "5par_marginals_fixed",
R+                      transform_to_marginal_normality = TRUE,
R+                      bw_method = "cv")
\end{example}

The \code{lg\_object} now contains the necessary bandwidths for this configuration, as can be seen by inspecting the contents of its
\code{bw}-element:

\begin{example}
R> lg_object$bw

$marginal
[1] 0.9989327 0.9875333

$marginal_convergence
[1] 0 0

$joint
  x1 x2       bw1      bw2 convergence
1  1  2 0.2946889 0.331971           0
\end{example}
This is itself a list, containing the crucial elements \code{marginal} for the \(p\) marginal bandwidths, and \code{joint} that contains the
\(p(p-1)/2\) bandwidth matrices, one for each pair of variables (which in this bivariate example just \emph{one} variable pair, (\code{x1,\ x2})). The convergence flags stem from the \code{optim()} and \code{optimize()} functions that we use to obtain the minimizer of \(CV(\cdot)\), and 0 indicates successful convergence.

\textbf{2. Using plug-in bandwidths.} Obtaining cross-validated bandwidths is unfortunately fairly slow on a standard computer. For sample sizes in the 500-1000 range, the process may take several minutes, which is unfeasible when embarking on analyses that require e.g. resampling. We have therefore implemented a quick plug-in bandwidth selector as well, that may suffice in many practical situations, especially at the initial or exploratory stage.

\citet{otne:tjos:2017} show that the simplified version of the local Gaussian fit have the same convergence rates as the corresponding nonparametric kernel density estimator, for which \citet{silv:1986} derives the plug-in formula \(b = 1.08\cdot sd(x) \cdot n^{-1/5}\). By
specifying \code{bw\_method\ =\ "plugin"}, the \code{lg\_main()}-function will select the bandwidths correspondingly, except that the exponent changes to \(-1/6\) for joint bandwidths, and the proportionality constant is by default set to \(1.75\). The latter number is the result of regressing \(b_{CV}\) on \(n^{-1/6}\) in a large simulation experiment covering various data generating processes \citep{otneim2016multivariate}. We see
the effect of switching to plugin bandwidths in the code below:

\begin{example}
R> lg_object <- lg_main(x,
R+                      est_method = "5par_marginals_fixed",
R+                      transform_to_marginal_normality = TRUE,
R+                      bw_method = "plugin")
R> lg_object$bw

$marginal
[1] 0.5703274 0.5703274

$marginal_convergence
[1] NA NA

$joint
  x1 x2       bw1       bw2 convergence
1  1  2 0.6875061 0.6875061          NA
\end{example}

\subsection{Summary of the initialization function}
\label{summary-of-the-initialization-function}

\renewcommand{\arraystretch}{1.2}
\begin{table}[t]
\centering
\begin{tabular}{lll}
\toprule
Argument & Explanation & Default value \\
\midrule
\code{x} & The data, an $n\times p$ matrix or data frame & \\
\code{bw{\textunderscore}method} & Method for calculating the bandwidths & \code{"plugin"} \\
\code{est{\textunderscore}method} & Estimation method & \code{"1par"} \\
\code{transform{\textunderscore}to{\textunderscore}} & & \\
\code{  marginal{\textunderscore}normality} & Transform the data? & \code{TRUE} \\
\code{bw} & The bandwidths to use if already calculated & \code{NULL} \\
\code{plugin{\textunderscore}constant{\textunderscore}marginal} & Prop. const. in plugin formula for marg. bw. & $1.75$ \\
\code{plugin{\textunderscore}exponent{\textunderscore}marginal} & Exponent in plugin formula for marg. bw.& $-1/5$ \\
\code{plugin{\textunderscore}constant{\textunderscore}joint} & Prop. const. in plugin formula for joint bw. & $1.75$ \\
\code{plugin{\textunderscore}exponent{\textunderscore}joint} & Exponent in plugin formula for joint bw.& $-1/6$ \\
\code{tol{\textunderscore}marginal} & Abs. tolerance when optimizing $CV(b)$, marg.& $10^{-3}$ \\
\code{tol{\textunderscore}joint} &Abs. tolerance when optimizing $CV(b)$, joint& $10^{-3}$ \\
\bottomrule
\end{tabular}
\caption{Arguments to the initialization function \code{lg{\textunderscore}main()}}
\label{tab:arguments-lg-main}
\end{table}

In the sub-section above we present the three most important arguments to \code{lg\_main()}. Each of them allows the user to configure one of the three crucial modelling choices. Let us complete this treatment by covering some possibilities to make further adjustments to those choices.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  The user may supply the bandwidths directly to \code{lg\_main()} by passing them to the \code{bw}-argument. They have to be in the   correct format though, which is a list containing the vector \code{\$marginal} if \code{est\_method\ =\ 5par\_marginals\_fixed}, and always a data frame \code{\$joint} specifying all variable pairs in the \code{x1} and \code{x2} columns and the corresponding bandwidths
  in the \code{bw1} and \code{bw2} columns. The function \code{bw\_simple()} will assist in creating bandwidth objects.
\item
  If \code{bw\_method\ =\ "plugin"} the user may change the proportionality constant and exponent in the plugin formula for the joint and, if applicable, the marginal bandwidths. See Table \ref{tab:arguments-lg-main} for the necessary argument names.
\item
  If \code{bw\_method\ =\ "cv"}, the user may change the numerical tolerance in the optimization of \(CV(b)\). See Table \ref{tab:arguments-lg-main} for the necessary argument names.
\end{enumerate}

\section{Statistical inference using the \texttt{lg}-package}
\label{chap:inference}

We proceed in this section to demonstrate how to implement each of the tasks that we discussed in Section \ref{chap:background}. The general pattern is to pass an lg-object, see the preceding section, to one of the estimation or test functions provided in the \CRANpkg{lg}-package. We will look at some financial data in the examples: the monthly returns on the S\&P500, FTSE100, DAX30 and TOPIX stock indices from January 1985 to March 2018. This data is available as an online appendix to this article.

\subsection{Density estimation}\label{chap:density}

\begin{figure}[t]
    \centering
    \begin{subfigure}[t]{0.48\textwidth}
      \includegraphics[width=\textwidth]{loccor-example}
      \caption{The estimated pairwise LGC for the stock \newline data along the diagonal.}
      \label{fig:lgc-plot1}
    \end{subfigure}
    \begin{subfigure}[t]{0.48\textwidth}
      \includegraphics[width=\textwidth]{density-example}
      \caption{The corresponding joint density estimate of the four return series, plotted along the diagonal in $\mathbb{R}^4$.}
      \label{fig:lgc-plot2}
    \end{subfigure}
    \caption{Important output from the \code{dlg()-function}.}
    \label{fig:lgc-plot}
  \end{figure}

We start by introducing a basic function for estimating the LGC on a grid as described by \citet{otne:tjos:2017}, and thus also a probability density estimate. We create a grid, \code{x0}, having the same number of columns as the data in the code below. Note that we use the pipe operator \code{\%\textgreater{}\%} from the \CRANpkg{magrittr}-package \citep{bach:wick:2014} as well as functions from the \CRANpkg{dplyr}-package \citep{wick:fran:mull:2018} for easy manipulation of data frames. We then pass the grid and the lg-object containing our modelling choices to the \code{dlg()}-function in order to do the estimation.

\begin{example}
R> # Create an lg-object
R> lg_object <- lg_main(x = stock_data %>% select(-Date),
R+                      est_method = "1par",
R+                      bw_method = "plugin",
R+                      transform_to_marginal_normality = TRUE)
R> 
R> # Construct a grid diagonally through the data.
R> grid_size <- 100
R> x0 <- stock_data %>% 
R+   select(-Date) %>% 
R+   apply(2, function(y) seq(from = -7,
R+                            to = 7,
R+                            length.out = grid_size))
R> 
R> # Estimate the local Gaussian correlation on the grid
R> density_object <- dlg(lg_object, grid = x0)
\end{example}
The last line of code creates a list containing a number of elements. The two most important are \code{\$loc\_cor}, which is a matrix of local correlations having one row per grid point and one column per \emph{pair of variables} (the columns correspond to the rows in \code{density\_object\$pairs}), as well as \code{\$f\_est}, which is a vector containing the estimate \(\widehat f_X(x)\) of the joint density \(f_X(x)\) in the grid points specified in \code{x0}. The estimated local correlations for this example is plotted in Figure \ref{fig:lgc-plot1}, and the corresponding density estimate is plotted (along the diagonal \(x_1=x_2=x_3=x_4=x\)) in Figure \ref{fig:lgc-plot2}.

The list \code{density\_object} contains the estimated standard deviations of the local correlations in \code{\$loc\_cor\_sd}, as well
as lower and upper confidence bands \code{\$loc\_cor\_lower} and \code{\$loc\_cor\_upper} at the 95\% level. We refer to Table \ref{tab:dlg} for a complete overview of the arguments to \code{dlg()}.

Note that the configuration \code{transform\_to\_marginal\_normality\ =\ TRUE} and \code{est\_method\ =\ 5par} in the bivariate case coincides with the situation considered by \citet{tjos:huft:2013}. In that case, \code{dlg()} serves as a wrapper for the function \code{localgauss()} in the \CRANpkg{localgauss}-package \citep{bere:klep:tjos:2014}.

Obtaining the estimate of a conditional density using the \citet{otne:tjos:2018} algorithm described in Section \ref{chap:background} is very similar, but one must take particular care of the \emph{ordering} of the variables in the data set. The estimation function, \code{clg()}, will \emph{always} assume that the free variables come first, and that the conditioning variables come last. Let us illustrate this in the following code chunk by estimating the joint conditional density of S\&P500 and FTSE100, given that DAX30 = TOPIX = 0.

\renewcommand{\arraystretch}{1.2}
\begin{table}[t!]
\centering
\begin{tabular}{lll}
\toprule
Argument & Explanation & Default value \\
\midrule
\code{lg{\textunderscore}object} & The lg-object created by \code{lg{\textunderscore}main()} & \\
\code{grid} & The evaluation points for the LGC & \code{NULL} \\
\code{level} & Level for confidence bands & 0.95\\ 
\code{normalization} & The estimated density does not integrate to one by & \\ \code{{  \textunderscore}points} & construction. \code{dlg()} will generate the given number   & \\ & of normal variables, having  the same moments as   & \\ &  the data, approximate  $\int \widehat f_X(x)\,\textrm{d}x$ by a Monte Carlo  & \\ & integral, and then normalize the density estimate  & \\ & accordingly & \\ \code{NULL} \\
\code{bootstrap} & Calculate bootstrapped confidence intervals instead& \\ &  of asymptotic expressions & \code{FALSE} \\
\code{B} & Number of bootstrap replicates & \code{500} \\
\bottomrule
\end{tabular}
\caption{Arguments to the \code{dlg()}-function}
\label{tab:dlg}
\end{table}

\begin{example}
R> # We must make sure that the free variables come first
R> returns1 <- stock_data %>% select(SP500, FTSE100, DAX30, TOPIX)
R> 
R> # Create the lg-object
R> lg_object <- lg_main(returns1,
R+                      est_method = "1par",
R+                      bw_method = "plugin",
R+                      transform_to_marginal_normality = TRUE)
R> 
R> # Create a grid
R> x0 <- returns1 %>% 
R+   select(SP500, FTSE100) %>% 
R+   apply(2, function(y) seq(from = -7,
R+                            to = 7,
R+                            length.out = grid_size))
R> 
R> # Calculate the conditional density
R> cond_density <- clg(lg_object, grid = x0, cond = c(0, 0))
\end{example}

The key argument in the call to \code{clg()} above is \code{cond\ =\ c(0,\ 0)}. This means that the last two variables are conditioning variables (and hence that the first \(4-2=2\) variables are free), and that the value of the conditioning variables are fixed at DAX30 = 0 and TOPIX = 0 respectively. This also means that the number of columns in the grid \code{x0} plus the number of elements in \code{cond} must equal the number of variables \(p\) in the data set, and the call to \code{clg()} will result in an error message if this requirement is not fulfilled. Apart from \code{cond}, the \code{clg()}-function takes exactly the same arguments as \code{dlg()} listed in Table \ref{tab:dlg}, and the conditional density estimate in our example is available in the vector \code{cond\_density\$f\_est}.

\subsection{Tests for independence}
\label{chap:indtests}

Three independence tests based on the LGC have appeared in the literature thus far:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  A test for independence between the stochastic variables \(X_1\) and \(X_2\) based on iid data, cf. \citet{bere:tjos:2014}.
\item
  A test for serial independence between \(X_t\) and \(X_{t-k}\) within a time series \(\{X_t\}\), cf. \citet{laca:tjos:2017}. 
\item
  A test for serial cross-dependence between \(X_t\) and \(Y_{t-k}\) for two time series \(\{X_t\}\) and \(\{Y_t\}\), cf. \citet{laca:tjos:2018}.
\end{enumerate}
As we noted in Section \ref{chap:background}, their practical implementations are very similar, and the \CRANpkg{lg}-package provides the function \code{ind\_test()} to perform the tests. Let us first consider the iid case, and generate 500 observations \code{test\_x} from the well known parabola model \(X_2 = X_1^2 + \varepsilon\), where both \(X_1\) and \(\varepsilon\) are independent and standard normal. In this case, \(X_1\) and \(X_2\) are uncorrelated, but obviously strongly dependent. \citet{bere:tjos:2014} considers mainly the full bivariate fit to the raw data, which we easily encode into the lg-object as before. The implementation of the test using 100 bootstrap replicates is shown below.

\begin{example}
R> lg_object <- lg_main(test_x, 
R+                      est_method = "5par",
R+                      transform_to_marginal_normality = TRUE)
R> test_result <- ind_test(lg_object, n_rep = 100)
R> test_result$p_value

[1] 0
\end{example}
This may take a few minutes due to the bootstrapping. We can further specify the function \(h\) and the integration area \(S\) in the test
statistic \eqref{eq:test-statistic1}, see Table \ref{tab:indtest} for details.

\renewcommand{\arraystretch}{1.2}
\begin{table}[t!]
\centering
\begin{tabular}{lll}
\toprule
Argument & Explanation & Default value \\
\midrule
\code{lg{\textunderscore}object} & The lg-object created by \code{lg{\textunderscore}main()} & \\
\code{h} & The function $h(\cdot)$ in (\ref{eq:test-statistic1}) & \code{function(x) x{\textasciicircum}2} \\
\code{S} & The integration area $S$ in (\ref{eq:test-statistic1}). Must & \code{function(x)} \\ & be a logical function on potential  & \code{  as.logical(rep(1,} \\ & grid points in $\mathbb{R}^2$ & \code{    nrow(x)))} \\
\code{bootstrap} & The bootstrap method, must & \\ \code{  {\textunderscore}type} & be either \code{"plain"}, \code{"block"} or \code{"stationary"} & \code{"plain"} \\
\code{block} & Block length for the block bootstrap, &\\ \code{  {\textunderscore}length}& mean block length for the stationary &\\& bootstrap. Calculated by \code{np::b.star()} &\\& \citep{hayf:raci:2008} if not provided & \code{NULL} \\
\code{n{\textunderscore}rep} & Number of bootstrap replicates & 1000 \\
\bottomrule
\end{tabular}
\caption{Arguments to the \code{ind{\textunderscore}test()}-function}
\label{tab:indtest}
\end{table}

The only difference when testing for serial independence within a time series \(\{X_t\}\) is to create a two-column data set consisting of
\(X_t\) and \(X_{t-k}\). For example, if we wish to perform this test for \(k=1\) for one of the variables in the stock-exchange series,
create the matrix of observations as below, and proceed exactly as in the iid case.

\begin{example}
R> returns2 <- stock_data %>% select(SP500) %>% 
R+   mutate(sp500_lagged = lag(SP500))
\end{example}

Finally, the only thing that we must alter in order to perform the third test for serial cross-dependence is the bootstrap method. \citet{laca:tjos:2018} consider two options here, the block bootstrap with either fixed \citep{kuns:1989} or random \citep{poli:roma:1994} block sizes, and this choice is specified by choosing \code{bootstrap\_type\ =\ "block"} or \code{bootstrap\_type\ =\ "stationary"}, respectively, in the call to \code{ind\_test()}.

\subsection{Test for financial contagion}\label{chap:cont}

Assume that we observe two financial time series \(\{X_t\}\) and \(\{Y_t\}\) at times \(t = 1,\ldots,T\), and that some crisis occurs at time \(T^* < T\). As described in Section \ref{chap:background}, \citet{stov:tjos:huft:2014} measure the local correlation between \(\{X_t\}\) and \(\{Y_t\}\) \emph{before} and \emph{after} \(T^*\), and take significant differences between these measurements as evidence against the null hypothesis of no linkage, or contagion, between the time series. In order to implement this test using the \CRANpkg{lg}-package, one must create \emph{two} lg-objects: one for the observations covering the non crisis period and one covering the crisis period. We do not enter a discussion here how to empirically identify such time periods; this is a job that must be done by the practitioner before performing the statistical test.

Let us illustrate the implementation of this test by looking at the same financial returns data that we have used in preceding sections, but this time we will, in the spirit of \citet{stov:tjos:huft:2014}, concentrate on \emph{GARCH(1,1)-filtrated daily returns on the S\&P500 and FTSE100 indices from 2 January 1985 to 29 April 1987} in order to test for financial contagion between the US and UK stock markets following the global stock market crash of 19 October 1987 (``Black Monday''). Assume that these observations are loaded into the R workspace as the \(n_1 \times 2\) data frame \code{x\_nc} containing the observations covering the \(n_1 = 728\) days preceding the crisis , and the \(n_2 \times 2\) data frame \code{x\_c} containing the observations covering the \(n_2 = 140\) consecutive trading days starting on \emph{Black Monday} (see the online code supplement for details concerning the GARCH-filtration and data processing). In the code below we construct one lg-object for each of these data frames with configuration matching the setup used by \citet{stov:tjos:huft:2014}, and perform the test by means of the \code{cont\_test()}-function.

This function returns a list containing the estimated \(p\)-value as well as other useful statistics, including the empirical local correlations measured in the two time periods. See Table \ref{tab:arguments-cont-test} for details concerning other arguments that may be passed to this function.

\renewcommand{\arraystretch}{1.2}
\begin{table}[t!]
\centering
\begin{tabular}{lll}
\toprule
Argument & Explanation & Default value \\
\midrule
\code{lg{\textunderscore}object{\textunderscore}nc} & The lg-object covering the non-crisis period & \\
\code{lg{\textunderscore}object{\textunderscore}c} & The lg-object covering the crisis period & \\
\code{grid{\textunderscore}range} & Range of diagonal for measuring the LGC & (5\%, 95\%) quantiles \\
\code{grid{\textunderscore}length} & The number of grid points to use & 30 \\
\code{n{\textunderscore}rep} & Number of bootstrap replicates & 1000 \\
\code{weight} & Weight function & \code{function(y)} \\ 
&& \code{  rep(1, nrow(y))} \\
\bottomrule
\end{tabular}
\caption{Arguments to the \code{cont{\textunderscore}test()-function}}
\label{tab:arguments-cont-test}
\end{table}

\begin{example}
R> # Create the two lg-objects
R> lg_object_nc <- lg_main(x_nc, 
R+                         est_method = "5par",
R+                         transform_to_marginal_normality = FALSE)
R> 
R> lg_object_c <- lg_main(x_c, 
R+                         est_method = "5par",
R+                         transform_to_marginal_normality = FALSE)
R> 
R> # Run the test with a limited number of bootstrap replicates for 
R> # demonstration purposes.
R> result <- cont_test(lg_object_nc, lg_object_c, n_rep = 20)
R> 
R> # Print out the p-value
R> result$p_value

[1] 0
\end{example}

\subsection{Partial local correlation}
\label{chap:partial}

Consider finally the work by \citet{otne:tjos:2019} who take the off-diagonal element in the local correlation matrix corresponding to the local conditional covariance matrix \eqref{eq:cond-sig} or \eqref{eq:cond-sig-trans} as a local measure of conditional dependence between two stochastic variables \(X^{(1)}\) and \(X^{(2)}\) given the stochastic vector \(X^{(3)}\), and furthermore, in the case with data having been transformed to marginal standard normality, take the statistic
\[T^{CI}_{n, b} = \int h(\widehat\alpha(z)) \, \textrm{d}F_n(z)\]
as a measure of \emph{global} conditional dependence. To this end, the \CRANpkg{lg}-package provides two functions. The first, \code{partial\_cor()}, calculates the local partial correlations as well as their estimated standard deviations, on a specified grid in the \((x^{(1)}, x^{(2)}, x^{(3)})\)-space, and is essentially a wrapper function for the \code{clg()}-function presented in Section \ref{chap:density}. See Table \ref{tab:arguments-partial-cor} for details. The second function, \code{ci\_test()}, performs a test for conditional independence between the first two variables in a data set given the remaining variables.

\renewcommand{\arraystretch}{1.2}
\begin{table}[p]
\centering
\begin{tabular}{lll}
\toprule
Argument & Explanation & Default value \\
\midrule
\texttt{lg{\textunderscore}object} & The lg-object created by \texttt{lg{\textunderscore}main()}  & \\
\texttt{grid} & The evaluation points for the LGPC, \textbf{must} be a & \\
& \qquad  data frame or matrix having 2 columns  & \texttt{NULL} \\
\texttt{cond} & Vector with fixed values for $X^{(3)}$ & \texttt{NULL} \\
\texttt{level} & Significance level for approximated confidence bands & 0.95 \\
\bottomrule
\end{tabular}
\caption{Arguments to the \texttt{partial{\textunderscore}cor()-function}}
\label{tab:arguments-partial-cor}
\end{table}

\begin{table}[p]
\centering
\begin{tabular}{lll}
\toprule
Argument & Explanation & Default value \\
\midrule
\texttt{lg{\textunderscore}object} & The lg-object created by \texttt{lg{\textunderscore}main()}  & \\
\texttt{h} & The function $h(\cdot)$ in (\ref{eq:ci-test-statistic}) & \texttt{function(x) x{\textasciicircum}2} \\
\texttt{n{\textunderscore}rep} & Number of bootstrap replicates & 500 \\
\bottomrule
\end{tabular}
\caption{Arguments to the \texttt{ci{\textunderscore}test()-function}}
\label{tab:arguments-ci-test}
\end{table}

It is well known in the econometrics literature that conditional independence tests are instrumental in the empirical detection of Granger causality \citep{gran:1980}. For example, if we continue to concentrate on the monthly stock returns data that we have already loaded into memory, we may test whether 
\begin{equation} 
\textrm{H}_0: \qquad R_{\textrm{FTSE100}, t} \perp R_{\textrm{SP500}, t-1} \,\, | \,\, R_{\textrm{FTSE100}, t-1}
\label{eq:cond-ind-test}
\end{equation} 
in the period starting in January 2009, the converse of which is a sufficient, but not necessary, condition for \(R_{\textrm{SP500}, t}\) Granger causing \(R_{\textrm{FTSE100}, t}\). We perform the test by running the code below, where \code{x} is a data frame having the following columns strictly ordered as \(R_{\textrm{FTSE100}, t}\), \(R_{\textrm{SP500}, t-1}\) and \(R_{\textrm{FTSE100}, t-1}\) (see the online supplement for the pre-processing of data).

The critical values of this test are calculated using the bootstrap under the null hypothesis by independently resampling replicates from
the conditional density estimates \(\widehat f_{X^{(1)|X^{(3)}}}(x^{(1)}|x^{(3)})\) and \(\widehat f_{X^{(2)|X^{(3)}}}(x^{(2)}|x^{(3)})\), as obtained by the \code{clg()}-function, using an approximated accept-reject algorithm. In order to avoid excessive optimization of the local likelihood function \eqref{eq:loclik}, we estimate \(f_{X^{(1)|X^{(3)}}}\) and \(f_{X^{(2)|X^{(3)}}}\) on the univariate regular grids \(x_0^{(1)}\) and \(x_0^{(2)}\) respectively (while keeping \(x^{(3)}\) fixed at the observed values of \(X^{(3)}\)), and produce interpolating functions \(\tilde f_{X^{(1)|X^{(3)}}}\) and \(\tilde f_{X^{(2)|X^{(3)}}}\) using cubic splines. It is much less computationally intensive to generate replicates from \(\tilde f\) than directly from \(\widehat f\).

We refer to the documentation of the \CRANpkg{lg}-package for details on how to finely tune the behavior of the bootstrapping algorithm by altering the arguments of the \code{ci\_test()}-function, and rather describe the arguments most suitable for modifications by the user in Table \ref{tab:arguments-ci-test}.

\begin{example}
R> # Create the lg-object
R> lg_object <- lg_main(returns4)
R> 
R> # Perform the test
R> test_result <- ci_test(lg_object, n_rep = 100)
R> 
R> # Print out result
R> test_result$p_value

[1] 0.71
\end{example}
The conditional independence test does not provide evidence against the null hypothesis \eqref{eq:cond-ind-test}.

\section{Graphics}\label{chap:graphics}

We conclude this article by describing the \code{corplot()} function, that may be used to draw dependence maps such as those displayed in Figure \ref{fig:initial-example}. \citet{bere:klep:tjos:2014} report on such capabilities in the \CRANpkg{localgauss}-package, but the possibility of creating dependence maps was unfortunately removed from \CRANpkg{localgauss} in the latest version 0.4.0 due to incompatibilities with the \CRANpkg{ggplot2} \citep{wick:2016} plotting engine. We make up for this loss by providing \code{corplot()}, a function that plots the estimated local correlations as provided by \code{dlg()}, or the estimated local \emph{partial} correlations as provided by \code{partial\_cor()}.

The plotting function is highly customizable and provides a number of options covering most basic graphical options. Users well versed in the \CRANpkg{ggplot2}-package may also modify the graphical object returned by corplot in the standard way by adding layers as demonstrate in the example below.

In the first example we generate a set of bivariate normally distributed data using the \CRANpkg{mvtnorm}-package \citep{gentz:etal:2018} and estimate the local Gaussian correlation on a regular \code{grid} using the \code{dlg()}-function. Passing the resulting \code{dlg\_object} to \code{corplot()} without further arguments results in Figure \ref{fig:graphics1}.

\begin{example}
R> grid <- expand.grid(seq(-3, 3, length.out = 7),
R+                     seq(-3, 3, length.out = 7))
R> 
R> x <- mvtnorm::rmvnorm(500, sigma = matrix(c(1, rho, rho, 1), 2))
R> lg_object <- lg_main(x, 
R+                      est_method = "5par", 
R+                      transform_to_marginal_normality = FALSE,
R+                      plugin_constant_joint = 4)
R> dlg_object <- dlg(lg_object, grid = grid)
R> 
R> # Make a dependence map using default setup
R> corplot(dlg_object1)
\end{example}

We may tweak the appearance of our dependence map by passing further arguments to \code{corplot()}. Some of the options are demonstrated in the code chunk below, in which we, for example, superimpose the observations (by setting \code{plot\_obs\ =\ TRUE)} as well as preventing the the estimated local correlations from being plotted in areas without data. The latter option is available through the argument \code{plot\_thres}, which works by calculating a bivariate kernel density estimate \(\tilde f(x_1, x_2)\) for the pair of  variables in question, and only allowing \(\widehat \rho(x_1,x_2)\) to be plotted if \(\tilde f(x_1,x_2)/\max\tilde f(\cdot) >\) \code{plot\_thres}. Adding layers to a dependence map using the ordinary \CRANpkg{ggplot2} syntax works as well, which we demonstrate in Figure \ref{fig:graphics2} by changing the \CRANpkg{ggplot2} theme.

The plotting function works in the same way when plotting the local partial correlations returned by \code{partial\_cor()}, and the arguments of \code{corplot()} are summarized in Table \ref{tab:arguments-corplot}.

\renewcommand{\arraystretch}{1.2}
\begin{table}[p]
\centering
\begin{tabular}{lll}
\toprule
Argument & Explanation & Default value \\
\midrule
\code{dlg{\textunderscore}object} & The object created by \code{dlg()} or \code{partial{\textunderscore}cor()}  & \\
\code{pair} & Which pair to plot if more than two variables & 1 \\
\code{gaussian{\textunderscore}scale} & Logical. Plot on the marginal st. normal scale? & \code{FALSE} \\
\code{plot{\textunderscore}colormap} & Logical. Plot the colormap? & \code{TRUE} \\
\code{plot{\textunderscore}obs} & Logical. Superimpose observations? & \code{FALSE} \\
\code{plot{\textunderscore}labels} & Logical. Plot labels on dependence map? & \code{TRUE} \\
\code{plot{\textunderscore}legend} & Logical. Add legend? & \code{FALSE} \\
\code{plot{\textunderscore}thres} & Threshold for plotting the estimated LGC & 0 \\
\code{alpha{\textunderscore}tile} & Transparency of color tiles & 0.8 \\
\code{alpha{\textunderscore}point} & Transparency of points & 0.8 \\
\code{low{\textunderscore}color} & Color representing $\widehat\rho = -1$ & \code{"blue"}\\
\code{high{\textunderscore}color} & Color representing $\widehat\rho = +1$ & \code{"red"}\\
\code{break{\textunderscore}int} & Break interval for color coding & 0.2 \\
\code{label{\textunderscore}size} & Size of labels in plot & 3 \\
\code{font{\textunderscore}family} & Font family for labels & \code{"sans"} \\
\code{point{\textunderscore}size} & Size of points, if plotted & \code{NULL} \\
\code{xlim}, \code{ylim} & Axis limits & \code{NULL} \\
\code{xlab}, \code{ylab} & Axis labels & \code{NULL} \\
\code{rholab} & Title of legend & \code{NULL} \\
\code{main}, \code{subtitle} & Title and subtitle of plot & \code{NULL} \\
\bottomrule
\end{tabular}
\caption{Arguments to the \code{corplot()}-function}
\label{tab:arguments-corplot}
\end{table}

\begin{example}
R> corplot(dlg_object1, 
R+         plot_obs = TRUE, 
R+         plot_thres = 0.01,
R+         plot_labels = FALSE,
R+         alpha_point = 0.3,
R+         main = "",
R+         xlab = "",
R+         ylab = "") +
R+   theme_classic()
\end{example}

\begin{figure}[p]
\centering 
\includegraphics[width = \textwidth]{graphics1} 
\caption{Dependence map produced by \code{corplot()} using the default configuration}
\label{fig:graphics1}
\end{figure}

\begin{figure}[p]
\centering 
\includegraphics[width = \textwidth]{graphics2} 
\caption{Dependence map produced by tweaking the arguments of \code{corplot()}}
\label{fig:graphics2}
\end{figure}

\bibliography{otneim}

\address{H{\aa}kon Otneim\\
  Department of Business and Management Science, \\
  NHH Norwegian School of Economics, \\
  Helleveien 30, 5045 BERGEN\\
  Norway\\
  \email{hakon.otneim@nhh.no}}

### 2021-02-08 otneim.R

@@ -1,0 +1,370 @@
+
+# Load packages ------
+library(lg)
+library(ggplot2)
+library(lubridate)
+library(tidyr)
+library(fGarch)
+library(dplyr)
+library(extrafont)
+loadfonts(quiet = TRUE)
+
+# Initial example 1 --------
+
+# Estimated local Gaussian correlation between two jointly normal variables 
+# having correlation equal to 0.5, using 1000 observations.
+
+set.seed(1)
+n   <- 1000
+rho <- .5
+
+# Simulated Gaussian data
+grid1 <- expand.grid(seq(-3, 3, length.out = 7),
+                     seq(-3, 3, length.out = 7))
+
+x1 <- mvtnorm::rmvnorm(500, sigma = matrix(c(1, rho, rho, 1), 2))
+lg_object1 <- lg_main(x1, 
+                      est_method = "5par", 
+                      transform_to_marginal_normality = FALSE,
+                      plugin_constant_joint = 4)
+dlg_object1 <- dlg(lg_object1, grid = grid1)
+
+pdf(file = "gaussian-example.pdf", height = 4, width = 5, family = "CM Roman")
+corplot(dlg_object1, plot_thres = .01,
+        xlab = "X1", ylab = "X2", 
+        main = "",
+        plot_obs = TRUE,
+        alpha_point = .1, 
+        plot_legend = FALSE,
+        label_size = 4) +
+    theme_classic() +
+    theme(legend.position = "none")
+dev.off()
+embed_fonts("gaussian-example.pdf")
+
+# Initial example 2 --------
+
+# Plotting the local dependence maps for the returns data
+x2 <- read.csv2("G5_daily_r.csv") %>% 
+    as_tibble %>% 
+    tail(n) %>% 
+    select(CAC40, FTSE100) 
+
+grid2 <- expand.grid(seq(-3.5, 3.5, length.out = 7),
+                     seq(-3, 3, length.out = 7))
+
+lg_object2 <- lg_main(x2, 
+                      est_method = "5par", 
+                      transform_to_marginal_normality = FALSE,
+                      plugin_constant_joint = 2)
+dlg_object2 <- dlg(lg_object2, grid = grid2)
+
+pdf(file = "returns-example.pdf", height = 4, width = 5, family = "CM Roman")
+corplot(dlg_object2, plot_thres = .0001,
+        xlab = "CAC40", ylab = "FTSE100", 
+        main = "",
+        plot_obs = TRUE,
+        alpha_point = .1, 
[…]

### 2021-02-12 response/response.Rmd
Revised section on statistical background, started response

---
title: ""
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\begin{center}
\Large \textbf{Response to reviewer comments to the manuscript \emph{2019-146: lg: An R package for Local Gaussian Approximations} submitted to \emph{The R Journal}}
\end{center}

I am very grateful to the Editor of *The R Journal* for considering this paper for publication, and to the reviewer for his/her careful reading and several useful comments. I have revised the paper according to the comments, as detailed in a point-by-point fashion below.

> *Summary*
>
> *The aim of this paper to introduce and to illustrate the use of the lg package for R which implements the computation and visualization of the concept of ‘local Gaussian correlation’. The local Gaussian correlation (LGC) is a measure of linear and non-linear dependence, proposed recently by Tjøstheim and Hufthammer (2013). Specifically, the paper presents how the GLC can be estimated from empirical or simulated data and how the ‘dependence structure’ described by the GLC can be visualized. Further, the application of statistical tests for the significance of correlation (i.e. GLC) is demonstrated.*

**My response:** This is a good summary that captures the essence of the paper.

> *Overall evaluation*
>
> *Measuring and visualizing the structure(s) of temporal dependencies and dependencies among variables – or aspects thereof – is highly relevant in all areas and sub-disciplines of statistics (e.g. biometry, econometrics etc.). Therefore, providing a software package that implements the local Gaussian correlation, a measure of linear and non-linear dependencies recently proposed by Tjøstheim and Hufthammer (2013) is very useful since it allows a broader audience to apply this concept in empirical research. In this regard presenting such a software package in a scientific journal seems to further support applied statisticians in using this new concept of dependence. However, the author fails to convince me that the manuscript constitutes a serious scientific piece at this stage which deserves publication in a well respected scientific journal. A number of reasons have lead to this conclusion.*

**My response:** The reviewer is exactly right. There is a lot of interest in the measurement, visualization and modeling of statistical dependence in many empirical disciplines. The purpose of this paper is to make recent developments within local Gaussian approximations that have appeared in the statistical literature in recent years available to a broader audience of practitioners. I appreciate the comments by the reviewer below, and I hope that the revised paper satisfies the high scientific standards of *The R Journal*.

> *To exemplify this, the section on ‘Statistical background’ reads as if compiled from different sources; In particular, the paragraphs are not well connected in my view.*

**My response:** The purpose of this section is to give the reader a brief overview of the statistical methods that have been implemented in the **lg**-package, without going too much in to the technical details at this stage. I have revised this section carefully, with particular emphasis on improving the transitions between the paragraphs.

>  *Further the explanation of the basic idea of the GLC measure in the section ‘Statistical background’ could be improved, thereby making the concept accessible for a broader audience and making this paper more ‘self-contained’.*

**My response:** Thank you for this input. The reviewer is correct to point out that the section on "Statistical background" is rather short. The composition of this section is the result of a careful trade-off. On the one hand I fully agree that a more comprehensive section would be desireable, because that would allow the reader to appreciate the technical nuances within the topic of the LGC in greater detail. On the other hand, it is the purpose of this paper to demonstrate the practical implementation of a rather large and diverse set of statistical methods. It is my conclusion that this section should not be much longer than it is in its current form (approx. 4 pages), which naturally puts a limit on the level of detail. I have instead opted to provide a rather acessible birds-eye view of the relevant literature, which I believe is useful to the readership of *The R Journal*. This section has been heavily revised, however, see my previous comment.

> *Also, the whole manuscript is not well-structured: The section with the illustration of the lg package is called ‘The lg-object’ and more importantly is confusing since, for instance, all three different estimation methods implemented are outlined, although not necessary for the (empirical) example chosen by the author;*

**My reponse:**

> *A section with a summary or concluding remarks is fully missing!*

**My response:**

> *Furthermore, the code chunks presented in the paper could be augmented by comments explaining what the different lines of code do; For example, in the section ‘Graphics’ it is not clear what expand.grid does.*

**My response:**

> *In all, this version of the manuscript, unfortunately, appears to be incomplete and as if submitted under time pressure.*

**My response:**

### 2021-02-12 otneim.tex
Revised section on statistical background, started response

@@
 Consider a random vector $X$ having the unknown probability density function $f_X(x)$.
-Based on a random sample $X_1,\ldots, X_n$ we wish to produce an estimate $\widehat f_X(x)$ of $f_X(x)$, to which end the statistical literature provides an abundance of methods.
-One may, for example, assume that $f_X \in F_{\theta}$, where $F_{\theta} = \{f(x;\theta), \theta \in \Theta\}$ is a family of density functions indexed by some parameter $\theta$ living in a parameter space $\Theta$, and estimate $\theta$ by means of the maximum likelihood method.
-On the other hand, we may estimate $f_X(\cdot)$ directly from the data using the classical kernel density estimator
+It is a standard task to estimate $f_X$ based on a random sample $X_1,\ldots, X_n$, and the statistical literature provides an abundance of methods to accomplish this.
+One may, for example, make the assumption that the unknown density function has a particular parametric form, $f_X \in F_{\theta}$, where $F_{\theta} = \{f(x;\theta), \theta \in \Theta\}$ is a family of probability density functions indexed by some parameter $\theta$, and where $\Theta$ is the parameter space.
+Under this assumption we will typically produce an estimate of the parameter $\theta$, written $\widehat \theta$, using the maximum likelihood method.
+The estimated probability density function is then given as $\widehat f_X(x) = f(x; \widehat\theta)$.
+
+A different approach is to estimate $f_X(\cdot)$ without any prior parametric assumptions. The classical method for non-parametric denstity estimation is the kernel estimator 
 $$\widehat f_X(x) = \frac{1}{nb}\sum_{i=1}^n K\left(\frac{X_i - x}{b}\right),$$
-where $K$ is a symmetric density function (the kernel) and $b$ is a tuning parameter (the bandwidth) controlling the smoothness of $\widehat f_X(\cdot)$.
-This requires very little prior assumptions.
-There is a massive literature on density estimation that include countless variations of the parametric and nonparametric approaches, as well as many other clever and creative techniques.
+where $K$ is a symmetric density function (the kernel) and $b$ is a tuning parameter (the bandwidth) that controls the smoothness of the estimate $\widehat f_X(\cdot)$.
+See \citet{silv:1986} for an introduction to this topic.
+There is also a massive statistical literature on density estimation containing extensions and improvements to the classical methods to be used in various practical situations.
 
 \citet{hjor:jone:1996} provide one such idea.
-They consider a parametric family $F_{\theta}$, but instead of searching for a single parameter value $\theta_0$ for which $f_X(x) = f(x;\theta_0)$ (or approximately so), they rather assert that different members of $F_{\theta}$ may approximate $f_X$, locally, in different parts of its domain.
-In other words, they seek to estimate a parameter \emph{function} $\theta_0(x)$ for which $f_X(x) = f(x;\theta_0(x))$ (or approximately so), and do this by pointwise maximizing a \emph{local likelihood function}:
+They consider a parametric family $F_{\theta}$, but instead of searching for a single parameter value $\theta_0$, for which $f_X(x) = f(x;\theta_0)$ (or approximately so), they rather assert that different members of $F_{\theta}$ may approximate $f_X$ locally in different parts of its domain.
+In other words, they seek to estimate a parameter \emph{function} $\theta_0(x)$ for which $f_X(x) = f(x;\theta_0(x))$ (or approximately so), and do this by maximizing a \emph{local likelihood function} in each point $x$:
 \begin{align}
@@
 
-\citet{tjos:huft:2013} consider the bivariate case $X = (X_1, X_2)$ and take $F_{\theta}$ to be the family of bivariate normal distribution consisting of densities on the form
+\citet{tjos:huft:2013} consider the bivariate case $X = (X_1, X_2)$ and take $F_{\theta}$ to be the family of bivariate normal distributions consisting of densities on the form
 
@@
 &= \frac{1}{2\pi \sigma_1 \sigma_2\sqrt{1-\rho^2}} \nonumber \\
-& \qquad \times\exp \left\{-\frac{1}{2(1-\rho^2)}\left(\frac{(x_1-\mu_1)^2}{\sigma_1^2}-2\rho\frac{(x_1-\mu_1)(x_2-\mu_2)}{\sigma_1 \sigma_2}+\frac{(x_2-\mu_2)^2}{\sigma_2^2}\right)\right\}.
+& \qquad \times\exp \left\{-\frac{1}{2(1-\rho^2)}\left(\frac{(x_1-\mu_1)^2}{\sigma_1^2}-2\rho\frac{(x_1-\mu_1)(x_2-\mu_2)}{\sigma_1 \sigma_2}+\frac{(x_2-\mu_2)^2}{\sigma_2^2}\right)\right\},
 \label{eq:gaussian}
 \end{align}
-Using a sample $\{X_{1i}, X_{2i}\}$, $i = 1, \ldots, n$, they estimate 
-$$\widehat\theta(x) = (\widehat\mu_1(x), \widehat\mu_2(x), \widehat\sigma_1(x), \widehat\sigma_2(x), \widehat\rho(x))$$
+where $\theta = (\mu_1,\mu_2,\sigma_1,\sigma_2,\rho)$ is the vector of parameters.
+Using a sample $\{X_{1i}, X_{2i}\}$, $i = 1, \ldots, n$, they estimate $\theta$ locally in the point $x$ by maximizing the local likelihood function \eqref{eq:loclik}, producing
+$$\widehat\theta(x) = (\widehat\mu_1(x), \widehat\mu_2(x), \widehat\sigma_1(x), \widehat\sigma_2(x), \widehat\rho(x)),$$
 and take special interest in the estimated correlation function $\widehat\rho(x)$ (i.e.
 the LGC) because it serves as an attractive local measure of statistical dependence between $X_1$ and $X_2$.
-They show that the LGC reveals many types of nonlinear statistical dependence not seen by the ordinary (global) Pearson's correlation coefficient.
-Furthermore, the LGC distinguishes between positive and negative dependence and reduces to Pearson's $\rho$ if $X_1$ and $X_2$ are jointly normal.
+They show that the LGC reveals many types of nonlinear statistical dependence that are not captured by the ordinary (global) Pearson correlation coefficient.
+Furthermore, the LGC distinguishes between positive and negative dependence and reduces to the Pearson $\rho$ if $X_1$ and $X_2$ are jointly normal.
 We refer to \cite{tjos:huft:2013} for a detailed treatment of the theoretical foundations of the LGC as well as several examples, and rather present two simple illustrations at this point in order to demonstrate the concept.
@@
 In Figure \ref{fig:initial-example} we have plotted the estimated LGC for two bivariate data sets; 1000 simulated observations from a binormal distribution having correlation equal to 0.5, and the daily return on the CAC40 and FTSE100 stock indices on 1000 consecutive trading days starting on May 5th 2014.
-In the first panel, we see that the estimated local correlation coincides with the global Pearson's $\rho$ except for the estimation error.
-In the second panel we see clearly that the local correlation, and thus, as we will see, the dependence, is stronger in the lower left and upper right tails than in central parts of the distribution.
-This phenomenon is of course well known in the financial literature, but can now also be easily measured and quantitatively interpreted by means of the LGC.
+In the first panel, we see that the estimated local correlation coincides with the global correlation, except for the estimation error.
+In the second panel we see clearly that the local correlation, and thus the dependence, is stronger in the lower left and upper right regions of the distribution than in central parts.
+The phenomenon of local dependence is well known in the financial literature, and using the LGC it can be measured, interpreted, and visualized in a natural way.
+The interpretation of this particular figure is that extreme observations on the two stock indices are more strongly dependent than the less extreme observations.
+
 One may obtain these particular estimates from the older \CRANpkg{localgauss}-package (as well as the \CRANpkg{lg}-package of course), but the plotting routine that was used to produce these figures is included in the \CRANpkg{lg}-package and will be described in more detail in Section \ref{chap:graphics}.
@@
 Taking the LGC as a measure of dependence opens up for a number of possibilities to construct statistical tests.
-\citet{bere:tjos:2014} derive from the representation (\ref{eq:gaussian}) that $\rho(x) \equiv 0$ implies that $X_1$ and $X_2$ are independent (the converse statement holds if $\mu_i(x_1, x_2) = \mu_i(x_i)$ and $\sigma_i(x_1, x_2) = \sigma_i(x_i)$ for $i=1,2$, and in general if the observations have been suitably transformed according to a procedure presented later in this section).
-It follows then that departures from $\widehat\rho(x) \equiv 0$ may be taken as evidence against $X_1$ and $X_2$ being statistically independent.
+\citet{bere:tjos:2014} show that $\rho(x) \equiv 0$ implies that $X_1$ and $X_2$ are independent.
+The converse statement holds if $\mu_i(x_1, x_2) = \mu_i(x_i)$ and $\sigma_i(x_1, x_2) = \sigma_i(x_i)$ for $i=1,2$, and in general if the observations have been suitably transformed according to a procedure presented later in this section.
+It follows then that departures from $\widehat\rho(x) \equiv 0$ may be taken as evidence against the hypothesis that $X_1$ and $X_2$ are statistically independent.
 \citet{bere:tjos:2014} formalize this notion by testing whether $\rho(x) \equiv 0$ for all $x\in S \subset \mathbb{R}^2$ using the test statistic
@@
 \end{equation}
-for some function $h$, for example $h(x) = x^2$ or $h(x) = |x|$.
-Critical values may be obtained by permutations of the data under the null hypothesis.
-We demonstrate the implementation of this test using the \CRANpkg{lg}-package in Section \ref{chap:indtests}.
+for some non-negative function $h$, for example $h(x) = x^2$ or $h(x) = |x|$.
+Critical values may be obtained by permutations of the data under the null hypothesis, and we demonstrate the implementation of this test using the \CRANpkg{lg}-package in Section \ref{chap:indtests}.
 
 Consider next the stationary time series $\{X_t\}$.
-The autocorrelation $\rho_k = \rho(X_t, X_{t-k})$ is a fundamental descriptive and analytical concept, but is, again, only capable to completely capture \emph{linear} serial dependence relationships.
-\citet{laca:tjos:2017} seek to remedy this by rather calculating the LGC to measure the \emph{local} dependence between $X_t$ and $X_{t-k}$, which in turn leads to a test for serial independence in the natural way.
-In fact, this work is mainly a theoretical exercise in order to accommodate dependent data.
-Testing for serial independence between $X_t$ and $X_{t-k}$ using observations $\{X_t, X_{t-k}\}_{t = 1}^T$ leads to the same test statistic and bootstrap procedure as the test for independence between $X_1$ and $X_2$ that we described above, and we present its implementation in Section \ref{chap:indtests} as well.
+The autocorrelation function (ACF) $\rho_k = \rho(X_t, X_{t-k})$ is a well known concept for describing the serial dependence in the time seris, but the ACF is, again, only capable to completely capture \emph{linear} serial dependence.
+\citet{laca:tjos:2017} seek to remedy this by rather calculating the the local correlation between $X_t$ and $X_{t-k}$.
+This leads to a test for serial independence in the natural way.
+In fact, this work is mainly a theoretical exercise in order to accommodate time series dependence.
+Testing for independence between $X_t$ and $X_{t-k}$ using observations $\{X_t, X_{t-k}\}_{t = k+1}^T$ leads to the same test statistic \eqref{eq:test-statistic1} and bootstrap procedure as the test for independence between $X_1$ and $X_2$ that we described above.
 
-\citet{laca:tjos:2018} extend this problem to test for serial cross dependence between two time series $\{X_t, Y_t\}$ by measuring the LGC between $X_t$ and $Y_{t-k}$.
-Indeed, departures from $\widehat\rho(x,y) \equiv 0$ is again taken as evidence against independence, and the test statistic (\ref{eq:test-statistic1}) provides an aggregate measure of just that in the specified region $S$.
-This time, however, we can not just obtain permutations of the data in order to produce replicates of the test statistic under the null hypothesis.
-\citet{laca:tjos:2018} accomplish this by rather suggesting two block bootstrap procedures using fixed and random block sizes.
-This test is implemented in the \CRANpkg{lg}-package, as we demonstrate in Section \ref{chap:indtests}.
+\citet{laca:tjos:2018} extend this problem to test for serial cross-dependence between two time series $\{X_t, Y_t\}$ by measuring the LGC between $X_t$ and $Y_{t-k}$.
+Departures from $\widehat\rho(x,y) \equiv 0$ is again taken as evidence against independence, and the test statistic (\ref{eq:test-statistic1}) provides an aggregate measure of this discrepancy in the specified region $S$.
+In this case, however, we can not obtain replicates of the test statistic under the null hypothesis by simple permutations of the data.
+\citet{laca:tjos:2018} suggest rather two block bootstrap procedures to this end, using fixed and random block sizes respectively.
+The tests for serial dependence, and serial cross-dependence are both implemented in the \CRANpkg{lg}-package, as we demonstrate in Section \ref{chap:indtests}.
 
-\citet{stov:tjos:huft:2014} measure and test for financial contagion using the LGC.
-They define contagion by "a significant increase in cross-market linkages after a shock to one country" \citep{forb:rigo:2002} and employ the LGC to quantify such linkage.
-Contrary to the independence tests above, \citet{stov:tjos:huft:2014} estimate the LGC on a grid $\{x_1, x_2\}_k$, $k=1,\ldots,K$ along the diagonal $D=\{(x_1,x_2): x_1 = x_2\}$ \emph{before} and \emph{after} some critical event in the financial markets, i.e.
-in the crisis (C) and non-crisis (NC) periods.
-They compare the two estimates using the test statistic
+We find another application of the local Gaussian approximation in the work by \citet{stov:tjos:huft:2014}, who measure and test for financial contagion.
+They define contagion by "a significant increase in cross-market linkages after a shock to one country" \citep{forb:rigo:2002} and employ the LGC to quantify this potential linkage.
+The authors estimate the LGC on a grid $\{x_1, x_2\}_k$, $k=1,\ldots,K$ along the diagonal $D=\{(x_1,x_2): x_1 = x_2\}$ \emph{before} and \emph{after} some critical event in the financial markets, denoted as the crisis (C) and the non-crisis (NC) periods respectively.
+They compare the two estimates using the following test statistic;
 
@@
 where $w(\cdot, \cdot)$ is a weight function that serve the same purpose as the integration area $S$ in (\ref{eq:test-statistic1}).
-In this case, a standard bootstrap will suffice in order to produce approximate replicates of $T_{n,b}^D$ under the null hypothesis of no financial contagion, and we demonstrate the implementation of this test using the \CRANpkg{lg}-package in Section \ref{chap:cont}.
+In this case, \citet{stov:tjos:huft:2014} show that a standard bootstrap will suffice in order to produce approximate replicates of $T_{n,b}^D$ under the null hypothesis of no financial contagion, and we demonstrate the implementation of this test using the \CRANpkg{lg}-package in Section \ref{chap:cont}.
 
-Although \citet{hjor:jone:1996} provide a general framework for locally parametric density estimation using using any parametric family, having any number of variables, it is evident that the method may struggle in multivariate applications much in the same way as the kernel density estimator does.
+Although the original work by \citet{hjor:jone:1996} provide a general framework for local likelihood density estimation, using any $p$-variate parametric family as the local family, it is evident that the method may struggle in multivariate applications much in the same way as the kernel density estimator does.
 This is a consequence of the curse of dimensionality, the effect of which is sought remedied by an algorithm provided by \citet{otne:tjos:2017}.
-The idea is to locally fit the $p$-variate normal distribution $\psi(\mu, \Sigma)$, where $\mu$ is the vector of $p$ expectations, and $\Sigma$ is the $p\times p$ covariance matrix (to which the correlation matrix $R$ corresponds), but under the following structural simplifications:
+The idea is to fit the $p$-variate normal distribution $\psi(\mu, \Sigma)$ locally, where $\mu$ is the vector of $p$ expectations, and $\Sigma$ is the $p\times p$ covariance matrix (to which the correlation matrix $R$ corresponds), but under the following structural simplifications:
 
@@
 \end{align}
-These local parameters are estimated by univariate marginal locally normal fits (eqs.
-\ref{eq:simp-mu} and \ref{eq:simp-sig}), and pairwise bivariate locally normal fits (eq.
-\ref{eq:simp-rho}), and in the latter case the estimates $\widehat\mu_i(x_i)$, $\widehat\mu_j(x_j)$, $\widehat\sigma_i(x_i)$ and $\widehat\sigma_j(x_j)$ are kept fixed.
-
+\citet{otne:tjos:2017} estimate the local parameters above by first obtaining univariate marginal locally Gaussian fits (eqs.
+\ref{eq:simp-mu} and \ref{eq:simp-sig}), and then pairwise bivariate locally Gaussian fits (eq.
+\ref{eq:simp-rho}).
+In the second step the estimates $\widehat\mu_i(x_i)$, $\widehat\mu_j(x_j)$, $\widehat\sigma_i(x_i)$ and $\widehat\sigma_j(x_j)$ are kept fixed in the estimation of the pairwise local correlation.
+They argue further that the following transformation technique will produce better density estimates in many situations.
 Denote by $F_i(x_i)$, $i = 1,\ldots, p$ the marginal distribution functions of the stochastic vector $X$, and by $\widehat F_i(x_i) = n^{-1}\sum_{i=1}^n 1(X_i \leq x_i)$ their empirical counterparts.
-\citet{otne:tjos:2017} argue that estimating the density $f_Z(z)$ of the vector $Z = \{\Phi^{-1}(F_i(X_i))\}_{i=1,\ldots,p}$, in practice approximated by
+They then estimate the density $f_Z(z)$ of the vector $Z = \{\Phi^{-1}(F_i(X_i))\}_{i=1,\ldots,p}$, in practice approximated by
 
@@
 \end{equation}
-where $\Psi(\cdot)$ is the univariate standard normal cdf, will lead to better density estimates in many situations. In that case, one may simplify even further and fix
+and where $\Psi(\cdot)$ is the univariate standard normal cdf. In that case, they simplify the estimation problem even further and fix
 
@@
 \end{equation}
-so that the only parameter functions left to estimate are the pairwise local Gaussian correlations $R(z) = \{\rho_{ij}(z_i, z_j)\}_{i<j}$, where we use the notation $Z$, $z_i$ and $z_j$ to signify that the estimation is performed on the (approximate) standard normal scale (or $Z$-scale).
-One then estimates the joint density $f_Z(z)$ of $Z$ as
+so that the only parameter functions left to estimate are the pairwise local Gaussian correlations $R(z) = \{\rho_{ij}(z_i, z_j)\}_{i<j}$.
+We use the notation $Z$, $z_i$ and $z_j$ to signify that the estimation is performed on the (approximate) standard normal scale, or $z$-scale for short.
+We can then estimate the joint density $f_Z(z)$ of $Z$ as
 \begin{equation}
@@
 \end{equation}
-where $\phi(\cdot)$ is the standard normal pdf. This estimator is implemented the \CRANpkg{lg}-package as described in Section \ref{chap:density}.
+where $\phi(\cdot)$ is the standard normal pdf. This estimator is implemented the \CRANpkg{lg}-package as demonstrated in Section \ref{chap:density}.
 
-One particular feature enjoyed by the jointly normally distributed vector $X$ is that, for any partitioning $\left(X^{(1)}, X^{(2)}\right)$, the conditional distribution of $X^{(1)}|X^{(2)} = x^{(2)}$ is also normal.
+One particular feature enjoyed by the jointly normally distributed vector $X$ is that, for any partitioning $X = \left(X^{(1)}, X^{(2)}\right)$, the conditional distribution of $X^{(1)}|X^{(2)} = x^{(2)}$ is also normal.
 In fact, if $X \sim \mathcal{N}(\mu,\Sigma)$, and $\mu$ and $\Sigma$ is partitioned according to $\left(X^{(1)}, X^{(2)}\right)$ as
@@
 \citet{otne:tjos:2018} demonstrate that this property may be translated into a corresponding local argument without modification.
-If the joint density $f_X(\cdot)$ is locally Gaussian, i.e.
-if
+That is, if the joint density $f_X(\cdot)$ can be written on a locally Gaussian form
 
@@
 
-then $X^{(1)}|X^{(2)} = x^{(2)}$ is also locally Gaussian with local parameters given by equations (\ref{eq:cond-mu}) and (\ref{eq:cond-sig}), except that all quantities are $x$-dependent.
+then the conditional density of $X^{(1)}|X^{(2)} = x^{(2)}$ can also be written on the same locally Gaussian form with local parameters given by equations (\ref{eq:cond-mu}) and (\ref{eq:cond-sig}), except that all quantities are $x$-dependent.
 If we use the transformation technique described above together with simplification (\ref{eq:simp}), the local versions of equations (\ref{eq:cond-mu}) and (\ref{eq:cond-sig}) simplify to
@@
 where we, again, switch to $z$-notation in order to make it clear that these quantities are estimated on the standard normal $z$-scale.
-An estimator $\widehat f_{X^{(1)}|X^{(2)}}(\cdot|\cdot)$ of the conditional density $f_{X^{(1)}|X^{(2)}}(\cdot|\cdot)$ follows immediately from an expression corresponding to (\ref{eq:backtrans}), and the \CRANpkg{lg}-package provides functions for implementing this estimator in R as described in Section \ref{chap:density}.
+An estimator $\widehat f_{X^{(1)}|X^{(2)}}(\cdot|\cdot)$ of the conditional density $f_{X^{(1)}|X^{(2)}}(\cdot|\cdot)$ follows immediately from an expression corresponding to (\ref{eq:backtrans}), and the \CRANpkg{lg}-package provides functions for implementing this estimator in R, and we describe the implementation of this functionality in Section \ref{chap:density}.
 
 Finally, we refer to \citet{otne:tjos:2019}, who take the local version of the conditional covariance matrix (\ref{eq:cond-sig}) (or (\ref{eq:cond-sig-trans}) in the transformed case) as a measure for conditional dependence, and thus as an instrument to test for conditional independence.
-Specifically, consider the stochastic vector $X = \left(X^{(1)}, X^{(2)}, X^{(3)}\right)$, where $X^{(1)}$ and $X^{(2)}$ are scalars and $X^{(3)}$ may be a vector.
-$X^{(1)}$ is conditionally independent from $X^{(2)}$ given $X^{(3)}$, written $X^{(1)} \perp X^{(2)} \,|\,X^{(3)}$ if the stochastic variables $X^{(1)}\,|\,X^{(3)}$ and $X^{(2)}\,|\,X^{(3)}$ are independent, or, equivalently, if the joint conditional density of $X^{(1)}$ and $X^{(2)}$ given $X^{(3)}$ can be written as the product
+Consider the stochastic vector $X = \left(X^{(1)}, X^{(2)}, X^{(3)}\right)$, where $X^{(1)}$ and $X^{(2)}$ are scalars and $X^{(3)}$ may be a vector.
+$X^{(1)}$ is conditionally independent from $X^{(2)}$ given $X^{(3)}$, written $X^{(1)} \perp X^{(2)} \,|\,X^{(3)}$, if the stochastic variables $X^{(1)}\,|\,X^{(3)}$ and $X^{(2)}\,|\,X^{(3)}$ are independent, or, equivalently, if the joint conditional density of $X^{(1)}$ and $X^{(2)}$ given $X^{(3)}$ can be written as the product
 
@@
 In this case, denote by $\alpha(z)$ the off-diagonal element in the $2\times2$ local correlation matrix $R^*(z)$ (which derives directly from $\Sigma^*(z)$ as given in (\ref{eq:cond-sig-trans})).
-If $X$ has a local Gaussian distribution, the conditional independence (\ref{eq:cond-ind}) is equivalent to $\alpha(z) \equiv 0$, departures from which we take as evidence against conditional dependence between $X^{(1)}$ and $X^{(2)}$ given $X^{(3)}$, and this discrepancy is quantified using the test functional
+If $X$ has a local Gaussian distribution, the conditional independence (\ref{eq:cond-ind}) is equivalent to $\alpha(z) \equiv 0$, and \citet{otne:tjos:2019} take departures from this relation as evidence against the hypothesis of conditional independence between $X^{(1)}$ and $X^{(2)}$ given $X^{(3)}$.
+The natural way to quantify this is the test functional
 

### 2021-02-12 otneim.R

@@ -101,7 +101,7 @@ embed_fonts("faithful.pdf")
 
 pdf(file = "faithful-transformed.pdf", height = 4, width = 5, family = "CM Roman")
 lg_object$transformed_data %>% 
-    as_tibble %>% 
+    as.data.frame %>% 
     ggplot +
     geom_point(aes(x = V1, y = V2)) +
     xlab("eruptions") +
@@ -231,7 +231,7 @@ set.seed(1)
 n <- 500
 x1 <- rnorm(n)
 x2 <- x1^2 + rnorm(n)
-test_x <- data_frame(x1 = x1, x2 = x2)
+test_x <- tibble(x1 = x1, x2 = x2)
 
 lg_object <- lg_main(test_x, 
                      est_method = "5par",
@@ -246,7 +246,7 @@ returns2 <- stock_data %>% select(SP500) %>%
 
 # Read the daily data
 returns3 <- read.csv2("G5_daily_r.csv") %>%     
-    as_data_frame  %>% 
+    as_tibble  %>% 
     dplyr::select(Date, SP500, FTSE100) %>%
     drop_na %>% 
     mutate(Date = dmy(Date)) %>%  
@@ -260,11 +260,9 @@ crisis_start <- ymd("1987-10-18")
 crisis_end   <- ymd("1988-04-29")
 
 garch_filtrate <- function(y) {
-    sink("/dev/null")
-    fit <- garchFit(~garch(1,1), 
+    invisible(capture.output(fit <- garchFit(~garch(1,1), 
                     data = y, 
-                    cond.dist = "std") 
-    sink()
+                    cond.dist = "std")))
     fit@residuals/fit@sigma.t
 }
 
@@ -289,7 +287,7 @@ x_nc <- data.frame(US = crash_data %>%
                        dplyr::filter(crisis == "BEFORE") %>% 
                        dplyr::select(return_filtered) %>% 
                        pull) %>% 
-    as_data_frame
+    as_tibble
 
 x_c <- data.frame(US = crash_data %>% 
                       dplyr::filter(exchange == "SP500") %>%
@@ -301,7 +299,7 @@ x_c <- data.frame(US = crash_data %>%
                       dplyr::filter(crisis == "AFTER") %>% 
                       dplyr::select(return_filtered) %>% 
                       pull) %>% 
-    as_data_frame
+    as_tibble
 
 # Create the two lg-objects
 lg_object_nc <- lg_main(x_nc, 

### 2021-02-16 response/response.Rmd
First pass of revision

@@
 
-**My response:** The purpose of this section is to give the reader a brief overview of the statistical methods that have been implemented in the **lg**-package, without going too much in to the technical details at this stage.
+**My response:** The purpose of this section is to give the reader a brief overview of the statistical methods that have been implemented in the **lg**-package, without going too much in to the technical details.
 I have revised this section carefully, with particular emphasis on improving the transitions between the paragraphs.
@@
 **My response:** Thank you for this input.
-The reviewer is correct to point out that the section on "Statistical background" is rather short.
+The reviewer is correct to point out that the section on "Statistical background" is fairly short and does not go into great detail.
 The composition of this section is the result of a careful trade-off.
-On the one hand I fully agree that a more comprehensive section would be desireable, because that would allow the reader to appreciate the technical nuances within the topic of the LGC in greater detail.
-On the other hand, it is the purpose of this paper to demonstrate the practical implementation of a rather large and diverse set of statistical methods.
+On the one hand I fully agree that a more comprehensive section would be desireable, because that would allow the reader to appreciate the technical nuances within the topic of the LGC.
+On the other hand, it is the purpose of this paper to demonstrate the practical implementation of a large and diverse set of statistical methods.
 It is my conclusion that this section should not be much longer than it is in its current form (approx.
@@
 I have instead opted to provide a rather acessible birds-eye view of the relevant literature, which I believe is useful to the readership of *The R Journal*.
-This section has been heavily revised, however, see my previous comment.
+This section has been heavily revised as prompted by the reviewer in his/her comment above..
 
@@
 
-**My reponse:**
+**My reponse:** The paper is structured in the same way as the package it describes.
+There are two steps to every analysis; in the first step the practitioner must make several modeling choices which to a large degree is guided by the respective references.
+This step is in common for all applications.
+I have created a new title for this section to clarify the structure; it is now called "The first step: Creating the lg-object".
+The various applications are then described separately and organized under informative headlines.
+
+I am not entirely sure how to address the reviewer's second remark.
+The examples in this paper are chosen to illustrate key steps in the implementation of the methods.
+It is not my intention to demonstrate all possible combinations of the modeling choices in examples, as that would likely reduce the value of the paper as a general, readable and fairly brief introduction to the relevant R-functions.
 
@@
 
-**My response:**
+**My response:** I have added some conluding remarks in the end of the paper.
 
@@
 
-**My response:**
+**My response:** Thank you for pointing this out. Several of the somewhat longer code chunks were not properly commented. Has been fixed.
 
@@
 
-**My response:**
+**My response:** I can ensure the reviewer and the editor that this is not the case.
+The manuscript has been prepared carefully, under no time pressure whatsoever.
+I do thank the reviewer for several useful comments, which I believe has contributed to a significant improvement in the presentation of the material.
+
+Here are my responses to the specific issues addressed by the reviewer:
+
+> *1.
+As far as I see is your manuscript missing a final section that summarizes the main ideas (e.g.
+functions here) from the paper and provides some concluding remarks and perhaps an outlook (e.g.
+what will be added to package in the future).
+I strongly recommend to provide such a section.*
+
+**My response:** Has been added.
+
+> *2. Are you sure that Equation (1) is correct? I think $x$ should be $x_0$ . Further I do not understand $y$ and assume $y$ should be $X_i$?*
+
+**My response:** Yes, this equation is correct.
+The $x$ is the location of the local estimate as defined in the discussion leading up to equation (1), and $y$ is the integration variable in the penalty term of the local likelihood function.
+This notation is consistent with the literature on local likelihood estimation.
+The second term does in fact not depend directly on the observations, and the reviewer is quite right to point out that this is not obvious.
+I have therefore added a comment about this, as well as a reference to a much more detailed justification of the local likelihood function.
+
+> *3.
+Are you sure that ‘the same test statistic and bootstrap procedure as for independence’ can be used with time series data?
+Is not block-bootstrap necessary in this case?
+Please explain briefly!
+(p.
+3)*
+
+**My response:** In this particular case we can use the standard bootstrap, because serial dependence in a time series can be investigated by considering the statistical dependence between $X_t$ and $X_{t-k}$ directly.
+This is a parallell to the autocorrelation function, which for a stationary time series is defined as the ordinary Pearson correlation $\gamma_k = Corr(X_t, X_{t-k})$.
+I have changed the formulation here so that it directly states that we obtain critical values by permutations.
+The block bootstrap is then introduced in the next paragraph, where it is needed to test for serial cross-dependence.
+
+> *4.
+I understand that the bandwidth controls how many observations are used to compute the correlation coefficient of one grid.
+Does lg also allow to extract the (exact) number of observations used to compute the different correlations coefficients (as depicted e.g.
+in Fig.
+1)?
+I feel that from the perspective of an applied statistician, this figure could provide an intuition about how reliable one or the other coefficient is.*
+
+**My response:** The bandwidth does not define a finite region over which the usual correlation is calculated; this is, it is not the conditional correlation, and the bandwidth does not correspond to an exact number of observations.
+Figure 1 shows the estimated local Gaussian correlation plotted out on a grid chosen by the user, and I have pointed this out in the first line of the paragraph that introduces this figure.
+
+Although one can discuss the "effective number of observations" in the classical nonparametric sense, this has not been done in the literature on the LGC, and hence not a consept in the *lg*-package.
+
+> *5. Is it possible to test for significant differences in the correlation coefficients, for instance, from the lower left, middle and upper right in Fig. 1?*
+
+**My response:** That may certainly be possible.
+To the best of my knowledge, however, this is not a test that has been introduced in the literature.
+It is therefore not implemented in the **lg**-package.
+
+> *6. Page 1: I suggest to change the statement ‘clever and creative techniques’ by a more scientific formulation.*
+
+**My response:** This statement has been taken out. Thank you.

### 2021-02-16 otneim.tex
First pass of revision

@@
 where, again, $K$ is a symmetric density function and $b$ is a bandwidth parameter that controls the smoothness of the estimate.
+The second term in the local likelihood function is a penalty that ensures that the estimated density $\widehat f_X(x) = f(x; \widehat\theta(x))$ converges correctly to the true density function $f_X(x)$ as the sample size $n$ increases to infinity and the bandwidth $b$ decreases towards zero.
+See \citet{hjor:jone:1996} for a detailed discussion about this construction.
 
@@
 
-In Figure \ref{fig:initial-example} we have plotted the estimated LGC for two bivariate data sets; 1000 simulated observations from a binormal distribution having correlation equal to 0.5, and the daily return on the CAC40 and FTSE100 stock indices on 1000 consecutive trading days starting on May 5th 2014.
+In Figure \ref{fig:initial-example} we have plotted the estimated LGC for two bivariate data sets on a grid; 1000 simulated observations from a binormal distribution having correlation equal to 0.5, and the daily return on the CAC40 and FTSE100 stock indices on 1000 consecutive trading days starting on May 5th 2014.
 In the first panel, we see that the estimated local correlation coincides with the global correlation, except for the estimation error.
@@
 
-\section{The \texttt{lg}-object} 
+\section{The first step: Creating the \texttt{lg}-object} 
 \label{chap:lgobject}
@@
 \begin{example}
+R> # Create the lg-object with bandwidths chosen by cross-validation
 R> lg_object <- lg_main(x,
@@
 \begin{example}
+R> # Print out the bandwidths
 R> lg_object$bw
@@
 \begin{example}
+R> # Make the lg-object with plugin bandwidths
 R> lg_object <- lg_main(x,
@@
 R+                      bw_method = "plugin")
+R> # Print out the bandwidths
 R> lg_object$bw
@@
 \begin{example}
+R> # Make the lg-object
 R> lg_object <- lg_main(test_x, 
@@
 R+                      transform_to_marginal_normality = TRUE)
+R> # Perform the independence test
 R> test_result <- ind_test(lg_object, n_rep = 100)
+R> # Print out the p-value of the test
 R> test_result$p_value
@@
 \begin{example}
+R> # Make a regular grid in the domain of the distribution
 R> grid <- expand.grid(seq(-3, 3, length.out = 7),
@@
 
+\section{Conclusion}
+
+The statistical literature has seen a number of applications of local Gaussian approximations in the last decade, covering several topics in dependence modeling and inference, as well as the estimation of multivariate density and conditional density functions.
+In this paper, we demonstrate the implementation of these methods i the R programming language using the \CRANpkg{lg}-package, as well as the graphical representation of the estimated local Gaussian correlation.
+The package is complete, in the sense that all major methods that have been published within this framework is now easily accessible to the practitioner.
+The package is also modular; Future methodological developments using local Gaussian approximations may be added to the package with little effort.
+
 \bibliography{otneim}

### 2021-02-16 otneim.tex
Revised conclusion

@@
 The package is complete, in the sense that all major methods that have been published within this framework is now easily accessible to the practitioner.
-The package is also modular; Future methodological developments using local Gaussian approximations may be added to the package with little effort.
+The package is also designed with a modular infrastructure that allows future methodological developments using local Gaussian approximations to be added to the package.
 

### 2021-02-16 response/response.Rmd
Full pass with spelling of the entire paper and response to reviewer

@@
 The composition of this section is the result of a careful trade-off.
-On the one hand I fully agree that a more comprehensive section would be desireable, because that would allow the reader to appreciate the technical nuances within the topic of the LGC.
+On the one hand I fully agree that a more comprehensive section would be desirable, because that would allow the reader to appreciate the technical nuances within the topic of the LGC in greater detail.
 On the other hand, it is the purpose of this paper to demonstrate the practical implementation of a large and diverse set of statistical methods.
@@
 4 pages), which naturally puts a limit on the level of detail.
-I have instead opted to provide a rather acessible birds-eye view of the relevant literature, which I believe is useful to the readership of *The R Journal*.
-This section has been heavily revised as prompted by the reviewer in his/her comment above..
+I have instead opted to provide a rather accessible birds-eye view of the relevant literature, which I believe is useful to the readership of *The R Journal*.
+This section has been heavily revised as prompted by the reviewer in his/her comment above.
 
@@
 There are two steps to every analysis; in the first step the practitioner must make several modeling choices which to a large degree is guided by the respective references.
-This step is in common for all applications.
-I have created a new title for this section to clarify the structure; it is now called "The first step: Creating the lg-object".
+This step is common to all applications.
+I have created a new title for this section to clarify the structure.
+It is now called "The first step: Creating the lg-object".
 The various applications are then described separately and organized under informative headlines.
@@
 
-**My response:** I have added some conluding remarks in the end of the paper.
+**My response:** I have added some concluding remarks in the end of the paper.
 
@@
 This notation is consistent with the literature on local likelihood estimation.
-The second term does in fact not depend directly on the observations, and the reviewer is quite right to point out that this is not obvious.
+Indeed, the second term does not depend on the data, and the reviewer is quite right to point out that this is not obvious.
 I have therefore added a comment about this, as well as a reference to a much more detailed justification of the local likelihood function.
@@
 **My response:** In this particular case we can use the standard bootstrap, because serial dependence in a time series can be investigated by considering the statistical dependence between $X_t$ and $X_{t-k}$ directly.
-This is a parallell to the autocorrelation function, which for a stationary time series is defined as the ordinary Pearson correlation $\gamma_k = Corr(X_t, X_{t-k})$.
+This is a parallel to the autocorrelation function, which for a stationary time series is defined as the ordinary Pearson correlation $\gamma_k = Corr(X_t, X_{t-k})$.
 I have changed the formulation here so that it directly states that we obtain critical values by permutations.
@@
 
-**My response:** The bandwidth does not define a finite region over which the usual correlation is calculated; this is, it is not the conditional correlation, and the bandwidth does not correspond to an exact number of observations.
+**My response:** The bandwidth does not define a finite region over which the usual correlation is calculated.
+The LGC is not the conditional correlation, and the bandwidth does not correspond to an exact number of observations.
 Figure 1 shows the estimated local Gaussian correlation plotted out on a grid chosen by the user, and I have pointed this out in the first line of the paragraph that introduces this figure.
 
-Although one can discuss the "effective number of observations" in the classical nonparametric sense, this has not been done in the literature on the LGC, and hence not a consept in the *lg*-package.
+Although one can discuss the "effective number of observations" in the classical nonparametric sense, this has not been done in the literature on the LGC, and hence is not a concept in the *lg*-package.
 
@@
 **My response:** This statement has been taken out. Thank you.
+
+I addition to the changes prompted by the reviewer, I have updated a few references and made several minor adjustments to the text throughout the article to improve the presentation.

### 2021-02-16 otneim.tex
Full pass with spelling of the entire paper and response to reviewer

--- 2021-02-16 otneim.tex
@@
 This includes estimation of the local Gaussian correlation itself, multivariate density estimation, conditional density estimation, various tests for independence and conditional independence, as well as a graphical module for creating dependence maps.
-This paper describes the \CRANpkg{lg} package, its principles and its practical use.
+This paper describes the \CRANpkg{lg} package, its principles, and its practical use.
 }
@@
 \section{Introduction}
+\label{chap:introduction}
 
-\citet{tjos:huft:2013} propose the \emph{local Gaussian correlation} (LGC) as a new measure of statistical dependence between two stochastic variables $X_1$ and $X_2$, which has the following important property yet unrivaled in the literature: It can separate between positive and negative nonlinear dependence while still reducing to the ordinary Pearson's correlation coefficient if $X_1$ and $X_2$ are jointly normally distributed.
+\citet{tjos:huft:2013} propose the \emph{local Gaussian correlation} (LGC) as a new measure of statistical dependence between two stochastic variables $X_1$ and $X_2$, which has the following important property yet unrivaled in the literature: It can separate between positive and negative nonlinear dependence while still reducing to the ordinary Pearson correlation coefficient if $X_1$ and $X_2$ are jointly normally distributed.
 The R-package \CRANpkg{localgauss} \citep{bere:klep:tjos:2014} provides two important functions in this context; one that calculates the sample LGC based on observed values of $(X_1,X_2)$, and one that uses the estimated LGC to perform a local test of independence between $X_1$ and $X_2$ as described in detail by \citet{bere:tjos:2014}.
@@
 
-This paper describes the \CRANpkg{lg}-package \citep{otne:2019}, which provides a unified framework to implement all these methods, as well as a tool for visualizing the LGC and LGPC as dependence maps (and conditional dependence maps).
+This paper describes the \CRANpkg{lg}-package \citep{otne:2019}, which provides a unified framework to implement all these methods, as well as a tool for visualizing the LGC and LGPC as dependence maps.
 \citet{jord:tjos:2017} use the LGC in spectral analysis of time series, but those methods have their own computational ecosystem in the \pkg{localgaussSpec}-package, \citet{jord:2018}.
@@
 
-A different approach is to estimate $f_X(\cdot)$ without any prior parametric assumptions. The classical method for non-parametric denstity estimation is the kernel estimator 
+A different approach is to estimate $f_X(\cdot)$ without any prior parametric assumptions. The classical method for non-parametric density estimation is the kernel estimator 
 $$\widehat f_X(x) = \frac{1}{nb}\sum_{i=1}^n K\left(\frac{X_i - x}{b}\right),$$
@@
 where, again, $K$ is a symmetric density function and $b$ is a bandwidth parameter that controls the smoothness of the estimate.
-The second term in the local likelihood function is a penalty that ensures that the estimated density $\widehat f_X(x) = f(x; \widehat\theta(x))$ converges correctly to the true density function $f_X(x)$ as the sample size $n$ increases to infinity and the bandwidth $b$ decreases towards zero.
+The second term in the local likelihood function is a penalty that ensures that the estimated density $\widehat f_X(x) = f(x; \widehat\theta(x))$ converges correctly to the true density function $f_X(x)$ as the sample size $n$ increases to infinity, and the bandwidth $b$ decreases towards zero.
 See \citet{hjor:jone:1996} for a detailed discussion about this construction.
@@
 In the first panel, we see that the estimated local correlation coincides with the global correlation, except for the estimation error.
-In the second panel we see clearly that the local correlation, and thus the dependence, is stronger in the lower left and upper right regions of the distribution than in central parts.
+In the second panel we see clearly that the local correlation, and thus the dependence, is stronger in the lower left and upper right regions of the distribution than in the central parts.
 The phenomenon of local dependence is well known in the financial literature, and using the LGC it can be measured, interpreted, and visualized in a natural way.
@@
 Consider next the stationary time series $\{X_t\}$.
-The autocorrelation function (ACF) $\rho_k = \rho(X_t, X_{t-k})$ is a well known concept for describing the serial dependence in the time seris, but the ACF is, again, only capable to completely capture \emph{linear} serial dependence.
+The autocorrelation function (ACF) $\rho_k = \rho(X_t, X_{t-k})$ is a well known concept for describing the serial dependence in the time series, but the ACF is, again, only capable to completely capture \emph{linear} serial dependence.
 \citet{laca:tjos:2017} seek to remedy this by rather calculating the the local correlation between $X_t$ and $X_{t-k}$.
@@
 
-Although the original work by \citet{hjor:jone:1996} provide a general framework for local likelihood density estimation, using any $p$-variate parametric family as the local family, it is evident that the method may struggle in multivariate applications much in the same way as the kernel density estimator does.
+Although the original work by \citet{hjor:jone:1996} provide a general framework for local likelihood density estimation using any $p$-variate parametric family as the local family, it is evident that the method may struggle in multivariate applications much in the same way as the kernel density estimator does.
 This is a consequence of the curse of dimensionality, the effect of which is sought remedied by an algorithm provided by \citet{otne:tjos:2017}.
@@
 They argue further that the following transformation technique will produce better density estimates in many situations.
+
 Denote by $F_i(x_i)$, $i = 1,\ldots, p$ the marginal distribution functions of the stochastic vector $X$, and by $\widehat F_i(x_i) = n^{-1}\sum_{i=1}^n 1(X_i \leq x_i)$ their empirical counterparts.
@@
 \begin{equation}
-\widehat f_Z(z) = \psi(z; \mu = 0, \sigma = 1, R = \widehat R(z)),
+\widehat f_Z(z) = \psi(z; \mu(z) = 0, \sigma(z) = 1, R = \widehat R(z)),
 \label{eq:transformed-density}
 \end{equation}
-and then substitute $f_Z$ for $\widehat f_Z$ in the following relation obtained by \citet{otne:tjos:2017} in order to estimate the unknown density $f_X$:
+where $\mu(z) = \{\mu_i(z)\}$ and $\sigma(z) = \{\sigma_i(z)\}$ for $i=1,\ldots,p$, and then substitute $f_Z$ for $\widehat f_Z$ in the following relation obtained by \citet{otne:tjos:2017} in order to estimate the unknown density $f_X$:
 \begin{equation}
@@
 where we, again, switch to $z$-notation in order to make it clear that these quantities are estimated on the standard normal $z$-scale.
-An estimator $\widehat f_{X^{(1)}|X^{(2)}}(\cdot|\cdot)$ of the conditional density $f_{X^{(1)}|X^{(2)}}(\cdot|\cdot)$ follows immediately from an expression corresponding to (\ref{eq:backtrans}), and the \CRANpkg{lg}-package provides functions for implementing this estimator in R, and we describe the implementation of this functionality in Section \ref{chap:density}.
+An estimator $\widehat f_{X^{(1)}|X^{(2)}}(\cdot|\cdot)$ of the conditional density $f_{X^{(1)}|X^{(2)}}(\cdot|\cdot)$ follows immediately from an expression corresponding to (\ref{eq:backtrans}), and the \CRANpkg{lg}-package provides functions for implementing this estimator in R.
+We describe the implementation of this functionality in Section \ref{chap:density}.
 
-Finally, we refer to \citet{otne:tjos:2019}, who take the local version of the conditional covariance matrix (\ref{eq:cond-sig}) (or (\ref{eq:cond-sig-trans}) in the transformed case) as a measure for conditional dependence, and thus as an instrument to test for conditional independence.
+Finally, we refer to \citet{otne:tjos:2019} who take the local version of the conditional covariance matrix (\ref{eq:cond-sig}) (or (\ref{eq:cond-sig-trans}) in the transformed case) as a measure for conditional dependence, and thus as an instrument to test for conditional independence.
 Consider the stochastic vector $X = \left(X^{(1)}, X^{(2)}, X^{(3)}\right)$, where $X^{(1)}$ and $X^{(2)}$ are scalars and $X^{(3)}$ may be a vector.
@@
 The local Gaussian correlation may be used to perform a number of statistical analyses, as is evident from the preceding section.
-The practitioner must first, however, make three quite specific modelling choices; namely (i) to choose an estimation method, i.e.
+The practitioner must first, however, make three quite specific modeling choices; namely (i) to choose an estimation method, i.e.
 the level of simplification in multivariate applications, (ii) to determine whether the data should be transformed towards marginal standard normality before estimating the LGC, and (iii) to choose a set of bandwidths, or at least a method for calculating bandwidths.
@@
 
-In the following we assume that one has a \emph{data set} \code{x} loaded into the R workspace, which must be an $n\times p$ matrix (one column per variable, one row per observation), possibly including \code{NA}`s which will be excluded, or a data frame having the same dimensions.
-The fundamental syntax for creating an lg-object is \code{lg\_object <- lg\_main(x)}, and we will in this section explain how the modelling decisions (i)-(iii) can be encoded into the lg-object.
+In the following we assume that one has a \emph{data set} \code{x} loaded into the R workspace, which must be an $n\times p$ matrix (one column per variable, one row per observation), possibly including \code{NA}s which will be excluded from the analysis, or a data frame having the same dimensions.
+The fundamental syntax for creating an lg-object is \code{lg\_object <- lg\_main(x)}, and we will in this section explain how the modeling decisions (i)-(iii) can be encoded into the lg-object by using appropriate arguments in this function.
 
@@
 Note that the full locally Gaussian fit for raw data is not available if the number of variables $p$ is greater than 2.
-The \code{lg\_main()}-function will check for this and throw an error if $p>2$ and \code{est\_method = "5par"}.
+The \code{lg\_main()}-function will check for this and print out an error message if $p>2$ and \code{est\_method = "5par"}.
 
@@
 the transformation (\ref{eq:trans}).
-In that case, we fix the marginal expectation functions and standard deviation functions to the constant values $0$ and $1$ respectively, rather then estimating them.
+In that case, we fix the marginal expectation functions and standard deviation functions to the constant values $0$ and $1$ respectively, and estimate only the pairwise local Gaussian correlations as in (\ref{eq:transformed-density}).
 To use this estimation method, create the lg-object by running
@@
 
-Note that the function call above will issue a warning if the option for transforming the data to marginal standard normality is not at the same time set to \code{TRUE}, see the next sub-section for details.
+Note that the function call above will issue a warning if the option for transforming the data to marginal standard normality is not at the same time set to \code{TRUE}, see the next sub-section on data transformation for details.
 
 \textbf{4.
-A full locally Gaussian fit for trivariate data.} If $p = 3$ \emph{and} we choose to transform the data to marginal standard normality (see the next sub-section), the transformed density $f_Z(\cdot)$ in (\ref{eq:transformed-density}) may be estimated by jointly estimating the three local correlations $\rho_{12}(z_1, z_2, z_3)$, $\rho_{13}(z_1, z_2, z_3)$ and $\rho_{23}(z_1, z_2, z_3)$.
+A full locally Gaussian fit for trivariate data.} If the number of variables $p$ is equal to $3$ \emph{and} we choose to transform the data to marginal standard normality (see the next sub-section), the transformed density $f_Z(\cdot)$ in (\ref{eq:transformed-density}) may be estimated by jointly estimating the three local correlations $\rho_{12}(z_1, z_2, z_3)$, $\rho_{13}(z_1, z_2, z_3)$ and $\rho_{23}(z_1, z_2, z_3)$.
 This estimation method was introduced recently by \citet{otne:tjos:2019} in order to increase power of their conditional independence test, but it can be used in any application described in this paper that consider trivariate data.
@@
 We have in practice seen two automatic bandwidth selectors employed in the applications referred to in Section \ref{chap:background}: a
-cross-validation procedure, which is slow, but accurate with respect to density estimation, and a plug-in bandwidth that is quick, but cruder. We use the argument
+cross-validation procedure, that is fairly slow to compute, but accurate with respect to density estimation, and a plug-in bandwidth that is much quicker to calculate, but less accurate with respect to density estimation.
+We use the argument
 \code{bw\_method} to the \code{lg\_main()}-function in order to choose between the two.
@@
 \(p(p-1)/2\) bandwidth matrices, one for each pair of variables (which in this bivariate example just \emph{one} variable pair, (\code{x1,\ x2})).
-The convergence flags stem from the \code{optim()} and \code{optimize()} functions that we use to obtain the minimizer of \(CV(\cdot)\), and 0 indicates successful convergence.
+The convergence flags stem from the built-in R functions \code{optim()} and \code{optimize()} that we use to obtain the minimizer of \(CV(\cdot)\), and 0 indicates successful convergence.
 
@@
 In the sub-section above we present the three most important arguments to \code{lg\_main()}.
-Each of them allows the user to configure one of the three crucial modelling choices.
+Each of them allows the user to configure one of the three crucial modeling choices.
 Let us complete this treatment by covering some possibilities to make further adjustments to those choices.
@@
   The user may supply the bandwidths directly to \code{lg\_main()} by passing them to the \code{bw}-argument.
-They have to be in the   correct format though, which is a list containing the vector \code{\$marginal} if \code{est\_method\ =\ 5par\_marginals\_fixed}, and always a data frame \code{\$joint} specifying all variable pairs in the \code{x1} and \code{x2} columns and the corresponding bandwidths
+They have to be in the   correct format though, which is a list containing the vector \code{\$marginal} if \code{est\_method\ =\ "5par\_marginals\_fixed"}, and always a data frame \code{\$joint} specifying all variable pairs in the \code{x1} and \code{x2} columns and the corresponding bandwidths
   in the \code{bw1} and \code{bw2} columns. The function \code{bw\_simple()} will assist in creating bandwidth objects.
@@
 We proceed in this section to demonstrate how to implement each of the tasks that we discussed in Section \ref{chap:background}.
-The general pattern is to pass an lg-object, see the preceding section, to one of the estimation or test functions provided in the \CRANpkg{lg}-package.
+The general pattern is to pass the lg-object to one of the estimation or test functions provided in the \CRANpkg{lg}-package.
 We will look at some financial data in the examples: the monthly returns on the S\&P500, FTSE100, DAX30 and TOPIX stock indices from January 1985 to March 2018.
-This data is available as an online appendix to this article.
+The data is available as an online appendix to this article.
 
 \subsection{Density estimation}\label{chap:density}
+
+We start by introducing a basic function for estimating the LGC on a grid as described by \citet{otne:tjos:2017}, and thus also a probability density estimate.
+We create a grid, \code{x0}, having the same number of columns as the data in the code below.
+Note that we use the pipe operator \code{\%\textgreater{}\%} from the \CRANpkg{magrittr}-package \citep{bach:wick:2014} as well as functions from the \CRANpkg{dplyr}-package \citep{wick:fran:mull:2018} for easy manipulation of data frames.
+We then pass the grid and the lg-object containing our modeling choices to the \code{dlg()}-function in order to do the estimation.
 
@@
   \end{figure}
-
-We start by introducing a basic function for estimating the LGC on a grid as described by \citet{otne:tjos:2017}, and thus also a probability density estimate.
-We create a grid, \code{x0}, having the same number of columns as the data in the code below.
-Note that we use the pipe operator \code{\%\textgreater{}\%} from the \CRANpkg{magrittr}-package \citep{bach:wick:2014} as well as functions from the \CRANpkg{dplyr}-package \citep{wick:fran:mull:2018} for easy manipulation of data frames.
-We then pass the grid and the lg-object containing our modelling choices to the \code{dlg()}-function in order to do the estimation.
 
@@
 This also means that the number of columns in the grid \code{x0} plus the number of elements in \code{cond} must equal the number of variables \(p\) in the data set, and the call to \code{clg()} will result in an error message if this requirement is not fulfilled.
-Apart from \code{cond}, the \code{clg()}-function takes exactly the same arguments as \code{dlg()} listed in Table \ref{tab:dlg}, and the conditional density estimate in our example is available in the vector \code{cond\_density\$f\_est}.
+The \code{clg()}-function takes mostly the same arguments as \code{dlg()} listed in Table \ref{tab:dlg}, and the conditional density estimate in our example is available in the vector \code{cond\_density\$f\_est}.
 
@@
 
-Consider finally the work by \citet{otne:tjos:2019} who take the off-diagonal element in the local correlation matrix corresponding to the local conditional covariance matrix \eqref{eq:cond-sig} or \eqref{eq:cond-sig-trans} as a local measure of conditional dependence between two stochastic variables \(X^{(1)}\) and \(X^{(2)}\) given the stochastic vector \(X^{(3)}\), and furthermore, in the case with data having been transformed to marginal standard normality, take the statistic
-\[T^{CI}_{n, b} = \int h(\widehat\alpha(z)) \, \textrm{d}F_n(z)\]
+Consider finally the work by \citet{otne:tjos:2019} who take the off-diagonal element in the local correlation matrix corresponding to the local conditional covariance matrix \eqref{eq:cond-sig} or \eqref{eq:cond-sig-trans} as a local measure of conditional dependence between two stochastic variables \(X^{(1)}\) and \(X^{(2)}\) given the stochastic vector \(X^{(3)}\).
+Furthermore, in the case with data having been transformed to marginal standard normality, they take the statistic
+\begin{equation}
+T^{CI}_{n, b} = \int h(\widehat\alpha(z)) \, \textrm{d}F_n(z)
+\label{eq:condintstat}
+\end{equation}
 as a measure of \emph{global} conditional dependence.
-To this end, the \CRANpkg{lg}-package provides two functions.
-The first, \code{partial\_cor()}, calculates the local partial correlations as well as their estimated standard deviations, on a specified grid in the \((x^{(1)}, x^{(2)}, x^{(3)})\)-space, and is essentially a wrapper function for the \code{clg()}-function presented in Section \ref{chap:density}.
+The \CRANpkg{lg}-package provides two key functions in this framework.
+The first, \code{partial\_cor()}, calculates the local partial correlations as well as their estimated standard deviations on a specified grid in the \((x^{(1)}, x^{(2)}, x^{(3)})\)-space, and is essentially a wrapper function for the \code{clg()}-function presented in Section \ref{chap:density}.
 See Table \ref{tab:arguments-partial-cor} for details.
-The second function, \code{ci\_test()}, performs a test for conditional independence between the first two variables in a data set given the remaining variables.
+The second function, \code{ci\_test()}, performs a test for conditional independence between the first two variables in a data set given the remaining variables using the test statistic (\ref{eq:condintstat}) and a special bootstrap procedure (described briefly below) for approximating the null distribution.
 
@@
 
-We refer to the documentation of the \CRANpkg{lg}-package for details on how to finely tune the behavior of the bootstrapping algorithm by altering the arguments of the \code{ci\_test()}-function, and rather describe the arguments most suitable for modifications by the user in Table \ref{tab:arguments-ci-test}.
+We refer to the documentation of the \CRANpkg{lg}-package for details on how to finely tune the behavior of the bootstrapping algorithm by altering the arguments of the \code{ci\_test()}-function, and limit our treatment to describing the arguments most suitable for modifications by the user in Table \ref{tab:arguments-ci-test}.
 
@@
 
-We conclude this article by describing the \code{corplot()} function, that may be used to draw dependence maps such as those displayed in Figure \ref{fig:initial-example}.
+We conclude this article by describing the \code{corplot()} function for drawing dependence maps such as those displayed in Figure \ref{fig:initial-example}.
 \citet{bere:klep:tjos:2014} report on such capabilities in the \CRANpkg{localgauss}-package, but the possibility of creating dependence maps was unfortunately removed from \CRANpkg{localgauss} in the latest version 0.4.0 due to incompatibilities with the \CRANpkg{ggplot2} \citep{wick:2016} plotting engine.
@@
 The plotting function is highly customizable and provides a number of options covering most basic graphical options.
-Users well versed in the \CRANpkg{ggplot2}-package may also modify the graphical object returned by corplot in the standard way by adding layers as demonstrate in the example below.
+Users well versed in the \CRANpkg{ggplot2}-package may also modify the graphical object returned by \code{corplot()} in the standard way by adding layers as demonstrated in the example below.
 
@@
 The package is complete, in the sense that all major methods that have been published within this framework is now easily accessible to the practitioner.
-The package is also modular; Future methodological developments using local Gaussian approximations may be added to the package with little effort.
+The package is also designed with a modular infrastructure that allows future methodological developments using local Gaussian approximations to be easily added to the package.
+
+\section{Acknowledgements}
+
+The author gratefully acknowledges the constructive comments made by an anonymous referee.
 

### 2021-02-17 submit/otneim.tex
Resubmitted to The R Journal

--- 2021-02-16 otneim.tex
@@
       \includegraphics[width=\textwidth]{gaussian-example}
-      \caption{Estimated local Gaussian correlations between two  jointly normal variables having correlation equal to 0.5, based on 1000 observations.}
+      \caption{Estimated local Gaussian correlations between \\ two jointly normal variables having correlation \\ equal to 0.5, based on 1000 observations.}
     \end{subfigure}
@@
 Consider next the stationary time series $\{X_t\}$.
-The autocorrelation function (ACF) $\rho_k = \rho(X_t, X_{t-k})$ is a well known concept for describing the serial dependence in the time series, but the ACF is, again, only capable to completely capture \emph{linear} serial dependence.
+The autocorrelation function (ACF) $\rho_k = \rho(X_t, X_{t-k})$ is a well known concept for describing the serial dependence in the time series, but the ACF is, again, only capable of completely capturing \emph{linear} serial dependence.
 \citet{laca:tjos:2017} seek to remedy this by rather calculating the the local correlation between $X_t$ and $X_{t-k}$.
@@
 We find another application of the local Gaussian approximation in the work by \citet{stov:tjos:huft:2014}, who measure and test for financial contagion.
-They define contagion by "a significant increase in cross-market linkages after a shock to one country" \citep{forb:rigo:2002} and employ the LGC to quantify this potential linkage.
+They define contagion as "a significant increase in cross-market linkages after a shock to one country" \citep{forb:rigo:2002} and employ the LGC to quantify this potential linkage.
 The authors estimate the LGC on a grid $\{x_1, x_2\}_k$, $k=1,\ldots,K$ along the diagonal $D=\{(x_1,x_2): x_1 = x_2\}$ \emph{before} and \emph{after} some critical event in the financial markets, denoted as the crisis (C) and the non-crisis (NC) periods respectively.
@@
 \end{equation}
-where $\phi(\cdot)$ is the standard normal pdf. This estimator is implemented the \CRANpkg{lg}-package as demonstrated in Section \ref{chap:density}.
+where $\phi(\cdot)$ is the standard normal pdf. This estimator is implemented in the \CRANpkg{lg}-package as demonstrated in Section \ref{chap:density}.
 
@@
 
-Consider finally the work by \citet{otne:tjos:2019} who take the off-diagonal element in the local correlation matrix corresponding to the local conditional covariance matrix \eqref{eq:cond-sig} or \eqref{eq:cond-sig-trans} as a local measure of conditional dependence between two stochastic variables \(X^{(1)}\) and \(X^{(2)}\) given the stochastic vector \(X^{(3)}\).
+Consider, finally, the work by \citet{otne:tjos:2019} who take the off-diagonal element in the local correlation matrix corresponding to the local conditional covariance matrix \eqref{eq:cond-sig} or \eqref{eq:cond-sig-trans} as a local measure of conditional dependence between two stochastic variables \(X^{(1)}\) and \(X^{(2)}\) given the stochastic vector \(X^{(3)}\).
 Furthermore, in the case with data having been transformed to marginal standard normality, they take the statistic
@@
 The statistical literature has seen a number of applications of local Gaussian approximations in the last decade, covering several topics in dependence modeling and inference, as well as the estimation of multivariate density and conditional density functions.
-In this paper, we demonstrate the implementation of these methods i the R programming language using the \CRANpkg{lg}-package, as well as the graphical representation of the estimated local Gaussian correlation.
+In this paper, we demonstrate the implementation of these methods in the R programming language using the \CRANpkg{lg}-package, as well as the graphical representation of the estimated local Gaussian correlation.
 The package is complete, in the sense that all major methods that have been published within this framework is now easily accessible to the practitioner.

### 2021-02-17 submit/otneim.R

@@ -1,0 +1,368 @@
+
+# Load packages ------
+library(lg)
+library(ggplot2)
+library(lubridate)
+library(tidyr)
+library(fGarch)
+library(dplyr)
+library(extrafont)
+loadfonts(quiet = TRUE)
+
+# Initial example 1 --------
+
+# Estimated local Gaussian correlation between two jointly normal variables 
+# having correlation equal to 0.5, using 1000 observations.
+
+set.seed(1)
+n   <- 1000
+rho <- .5
+
+# Simulated Gaussian data
+grid1 <- expand.grid(seq(-3, 3, length.out = 7),
+                     seq(-3, 3, length.out = 7))
+
+x1 <- mvtnorm::rmvnorm(500, sigma = matrix(c(1, rho, rho, 1), 2))
+lg_object1 <- lg_main(x1, 
+                      est_method = "5par", 
+                      transform_to_marginal_normality = FALSE,
+                      plugin_constant_joint = 4)
+dlg_object1 <- dlg(lg_object1, grid = grid1)
+
+pdf(file = "gaussian-example.pdf", height = 4, width = 5, family = "CM Roman")
+corplot(dlg_object1, plot_thres = .01,
+        xlab = "X1", ylab = "X2", 
+        main = "",
+        plot_obs = TRUE,
+        alpha_point = .1, 
+        plot_legend = FALSE,
+        label_size = 4) +
+    theme_classic() +
+    theme(legend.position = "none")
+dev.off()
+embed_fonts("gaussian-example.pdf")
+
+# Initial example 2 --------
+
+# Plotting the local dependence maps for the returns data
+x2 <- read.csv2("G5_daily_r.csv") %>% 
+    as_tibble %>% 
+    tail(n) %>% 
+    select(CAC40, FTSE100) 
+
+grid2 <- expand.grid(seq(-3.5, 3.5, length.out = 7),
+                     seq(-3, 3, length.out = 7))
+
+lg_object2 <- lg_main(x2, 
+                      est_method = "5par", 
+                      transform_to_marginal_normality = FALSE,
+                      plugin_constant_joint = 2)
+dlg_object2 <- dlg(lg_object2, grid = grid2)
+
+pdf(file = "returns-example.pdf", height = 4, width = 5, family = "CM Roman")
+corplot(dlg_object2, plot_thres = .0001,
+        xlab = "CAC40", ylab = "FTSE100", 
+        main = "",
+        plot_obs = TRUE,
+        alpha_point = .1, 
[…]

### 2021-08-04 otneim.tex
Second revision submitted to the R Journal

@@
 
-In Figure \ref{fig:initial-example} we have plotted the estimated LGC for two bivariate data sets on a grid; 1000 simulated observations from a binormal distribution having correlation equal to 0.5, and the daily return on the CAC40 and FTSE100 stock indices on 1000 consecutive trading days starting on May 5th 2014.
+In Figure \ref{fig:initial-example} we have plotted the estimated LGC for two bivariate data sets on a grid; 1000 simulated observations from a binormal distribution having correlation equal to 0.5, and the daily return on the CAC40 and FTSE100 stock indices on 1000 consecutive trading days starting on May 5th 2014 \citep{datastream}.
 In the first panel, we see that the estimated local correlation coincides with the global correlation, except for the estimation error.
@@
 \citet{bere:tjos:2014} show that $\rho(x) \equiv 0$ implies that $X_1$ and $X_2$ are independent.
-The converse statement holds if $\mu_i(x_1, x_2) = \mu_i(x_i)$ and $\sigma_i(x_1, x_2) = \sigma_i(x_i)$ for $i=1,2$, and in general if the observations have been suitably transformed according to a procedure presented later in this section.
+They show further that independence between $X_1$ and $X_2$ implies $\rho(x) \equiv 0$ if the population values of the local mean and standard deviation functions satisfy the following conditions: $\mu_i(x_1, x_2) = \mu_i(x_i)$ and $\sigma_i(x_1, x_2) = \sigma_i(x_i)$ for $i=1,2$.
+Equivalence between independence and $\rho(x) \equiv 0$ holds in general if the observations have been suitably transformed according to a procedure presented later in this section.
 It follows then that departures from $\widehat\rho(x) \equiv 0$ may be taken as evidence against the hypothesis that $X_1$ and $X_2$ are statistically independent.
@@
 We find another application of the local Gaussian approximation in the work by \citet{stov:tjos:huft:2014}, who measure and test for financial contagion.
-They define contagion by "a significant increase in cross-market linkages after a shock to one country" \citep{forb:rigo:2002} and employ the LGC to quantify this potential linkage.
+They define contagion by "a significant increase in cross-market linkages after a shock to one country" \citep[p.
+2223]{forb:rigo:2002} and employ the LGC to quantify this potential linkage.
 The authors estimate the LGC on a grid $\{x_1, x_2\}_k$, $k=1,\ldots,K$ along the diagonal $D=\{(x_1,x_2): x_1 = x_2\}$ \emph{before} and \emph{after} some critical event in the financial markets, denoted as the crisis (C) and the non-crisis (NC) periods respectively.
@@
 \begin{align}
-\mu_i(x) &= \mu_i(x_1, \ldots, x_p) \stackrel{\textrm{def}}{=} \mu_(x_i) \label{eq:simp-mu}\\
+\mu_i(x) &= \mu_i(x_1, \ldots, x_p) \stackrel{\textrm{def}}{=} \mu_i(x_i) \label{eq:simp-mu}\\
 \sigma_i(x) &= \sigma_i(x_1, \ldots, x_p)  \stackrel{\textrm{def}}{=} \sigma_i(x_i) \label{eq:simp-sig} \\
@@
 They argue further that the following transformation technique will produce better density estimates in many situations.
+The motivation for introducing the simplifications defined in equations \ref{eq:simp-mu}-\ref{eq:simp-rho} can be compared to the practical advantages of estimating additive regression models, where $\textrm{E}(Y) = f(x_1, \ldots, x_p) \stackrel{\textrm{def}}{=} f_1(x_1) + \cdots + f_p(x_p)$.
 
@@
 $$f_X(x) = \psi(x, \mu(x), \Sigma(x)),$$
-
 then the conditional density of $X^{(1)}|X^{(2)} = x^{(2)}$ can also be written on the same locally Gaussian form with local parameters given by equations (\ref{eq:cond-mu}) and (\ref{eq:cond-sig}), except that all quantities are $x$-dependent.
@@
 
-\section{The first step: Creating the \texttt{lg}-object} 
+\section{The first step: Creating the lg-object} 
 \label{chap:lgobject}
@@
 The general pattern is to pass the lg-object to one of the estimation or test functions provided in the \CRANpkg{lg}-package.
-We will look at some financial data in the examples: the monthly returns on the S\&P500, FTSE100, DAX30 and TOPIX stock indices from January 1985 to March 2018.
-The data is available as an online appendix to this article.
+We will look at some financial data in the examples: the monthly returns on the S\&P500, FTSE100, DAX30 and TOPIX stock indices from January 1985 to March 2018 \citep{datastream}.
 
@@
       \includegraphics[width=\textwidth]{loccor-example}
-      \caption{The estimated pairwise LGC for the stock \newline data along the diagonal.}
+      \caption{The estimated pairwise LGC for the stock \newline data along the diagonal indicating that the pairs \\ of stocks are more strongly dependent in the lower \\ part of the distribution than the upper part.}
       \label{fig:lgc-plot1}
@@
 \end{example}
-This may take a few minutes due to the bootstrapping. We can further specify the function \(h\) and the integration area \(S\) in the test
+This may take a few minutes to run on a desktop computer due to the bootstrapping.
+The small $p$-value indicate that we reject the null-hypothesis of independence between $X_1$ and $X_2$ in the parabola model defined above.
+We can further specify the function \(h\) and the integration area \(S\) in the test
 statistic \eqref{eq:test-statistic1}, see Table \ref{tab:indtest} for details.
@@
 \end{example}
+
+The small $p$-value means that we reject the null-hypothesis of no financial contagion between the time series after the crisis.
 
@@
 R> # Make a dependence map using default setup
-R> corplot(dlg_object1)
+R> corplot(dlg_object)
 \end{example}

### 2021-08-04 response2/response2.Rmd
Second revision submitted to the R Journal

---
title: ""
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\begin{center}
\Large \textbf{Response to reviewer comments to the manuscript \emph{2019-146: lg: An R package for Local Gaussian Approximations} submitted to \emph{The R Journal} -- Round 2}
\end{center}

I am grateful for the opportunity to revise the paper for publication in *The R Journal*. Below are my point-by-point response to the comments made by the reviewer.

> *1. What does “$\mu_i(x_1, x_2) = \mu_i(x_i)$ and $\sigma_i(x_1, x_2) = \sigma_i(x_i)$ for $i = 1, 2$” imply? According to your text – if I understand correctly – it implies dependence. However, I feel, the opposite is true. Have a look please. Compare also to Berentsen & Tjøstheim (2014), p. 788.*

**My response:** I presume that the reviewer is referring to the paragraph leading up to equation (3). The statement refers to a specific condition on the true mean and variance functions of a bivariate population that leads to equivalence between independence and constant zero local correlation. The condition $\mu_i(x_1, x_2) = \mu_i(x_i)$ for $i=1,2$ means that $\mu_1$ is a function of only $x_1$, and that $\mu_2$ is a function of only $x_2$, and the same for the $\sigma$-functions. I have written out this paragraph with more details in order to make this point clear.

> *2. Please confirm that Eq. (3) is complete. Compare Berentsen & Tjøstheim (2014) pp. 790-791.*

**My response:** The equation is correct, and corresponds exactly to the definition of the test statistic defined in the beginning of Section 3.2 of Berentsen and Tjøstheim (2014). The notation in my paper ($\widehat \rho = \rho_{n,b}$) is more consistent with the recent publications in this field.

> *3. There is a typesetting error in Eq. (4).*

**My response:** Has been fixed, thank you.

> *4. I assume there is a (little) typo in the code chunk on page 14: You assign a value/information to an object called ‘dlg_object’ but then use ‘dlg_object1’, also in the following code chunks.*

**My response:** Has been fixed, thank you.

> *5. Please consider adding a page number to the quote from Forbes & Rigobon (2002) on p. 3.*

**My response:** Has been added, thank you.

> *6. Please add interpretations of the empirical findings in the empirical application(s).*

**My response:** Thank you for this comment. The following empirical examples did not have practical interpretations, but they have now been added:

- the caption of Figure 3,
- the independence test on the parabola model on page 11, and
- the test for financial contagion on page 13.

### 2021-08-16 otneim.tex
Updated according to a third review

@@
 In Figure \ref{fig:initial-example} we have plotted the estimated LGC for two bivariate data sets on a grid; 1000 simulated observations from a binormal distribution having correlation equal to 0.5, and the daily return on the CAC40 and FTSE100 stock indices on 1000 consecutive trading days starting on May 5th 2014 \citep{datastream}.
-In the first panel, we see that the estimated local correlation coincides with the global correlation, except for the estimation error.
+In the first panel, we see that the estimated local correlation coincides with the global correlation, except for the estimation error which is comparable to the uncertainty observed in other non-parametric estimation methods such as the kernel density estimator (see, for instance, \citet{otne:tjos:2017} for a formal asymptotic analysis of relevant convergence rates).
 In the second panel we see clearly that the local correlation, and thus the dependence, is stronger in the lower left and upper right regions of the distribution than in the central parts.
@@
 Given a data set \code{x} having $n$ rows and $p \geq 2$ columns, the user must choose between four distinct estimation methods, and specify this choice by using the argument \code{est\_method} to the \code{lg\_main()}-function.
-We look at the built-in bivariate data set `faithful` and load the \CRANpkg{lg}-package in order to demonstrate the implementation:
+We look at the built-in bivariate data set `faithful`, which records the waiting time between eruptions and the duration of the eruption for the Old Faithful geyser in the Yellowstone National Park, USA (see the help file in R for more details: \code{?faithful}), and load the \CRANpkg{lg}-package in order to demonstrate the implementation:
 
@@
 \end{example}
+This command will throw an error if the data set \code{x} does not have exactly three columns.
 
@@
 Finally, the only thing that we must alter in order to perform the third test for serial cross-dependence is the bootstrap method.
-\citet{laca:tjos:2018} consider two options here, the block bootstrap with either fixed \citep{kuns:1989} or random \citep{poli:roma:1994} block sizes, and this choice is specified by choosing \code{bootstrap\_type\ =\ "block"} or \code{bootstrap\_type\ =\ "stationary"}, respectively, in the call to \code{ind\_test()}.
+In the applications above, it suffices to use the standard bootstrap, where we resammple with replacement from the data.
+This is implemented in the \code{ind\_test()}-function by setting the \code{bootstrap\_type}-argument to \code{"plain"}, which is the default option.
+When testing for serial cross-dependence we need to use a block-bootstrap procedure, and \citet{laca:tjos:2018} consider two options here: The block bootstrap with either fixed \citep{kuns:1989} or random \citep{poli:roma:1994} block sizes, and this choice is specified by choosing \code{bootstrap\_type\ =\ "block"} or \code{bootstrap\_type\ =\ "stationary"}, respectively, in the call to \code{ind\_test()}.
+\citet{laca:tjos:2018} do not report significant differences in test performance using the different bootstrap types.
 
@@
 R> # demonstration purposes.
-R> result <- cont_test(lg_object_nc, lg_object_c, n_rep = 20)
+R> result <- cont_test(lg_object_nc, lg_object_c, n_rep = 100)
 R> 
@@
 
-[1] 0
+[1] 0.01
 \end{example}
@@
 
-[1] 0.71
+[1] 0.51
 \end{example}

### 2021-08-16 response3/response3.Rmd
Updated according to a third review

---
title: ""
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\begin{center}
\Large \textbf{Response to reviewer comments to the manuscript \emph{2019-146: lg: An R package for Local Gaussian Approximations} submitted to \emph{The R Journal} -- Round 3}
\end{center}

I am grateful for the opportunity to revise the paper for publication in *The R Journal*. Below are my point-by-point response to the comments made by the latest reviewer report.

> *Overview: The article describes how to use the lg package for local gaussian approximations. The proposed methodology provides insights into the correlation structure of data beyond the usual central correlation of Spearman. I think that this methodology is interesting, in particular in the context of financial analysis where tail dependence structures are of importance. I therefore think that the package will be used and the article will be consulted and cited.*

**My response:** Thank you for your kind remarks. The methodology has indeed been applied to financial analysis of various types.

> *I find the article clear and well written in general.*

**My response:** Thank you.

> *I would appreciate a few words on the "faithful" dataset which is used in the examples. I didn't know what it contains and had to find that out first somewhere else.*

**My response:** I have added a brief description of the data, as well as a reference to the description of the data in the R documentation.

> *I would suggest to include a few more words to explain the differences between the bootstrap procedures. In particular, I'm not sure about the difference between the "plain" and the "stationary" implementation.*

**My response:** A good suggestion. I have expanded the relevant paragraph so that the `plain`-option is explicitly mentioned and described as standard data resampling with replacement.

> *Some information about the minimum number of observations needed to reliably estimate the tails and boundary regions should be added. I modified the introductory example to x1 <- matrix(rnorm(1000), 500,2) which indicated significant dependence in some parts of the multivariate distribution, in particular at the thin ends. Of course, my data here are independent and this is a small sample issue which should be addressed in the paper as the method suggests dependence which is not there.*

**My response:** Thank you for raising this important point. I did the same thing, executed the following code, and obtained the figure below:

```{r, message = FALSE, warning = FALSE}
library(lg)
library(ggplot2)
library(dplyr)

set.seed(1)

# Simulated Gaussian data
grid1 <- expand.grid(seq(-3, 3, length.out = 7),
                     seq(-3, 3, length.out = 7))

x1 <- matrix(rnorm(1000), 500,2)
lg_object1 <- lg_main(x1, 
                      est_method = "5par", 
                      transform_to_marginal_normality = FALSE,
                      plugin_constant_joint = 4)
dlg_object1 <- dlg(lg_object1, grid = grid1)

corplot(dlg_object1, plot_thres = .01,
        xlab = "X1", ylab = "X2", 
        main = "",
        plot_obs = TRUE,
        alpha_point = .1, 
        plot_legend = FALSE,
        label_size = 4) +
    theme_classic() +
    theme(legend.position = "none")
```

If we have used the same random seed this figure should be identical to what the reviewer has obtained. Whether the observed local correlations represent *significant* (in the statistical sense) dependence can be investigated by looking at the estimated standard deviations, and lower/upper confidence limits which is reporten on the 95% confidence level by default, available in the `dlg_object1` in the `$loc_cor_sd`, `$loc_cor_lower` and `$loc_cor_upper` slots respectively. By plotting the estimated correlations with their confidence intervals, as I have done below, we see that zero (represented as a horizontal line) is included in almost all of the intervals.

```{r}
data.frame(index = 1:length(dlg_object1$f_est),
           rho = dlg_object1$loc_cor, 
           lower = dlg_object1$loc_cor_lower, 
           upper = dlg_object1$loc_cor_upper) %>% 
  ggplot(aes(x = index)) +
  geom_point(aes(y = rho)) +
  geom_errorbar(aes(ymin = lower, ymax = upper)) +
  geom_hline(yintercept = 0)
```

The global Pearson correlation coefficient for this data is $-0.041$. The corresponding independence test does not reject independence for this data. I get a $p$-value of 0.63 using 100 resamples.

The conclusion is that the dependence map in this case in fact does not indicate dependence that is statistically significant.

The comment made by the reviewer is relevant though, and represents a question that a reader of this paper might be interested in. Several of the main references that analyze the local Gaussian correlation provide asymptotic analyses of convergence rates under various common sets of assumptions, including the iid case as well as time series dependence. It turns out that the local Gaussian correlation estimates converge with the same speed, and hence require approximately the same sample sizes, as other well known non-parametric estimation methods that require a smoothing bandwidth, such as the kernel density estimator. The uncertainty is naturally greater in the tails of the distribution as the reviewer indicates.

I have added a comment that hopefully clears up this point in the discussion of this example. The latter part of the following sentence is new:

*"In the first panel, we see that the estimated local correlation coincides with the global correlation, except for the estimation error which is comparable to the uncertainty observed in other non-parametric estimation methods such as the kernel density estimator (see, for instance, Otneim and Tjøstheim (2017) for a formal asymptotic analysis of relevant convergence rates). "*

> *I tested the package on an openSuse 10 tumbleweed system (state 20210724) on a thinkpad i7-8650U with 16 GB RAM with R version 4.1.0.*

**My response:** I appreciate that the reviewer has taken the time to test the code file.

> *line 90: `lg_object4 <- lg_main(x, est_method = "trivariate")` cannot be created as x contains 2 columns only*

**My response:** Yes this is correct. At this point in the script, the data `x` is the bivariate `faithful `data set, for which we cannot fit a trivariate density estimate. That is the reason why the reviewer gets an error. I have commented out this particular line in the script and added an explanation. I have also pointed out in the paper that the function call will result in an error if the data has the wrong dimension.

> *line 132: `stock_data_d` is not used, could be dropped*

**My response:** Correct, has been taken out.

> *line 239: the independence test (ind_test()) takes quite some time. Is there any possibility to speed that up by using parallel computing? It took 6.805 minutes on my computer, even longer with the return data. I appreciate the warning that it takes time given in the paper.*

**My response:** Indeed, calculating critical values using bootstrapping takes quite a while, even for relatively modest sample sizes. This is noted in the paper, as the reviewer points out. It is possible to speed up the calculations using parallel computing, and this was actually implemented in an earlier version of the package. Unfortunately, this turned out to be a very unstable feature, because the back-ends that have been developed in various packages were too system dependent to work consistently on all platforms. It was therefore decided to take parallelization out of the package in order to keep it robust and future-proof.

It is, however, perfectly possible to wrap the functions in this package in the appropriate parallelized loops in order to speed up the process. As this is somewhat advanced, and would require more than just a brief comment, I have chosen not to explicitly point to this possibility in the paper.

> *line 269: R gave me a warning on the use of mutate().*

**My response:** I confirm this behaviour on my side, and it turns out that this warning appeared after the 4.0 update of R (see `https://github.com/bstewart/stm/issues/222` for a discussion about this in relation to another package), and relates to the use of formulas in the `fGarch`-package. Fixing it requires an update to this package, so unfortunately I can not get rid of it in this example. It has not changed the results.

I could of course calculate the GARCH-residuals locally on my computer and then load them directly into R, but that would make it harder for a reader to apply the method to his or her own data according to the original paper; Støve et. al. (2014).

I have instead added a comment in the script making the user aware of the issue, and clearly stating that the warning does not affects the result of the analysis.

> *I then obtained a p-value of 0.05 instead of the exact 0 given in the paper on page 13.*

**My response:** Yes, this was a typo in the paper. I re-did the example, but also changed the number of bootstrap samples to 100 in order to make it consistent with the other examples in the paper. The $p$-value of the test in the example is now $0.01$.

> *line 333: code interrupts with the following error message: Error in logspline::logspline(x[, i]) : Not enough unique values*

**My response:** Yes, thank you for noticing. In order to make the examples as lightweight as possible I have demonstrated the methods on very small samples. The sample was a bit too small in this example, which caused problems in one of the dependencies; the `logspline`-package for estimating the marginal densities. I increased the length of the multivariate time series for this example slightly, which has eliminated the problem.

In the last two points, the reviewer was not able to replicate the results in the paper exactly, which is probably due to a mistake in the original submission. The results should now be exactly replicable, also the parts that involve randomness due to bootstrapping.

> *I also got the following error (which is admittedly not related to the project but merely on trying to make the graphs look nicer).*

> `# Initial example 1 --------`

> `Error in pdf(file = "gaussian-example.pdf", height = 4, width = 5, family = "CM Roman")`

> *removing the command family = "CM Roman" solved the issue.*

**My response:** Yes, my apologies. My intention was to make the figures exactly replicate as presented in the paper, but this requires a specific font to be installed on the user system. This is not important in practice, so I have commented out the relevant lines for figure export, and added a remark about this on line 10 in the script.

### 2021-08-16 otneim.R

@@ -6,8 +6,13 @@ library(lubridate)
 library(tidyr)
 library(fGarch)
 library(dplyr)
-library(extrafont)
-loadfonts(quiet = TRUE)
+
+# Font package for exact replication of the plots. You need to install the font
+# CM Roman for this to work, as well as the pdf export commands surrounding the
+# plotting in the code below. 
+
+# library(extrafont) 
+# loadfonts(quiet = TRUE)
 
 # Initial example 1 --------
 
@@ -29,7 +34,7 @@ lg_object1 <- lg_main(x1,
                       plugin_constant_joint = 4)
 dlg_object1 <- dlg(lg_object1, grid = grid1)
 
-pdf(file = "gaussian-example.pdf", height = 4, width = 5, family = "CM Roman")
+# pdf(file = "gaussian-example.pdf", height = 4, width = 5, family = "CM Roman")
 corplot(dlg_object1, plot_thres = .01,
         xlab = "X1", ylab = "X2", 
         main = "",
@@ -39,8 +44,8 @@ corplot(dlg_object1, plot_thres = .01,
         label_size = 4) +
     theme_classic() +
     theme(legend.position = "none")
-dev.off()
-embed_fonts("gaussian-example.pdf")
+# dev.off()
+# embed_fonts("gaussian-example.pdf")
 
 # Initial example 2 --------
 
@@ -59,7 +64,7 @@ lg_object2 <- lg_main(x2,
                       plugin_constant_joint = 2)
 dlg_object2 <- dlg(lg_object2, grid = grid2)
 
-pdf(file = "returns-example.pdf", height = 4, width = 5, family = "CM Roman")
+# pdf(file = "returns-example.pdf", height = 4, width = 5, family = "CM Roman")
 corplot(dlg_object2, plot_thres = .0001,
         xlab = "CAC40", ylab = "FTSE100", 
         main = "",
@@ -71,8 +76,8 @@ corplot(dlg_object2, plot_thres = .0001,
         label_size = 3.5) +
     theme_classic() +
     theme(legend.position = "none")
-dev.off()
-embed_fonts("returns-example.pdf")
+# dev.off()
+# embed_fonts("returns-example.pdf")
 
 # The lg-object ---------
 
@@ -87,19 +92,21 @@ lg_object2 <- lg_main(x, est_method = "5par_marginals_fixed")
 
 lg_object3 <- lg_main(x, est_method = "1par")
 
[…]

### 2021-08-17 otneim.tex
Third submission for the R Journal

@@
 Given a data set \code{x} having $n$ rows and $p \geq 2$ columns, the user must choose between four distinct estimation methods, and specify this choice by using the argument \code{est\_method} to the \code{lg\_main()}-function.
-We look at the built-in bivariate data set `faithful`, which records the waiting time between eruptions and the duration of the eruption for the Old Faithful geyser in the Yellowstone National Park, USA (see the help file in R for more details: \code{?faithful}), and load the \CRANpkg{lg}-package in order to demonstrate the implementation:
+We look at the built-in bivariate data set \code{faithful}, which records the waiting time between eruptions and the duration of the eruption for the Old Faithful geyser in the Yellowstone National Park, USA (see the help file in R for more details: \code{?faithful}), and load the \CRANpkg{lg}-package in order to demonstrate the implementation:
 
@@
 Finally, the only thing that we must alter in order to perform the third test for serial cross-dependence is the bootstrap method.
-In the applications above, it suffices to use the standard bootstrap, where we resammple with replacement from the data.
+In the applications above, it suffices to use the standard bootstrap, where we resample with replacement from the data.
 This is implemented in the \code{ind\_test()}-function by setting the \code{bootstrap\_type}-argument to \code{"plain"}, which is the default option.

### 2021-08-17 response3/response3.Rmd
Third submission for the R Journal

@@
 If we have used the same random seed this figure should be identical to what the reviewer has obtained.
-Whether the observed local correlations represent *significant* (in the statistical sense) dependence can be investigated by looking at the estimated standard deviations, and lower/upper confidence limits which is reporten on the 95% confidence level by default, available in the `dlg_object1` in the `$loc_cor_sd`, `$loc_cor_lower` and `$loc_cor_upper` slots respectively.
-By plotting the estimated correlations with their confidence intervals, as I have done below, we see that zero (represented as a horizontal line) is included in almost all of the intervals.
+Whether the observed local correlations represent *significant*  dependence (in the statistical sense) can be investigated by looking at the estimated standard deviations, and lower/upper confidence limits which is reported on the 95% confidence level by default, available in the `dlg_object1` in the `$loc_cor_sd`, `$loc_cor_lower` and `$loc_cor_upper` slots respectively.
+By plotting the estimated correlations with their confidence intervals, as I have done below (the $x$-axis is just an index of the grid points), we see that zero (represented as a horizontal line) is included in almost all of the intervals.
 
@@
 
-**My response:** I confirm this behaviour on my side, and it turns out that this warning appeared after the 4.0 update of R (see `https://github.com/bstewart/stm/issues/222` for a discussion about this in relation to another package), and relates to the use of formulas in the `fGarch`-package.
-Fixing it requires an update to this package, so unfortunately I can not get rid of it in this example.
+**My response:** I confirm this behavior on my side, and it turns out that this warning appeared after the 4.0 update of R (see `https://github.com/bstewart/stm/issues/222` for a discussion about this in relation to another package), and relates to the use of formulas in the `fGarch`-package.
+Fixing it requires an update to that package, so unfortunately I can not get rid of it in this example.
 It has not changed the results.
@@
 The sample was a bit too small in this example, which caused problems in one of the dependencies; the `logspline`-package for estimating the marginal densities.
-I increased the length of the multivariate time series for this example slightly, which has eliminated the problem.
+I increased the length of the multivariate time series for this example, which has eliminated the problem.
 
@@
 This is not important in practice, so I have commented out the relevant lines for figure export, and added a remark about this on line 10 in the script.
+
+Finally, I have changed the order of lines 182 and 183 in the script in order to take care of a warning message regarding column names, resulting from a recent update in the `tibble`-package.

### 2021-08-17 otneim.R

@@ -179,8 +179,8 @@ var_pairs <- density_object$bw$joint %>%
 
 # pdf(file = "loccor-example.pdf", height = 3, width = 5, family = "CM Roman")
 density_object$loc_cor %>% 
-    as_tibble %>% 
     (function(x) {colnames(x) <- (var_pairs %>% pull); x}) %>% 
+    as_tibble %>%
     mutate(x0 = x0[,1]) %>% 
     gather(pair, cor, -x0) %>% 
     ggplot() +
@@ -353,12 +353,12 @@ lg_object <- lg_main(x,
 dlg_object <- dlg(lg_object, grid = grid)
 
 # Make a dependence map using default setup
-pdf("graphics1.pdf", height = 5, width = 7, family = "CM Roman")
+# pdf("graphics1.pdf", height = 5, width = 7, family = "CM Roman")
 corplot(dlg_object1)
-dev.off()
-embed_fonts("graphics1.pdf")
+# dev.off()
+# embed_fonts("graphics1.pdf")
 
-pdf("graphics2.pdf", height = 5, width = 7, family = "CM Roman")
+# pdf("graphics2.pdf", height = 5, width = 7, family = "CM Roman")
 corplot(dlg_object1, 
         plot_obs = TRUE, 
         plot_thres = 0.01,
@@ -368,5 +368,5 @@ corplot(dlg_object1,
         xlab = "",
         ylab = "") +
     theme_classic()
-dev.off()
-embed_fonts("graphics2.pdf")
+# dev.off()
+# embed_fonts("graphics2.pdf")

### 2021-09-09 otneim.tex
Started on the bibfile

@@
 \citet{tjos:huft:2013} propose the \emph{local Gaussian correlation} (LGC) as a new measure of statistical dependence between two stochastic variables $X_1$ and $X_2$, which has the following important property yet unrivaled in the literature: It can separate between positive and negative nonlinear dependence while still reducing to the ordinary Pearson correlation coefficient if $X_1$ and $X_2$ are jointly normally distributed.
-The R-package \CRANpkg{localgauss} \citep{bere:klep:tjos:2014} provides two important functions in this context; one that calculates the sample LGC based on observed values of $(X_1,X_2)$, and one that uses the estimated LGC to perform a local test of independence between $X_1$ and $X_2$ as described in detail by \citet{bere:tjos:2014}.
+The R-package \CRANpkg{localgauss} \citep{bere:klep:tjos:2014} provides two important functions in this context; one that calculates the sample LGC based on observed values of $\left(X_1,X_2\right)$, and one that uses the estimated LGC to perform a local test of independence between $X_1$ and $X_2$ as described in detail by \citet{bere:tjos:2014}.
 
@@
 
-Consider a random vector $X$ having the unknown probability density function $f_X(x)$.
+Consider a random vector $X$ having the unknown probability density function $f_X\left(x\right)$.
 It is a standard task to estimate $f_X$ based on a random sample $X_1,\ldots, X_n$, and the statistical literature provides an abundance of methods to accomplish this.
-One may, for example, make the assumption that the unknown density function has a particular parametric form, $f_X \in F_{\theta}$, where $F_{\theta} = \{f(x;\theta), \theta \in \Theta\}$ is a family of probability density functions indexed by some parameter $\theta$, and where $\Theta$ is the parameter space.
+One may, for example, make the assumption that the unknown density function has a particular parametric form, $f_X \in F_{\theta}$, where $F_{\theta} = \left\{f(x;\theta), \theta \in \Theta\right\}$ is a family of probability density functions indexed by some parameter $\theta$, and where $\Theta$ is the parameter space.
 Under this assumption we will typically produce an estimate of the parameter $\theta$, written $\widehat \theta$, using the maximum likelihood method.
-The estimated probability density function is then given as $\widehat f_X(x) = f(x; \widehat\theta)$.
+The estimated probability density function is then given as $\widehat f_X\left(x\right) = f\left(x; \widehat\theta\right)$.
 
-A different approach is to estimate $f_X(\cdot)$ without any prior parametric assumptions. The classical method for non-parametric density estimation is the kernel estimator 
-$$\widehat f_X(x) = \frac{1}{nb}\sum_{i=1}^n K\left(\frac{X_i - x}{b}\right),$$
-where $K$ is a symmetric density function (the kernel) and $b$ is a tuning parameter (the bandwidth) that controls the smoothness of the estimate $\widehat f_X(\cdot)$.
+A different approach is to estimate $f_X\left(\cdot\right)$ without any prior parametric assumptions. The classical method for non-parametric density estimation is the kernel estimator 
+$$\widehat f_X\left(x\right) = \frac{1}{nb}\sum_{i=1}^n K\left(\frac{X_i - x}{b}\right),$$
+where $K$ is a symmetric density function (the kernel) and $b$ is a tuning parameter (the bandwidth) that controls the smoothness of the estimate $\widehat f_X\left(\cdot\right)$.
 See \citet{silv:1986} for an introduction to this topic.
@@
 \citet{hjor:jone:1996} provide one such idea.
-They consider a parametric family $F_{\theta}$, but instead of searching for a single parameter value $\theta_0$, for which $f_X(x) = f(x;\theta_0)$ (or approximately so), they rather assert that different members of $F_{\theta}$ may approximate $f_X$ locally in different parts of its domain.
-In other words, they seek to estimate a parameter \emph{function} $\theta_0(x)$ for which $f_X(x) = f(x;\theta_0(x))$ (or approximately so), and do this by maximizing a \emph{local likelihood function} in each point $x$:
+They consider a parametric family $F_{\theta}$, but instead of searching for a single parameter value $\theta_0$, for which $f_X\left(x\right) = f\left(x;\theta_0\right)$ (or approximately so), they rather assert that different members of $F_{\theta}$ may approximate $f_X$ locally in different parts of its domain.
+In other words, they seek to estimate a parameter \emph{function} $\theta_0\left(x\right)$ for which $f_X\left(x\right) = f\left(x;\theta_0\left(x\right)\right)$ (or approximately so), and do this by maximizing a \emph{local likelihood function} in each point $x$:
 \begin{align}
-\widehat\theta(x) 
-&= \arg\max_{\theta \in \Theta} L_n(\theta, x) \nonumber\\ 
-&= \arg\max_{\theta \in \Theta} \frac{1}{nb}\sum_{i=1}^n K\left(\frac{X_i - x}{b}\right)\log f(X_i; \theta) - \int \frac{1}{b}K\left(\frac{y - x}{b}\right)f(y;\theta)\,\textrm{d}y,
+\widehat\theta\left(x\right) 
+&= \arg\max_{\theta \in \Theta} L_n\left(\theta, x\right) \nonumber\\ 
+&= \arg\max_{\theta \in \Theta} \frac{1}{nb}\sum_{i=1}^n K\left(\frac{X_i - x}{b}\right)\log f\left(X_i; \theta\right) - \int \frac{1}{b}K\left(\frac{y - x}{b}\right)f\left(y;\theta\right)\,\textrm{d}y,
 \label{eq:loclik}
@@
 where, again, $K$ is a symmetric density function and $b$ is a bandwidth parameter that controls the smoothness of the estimate.
-The second term in the local likelihood function is a penalty that ensures that the estimated density $\widehat f_X(x) = f(x; \widehat\theta(x))$ converges correctly to the true density function $f_X(x)$ as the sample size $n$ increases to infinity, and the bandwidth $b$ decreases towards zero.
+The second term in the local likelihood function is a penalty that ensures that the estimated density $\widehat f_X\left(x\right) = f\left(x; \widehat\theta\left(x\right)\right)$ converges correctly to the true density function $f_X\left(x\right)$ as the sample size $n$ increases to infinity, and the bandwidth $b$ decreases towards zero.
 See \citet{hjor:jone:1996} for a detailed discussion about this construction.
 
-\citet{tjos:huft:2013} consider the bivariate case $X = (X_1, X_2)$ and take $F_{\theta}$ to be the family of bivariate normal distributions consisting of densities on the form
+\citet{tjos:huft:2013} consider the bivariate case $X = \left(X_1, X_2\right)$ and take $F_{\theta}$ to be the family of bivariate normal distributions consisting of densities on the form
 
 \begin{align}
-f(x;\theta) 
-&= \psi(x_1, x_2;\mu_1, \mu_2, \sigma_1, \sigma_2, \rho)  \nonumber\\ 
+f\left(x;\theta\right) 
+&= \psi\left(x_1, x_2;\mu_1, \mu_2, \sigma_1, \sigma_2, \rho\right)  \nonumber\\ 
 &= \frac{1}{2\pi \sigma_1 \sigma_2\sqrt{1-\rho^2}} \nonumber \\
-& \qquad \times\exp \left\{-\frac{1}{2(1-\rho^2)}\left(\frac{(x_1-\mu_1)^2}{\sigma_1^2}-2\rho\frac{(x_1-\mu_1)(x_2-\mu_2)}{\sigma_1 \sigma_2}+\frac{(x_2-\mu_2)^2}{\sigma_2^2}\right)\right\},
+& \qquad \times\exp \left\{-\frac{1}{2(1-\rho^2)}\left(\frac{\left(x_1-\mu_1\right)^2}{\sigma_1^2}-2\rho\frac{\left(x_1-\mu_1\right)\left(x_2-\mu_2\right)}{\sigma_1 \sigma_2}+\frac{\left(x_2-\mu_2\right)^2}{\sigma_2^2}\right)\right\},
 \label{eq:gaussian}
 \end{align}
-where $\theta = (\mu_1,\mu_2,\sigma_1,\sigma_2,\rho)$ is the vector of parameters.
-Using a sample $\{X_{1i}, X_{2i}\}$, $i = 1, \ldots, n$, they estimate $\theta$ locally in the point $x$ by maximizing the local likelihood function \eqref{eq:loclik}, producing
-$$\widehat\theta(x) = (\widehat\mu_1(x), \widehat\mu_2(x), \widehat\sigma_1(x), \widehat\sigma_2(x), \widehat\rho(x)),$$
-and take special interest in the estimated correlation function $\widehat\rho(x)$ (i.e.
+where $\theta = \left(\mu_1,\mu_2,\sigma_1,\sigma_2,\rho\right)$ is the vector of parameters.
+Using a sample $\left\{X_{1i}, X_{2i}\right\}$, $i = 1, \ldots, n$, they estimate $\theta$ locally in the point $x$ by maximizing the local likelihood function \eqref{eq:loclik}, producing
+$$\widehat\theta\left(x\right) = \left(\widehat\mu_1\left(x\right), \widehat\mu_2\left(x\right), \widehat\sigma_1\left(x\right), \widehat\sigma_2\left(x\right), \widehat\rho\left(x\right)\right),$$
+and take special interest in the estimated correlation function $\widehat\rho\left(x\right)$ (i.e.
 the LGC) because it serves as an attractive local measure of statistical dependence between $X_1$ and $X_2$.
@@
     \centering
-    \begin{subfigure}[t]{0.48\textwidth}
+    \begin{minipage}[b]{0.48\textwidth}
       \includegraphics[width=\textwidth]{gaussian-example}
-      \caption{Estimated local Gaussian correlations between two  jointly normal variables having correlation equal to 0.5, based on 1000 observations.}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.48\textwidth}
+      \subcaption{Estimated local Gaussian correlations between  \newline two jointly normal variables having correlation  \newline  equal to 0.5, based on 1000 observations.}
+    \end{minipage}
+    \begin{minipage}[b]{0.48\textwidth}
       \includegraphics[width=\textwidth]{returns-example}
-      \caption{Local Gaussian correlation between daily returns from the CAC40 and FTSE100 stock indices, based on 1000 consecutive trading days.}
-    \end{subfigure}
+      \subcaption{Local Gaussian correlation between daily returns from the CAC40 and FTSE100 stock indices, based on 1000 consecutive trading days.}
+    \end{minipage}
     \caption{Two dependence maps}
@@
 Taking the LGC as a measure of dependence opens up for a number of possibilities to construct statistical tests.
-\citet{bere:tjos:2014} show that $\rho(x) \equiv 0$ implies that $X_1$ and $X_2$ are independent.
-They show further that independence between $X_1$ and $X_2$ implies $\rho(x) \equiv 0$ if the population values of the local mean and standard deviation functions satisfy the following conditions: $\mu_i(x_1, x_2) = \mu_i(x_i)$ and $\sigma_i(x_1, x_2) = \sigma_i(x_i)$ for $i=1,2$.
-Equivalence between independence and $\rho(x) \equiv 0$ holds in general if the observations have been suitably transformed according to a procedure presented later in this section.
-It follows then that departures from $\widehat\rho(x) \equiv 0$ may be taken as evidence against the hypothesis that $X_1$ and $X_2$ are statistically independent.
-\citet{bere:tjos:2014} formalize this notion by testing whether $\rho(x) \equiv 0$ for all $x\in S \subset \mathbb{R}^2$ using the test statistic
+\citet{bere:tjos:2014} show that $\rho\left(x\right) \equiv 0$ implies that $X_1$ and $X_2$ are independent.
+They show further that independence between $X_1$ and $X_2$ implies $\rho\left(x\right) \equiv 0$ if the population values of the local mean and standard deviation functions satisfy the following conditions: $\mu_i\left(x_1, x_2\right) = \mu_i\left(x_i\right)$ and $\sigma_i\left(x_1, x_2\right) = \sigma_i\left(x_i\right)$ for $i=1,2$.
+Equivalence between independence and $\rho\left(x\right) \equiv 0$ holds in general if the observations have been suitably transformed according to a procedure presented later in this section.
+It follows then that departures from $\widehat\rho\left(x\right) \equiv 0$ may be taken as evidence against the hypothesis that $X_1$ and $X_2$ are statistically independent.
+\citet{bere:tjos:2014} formalize this notion by testing whether $\rho\left(x\right) \equiv 0$ for all $x\in S \subset \mathbb{R}^2$ using the test statistic
 
 \begin{equation}
-T_{n,b} = \int_S h\left(\widehat \rho(x) \right) \,\textrm{d}F_n(x)
+T_{n,b} = \int_S h\left(\widehat \rho\left(x\right) \right) \,\textrm{d}F_n(x)
 \label{eq:test-statistic1}
 \end{equation}
-for some non-negative function $h$, for example $h(x) = x^2$ or $h(x) = |x|$.
+for some non-negative function $h$, for example $h\left(x\right) = x^2$ or $h\left(x\right) = |x|$.
 Critical values may be obtained by permutations of the data under the null hypothesis, and we demonstrate the implementation of this test using the \CRANpkg{lg}-package in Section \ref{chap:indtests}.
 
-Consider next the stationary time series $\{X_t\}$.
-The autocorrelation function (ACF) $\rho_k = \rho(X_t, X_{t-k})$ is a well known concept for describing the serial dependence in the time series, but the ACF is, again, only capable to completely capture \emph{linear} serial dependence.
+Consider next the stationary time series $\left\{X_t\right\}$.
+The autocorrelation function (ACF) $\rho_k = \rho\left(X_t, X_{t-k}\right)$ is a well known concept for describing the serial dependence in the time series, but the ACF is, again, only capable to completely capture \emph{linear} serial dependence.
 \citet{laca:tjos:2017} seek to remedy this by rather calculating the the local correlation between $X_t$ and $X_{t-k}$.
@@
 In fact, this work is mainly a theoretical exercise in order to accommodate time series dependence.
-Testing for independence between $X_t$ and $X_{t-k}$ using observations $\{X_t, X_{t-k}\}_{t = k+1}^T$ leads to the same test statistic \eqref{eq:test-statistic1} and bootstrap procedure as the test for independence between $X_1$ and $X_2$ that we described above.
+Testing for independence between $X_t$ and $X_{t-k}$ using observations $\left\{X_t, X_{t-k}\right\}_{t = k+1}^T$ leads to the same test statistic \eqref{eq:test-statistic1} and bootstrap procedure as the test for independence between $X_1$ and $X_2$ that we described above.
 
-\citet{laca:tjos:2018} extend this problem to test for serial cross-dependence between two time series $\{X_t, Y_t\}$ by measuring the LGC between $X_t$ and $Y_{t-k}$.
-Departures from $\widehat\rho(x,y) \equiv 0$ is again taken as evidence against independence, and the test statistic (\ref{eq:test-statistic1}) provides an aggregate measure of this discrepancy in the specified region $S$.
+\citet{laca:tjos:2018} extend this problem to test for serial cross-dependence between two time series $\left\{X_t, Y_t\right\}$ by measuring the LGC between $X_t$ and $Y_{t-k}$.
+Departures from $\widehat\rho\left(x,y\right) \equiv 0$ is again taken as evidence against independence, and the test statistic (\ref{eq:test-statistic1}) provides an aggregate measure of this discrepancy in the specified region $S$.
 In this case, however, we can not obtain replicates of the test statistic under the null hypothesis by simple permutations of the data.
@@
 2223]{forb:rigo:2002} and employ the LGC to quantify this potential linkage.
-The authors estimate the LGC on a grid $\{x_1, x_2\}_k$, $k=1,\ldots,K$ along the diagonal $D=\{(x_1,x_2): x_1 = x_2\}$ \emph{before} and \emph{after} some critical event in the financial markets, denoted as the crisis (C) and the non-crisis (NC) periods respectively.
+The authors estimate the LGC on a grid $\left\{x_1, x_2\right\}_k$, $k=1,\ldots,K$ along the diagonal $D=\left\{\left(x_1,x_2\right): x_1 = x_2\right\}$ \emph{before} and \emph{after} some critical event in the financial markets, denoted as the crisis (C) and the non-crisis (NC) periods respectively.
 They compare the two estimates using the following test statistic;
+$$T_{n,b}^D = \sum_{k=1}^K \left\{\widehat\rho_C\left(x_k, x_k\right) - \widehat\rho_{NC}\left(x_k, x_k\right) \right\}w\left(x_k,x_k\right),$$
 
-$$T_{n,b}^D = \sum_{k=1}^K \left\{\widehat\rho_C(x_k, x_k) - \widehat\rho_{NC}(x_k, x_k) \right\}w(x_k,x_k),$$
-
-where $w(\cdot, \cdot)$ is a weight function that serve the same purpose as the integration area $S$ in (\ref{eq:test-statistic1}).
+where $w\left(\cdot, \cdot\right)$ is a weight function that serve the same purpose as the integration area $S$ in (\ref{eq:test-statistic1}).
 In this case, \citet{stov:tjos:huft:2014} show that a standard bootstrap will suffice in order to produce approximate replicates of $T_{n,b}^D$ under the null hypothesis of no financial contagion, and we demonstrate the implementation of this test using the \CRANpkg{lg}-package in Section \ref{chap:cont}.
@@
 This is a consequence of the curse of dimensionality, the effect of which is sought remedied by an algorithm provided by \citet{otne:tjos:2017}.
-The idea is to fit the $p$-variate normal distribution $\psi(\mu, \Sigma)$ locally, where $\mu$ is the vector of $p$ expectations, and $\Sigma$ is the $p\times p$ covariance matrix (to which the correlation matrix $R$ corresponds), but under the following structural simplifications:
+The idea is to fit the $p$-variate normal distribution $\psi\left(\mu, \Sigma\right)$ locally, where $\mu$ is the vector of $p$ expectations, and $\Sigma$ is the $p\times p$ covariance matrix (to which the correlation matrix $R$ corresponds), but under the following structural simplifications:
 
 \begin{align}
-\mu_i(x) &= \mu_i(x_1, \ldots, x_p) \stackrel{\textrm{def}}{=} \mu_i(x_i) \label{eq:simp-mu}\\
-\sigma_i(x) &= \sigma_i(x_1, \ldots, x_p)  \stackrel{\textrm{def}}{=} \sigma_i(x_i) \label{eq:simp-sig} \\
-\rho_{ij}(x) &= \rho_{ij}(x_1, \ldots, x_p) \stackrel{\textrm{def}}{=} \rho_{ij}(x_i, x_j). \label{eq:simp-rho}
+\mu_i\left(x\right) &= \mu_i\left(x_1, \ldots, x_p\right) \stackrel{\textrm{def}}{=} \mu_i\left(x_i\right) \label{eq:simp-mu}\\
+\sigma_i\left(x\right) &= \sigma_i\left(x_1, \ldots, x_p\right)  \stackrel{\textrm{def}}{=} \sigma_i\left(x_i\right) \label{eq:simp-sig} \\
+\rho_{ij}\left(x\right) &= \rho_{ij}\left(x_1, \ldots, x_p\right) \stackrel{\textrm{def}}{=} \rho_{ij}\left(x_i, x_j\right). \label{eq:simp-rho}
 \end{align}
@@
 \ref{eq:simp-rho}).
-In the second step the estimates $\widehat\mu_i(x_i)$, $\widehat\mu_j(x_j)$, $\widehat\sigma_i(x_i)$ and $\widehat\sigma_j(x_j)$ are kept fixed in the estimation of the pairwise local correlation.
+In the second step the estimates $\widehat\mu_i\left(x_i\right)$, $\widehat\mu_j\left(x_j\right)$, $\widehat\sigma_i\left(x_i\right)$ and $\widehat\sigma_j\left(x_j\right)$ are kept fixed in the estimation of the pairwise local correlation.
 They argue further that the following transformation technique will produce better density estimates in many situations.
-The motivation for introducing the simplifications defined in equations \ref{eq:simp-mu}-\ref{eq:simp-rho} can be compared to the practical advantages of estimating additive regression models, where $\textrm{E}(Y) = f(x_1, \ldots, x_p) \stackrel{\textrm{def}}{=} f_1(x_1) + \cdots + f_p(x_p)$.
+The motivation for introducing the simplifications defined in equations \ref{eq:simp-mu}-\ref{eq:simp-rho} can be compared to the practical advantages of estimating additive regression models, where $\textrm{E}\left(Y\right) = f\left(x_1, \ldots, x_p\right) \stackrel{\textrm{def}}{=} f_1\left(x_1\right) + \cdots + f_p\left(x_p\right)$.
 
-Denote by $F_i(x_i)$, $i = 1,\ldots, p$ the marginal distribution functions of the stochastic vector $X$, and by $\widehat F_i(x_i) = n^{-1}\sum_{i=1}^n 1(X_i \leq x_i)$ their empirical counterparts.
-They then estimate the density $f_Z(z)$ of the vector $Z = \{\Phi^{-1}(F_i(X_i))\}_{i=1,\ldots,p}$, in practice approximated by
+Denote by $F_i\left(x_i\right)$, $i = 1,\ldots, p$ the marginal distribution functions of the stochastic vector $X$, and by $\widehat F_i\left(x_i\right) = n^{-1}\sum_{i=1}^n 1\left(X_i \leq x_i\right)$ their empirical counterparts.
+They then estimate the density $f_Z\left(z\right)$ of the vector $Z = \left\{\Phi^{-1}\left(F_i\left(X_i\right)\right)\right\}_{i=1,\ldots,p}$, in practice approximated by
 
 \begin{equation}
-\widehat Z = \{\Phi^{-1}(\widehat F_i(X_i))\}_{i=1,\ldots,p}, 
+\widehat Z = \left\{\Phi^{-1}\left(\widehat F_i\left(X_i\right)\right)\right\}_{i=1,\ldots,p}, 
 \label{eq:trans}
 \end{equation}
-and where $\Psi(\cdot)$ is the univariate standard normal cdf. In that case, they simplify the estimation problem even further and fix
+and where $\Psi\left(\cdot\right)$ is the univariate standard normal cdf. In that case, they simplify the estimation problem even further and fix
 
 \begin{equation}
-\mu_i(z_i) \stackrel{\textrm{def}}{=} 0 \,\, \textrm{ and } \sigma_i(z_i) \stackrel{\textrm{def}}{=} 1, \,\, i = 1,\ldots,p,
+\mu_i\left(z_i\right) \stackrel{\textrm{def}}{=} 0 \,\, \textrm{ and } \sigma_i\left(z_i\right) \stackrel{\textrm{def}}{=} 1, \,\, i = 1,\ldots,p,
 \label{eq:simp}
 \end{equation}
-so that the only parameter functions left to estimate are the pairwise local Gaussian correlations $R(z) = \{\rho_{ij}(z_i, z_j)\}_{i<j}$.
+so that the only parameter functions left to estimate are the pairwise local Gaussian correlations $R\left(z\right) = \left\{\rho_{ij}\left(z_i, z_j\right)\right\}_{i<j}$.
 We use the notation $Z$, $z_i$ and $z_j$ to signify that the estimation is performed on the (approximate) standard normal scale, or $z$-scale for short.
-We can then estimate the joint density $f_Z(z)$ of $Z$ as
+We can then estimate the joint density $f_Z\left(z\right)$ of $Z$ as
 \begin{equation}
-\widehat f_Z(z) = \psi(z; \mu(z) = 0, \sigma(z) = 1, R = \widehat R(z)),
+\widehat f_Z\left(z\right) = \psi\left(z; \mu\left(z\right) = 0, \sigma\left(z\right) = 1, R = \widehat R\left(z\right)\right),
 \label{eq:transformed-density}
 \end{equation}
-where $\mu(z) = \{\mu_i(z)\}$ and $\sigma(z) = \{\sigma_i(z)\}$ for $i=1,\ldots,p$, and then substitute $f_Z$ for $\widehat f_Z$ in the following relation obtained by \citet{otne:tjos:2017} in order to estimate the unknown density $f_X$:
+where $\mu\left(z\right) = \left\{\mu_i\left(z\right)\right\}$ and $\sigma\left(z\right) = \left\{\sigma_i\left(z\right)\right\}$ for $i=1,\ldots,p$, and then substitute $f_Z$ for $\widehat f_Z$ in the following relation obtained by \citet{otne:tjos:2017} in order to estimate the unknown density $f_X$:
 \begin{equation}
-f(x) = f_{Z}\big(\Phi^{-1}(F_1(x_1)), \ldots, \Phi^{-1}(F_p(x_p))\big) \times \prod_{i=1}^p \frac{f_i(x_i)}{\phi\big(\Phi^{-1}(F_i(x_i))\big)},
+f\left(x\right) = f_{Z}\big(\Phi^{-1}\left(F_1\left(x_1\right)\right), \ldots, \Phi^{-1}\left(F_p\left(x_p\right)\right)\big) \times \prod_{i=1}^p \frac{f_i\left(x_i\right)}{\phi\big(\Phi^{-1}\left(F_i\left(x_i\right)\right)\big)},
 \label{eq:backtrans}
 \end{equation}
-where $\phi(\cdot)$ is the standard normal pdf. This estimator is implemented the \CRANpkg{lg}-package as demonstrated in Section \ref{chap:density}.
+where $\phi\left(\cdot\right)$ is the standard normal pdf. This estimator is implemented the \CRANpkg{lg}-package as demonstrated in Section \ref{chap:density}.
 
-One particular feature enjoyed by the jointly normally distributed vector $X$ is that, for any partitioning $X = \left(X^{(1)}, X^{(2)}\right)$, the conditional distribution of $X^{(1)}|X^{(2)} = x^{(2)}$ is also normal.
-In fact, if $X \sim \mathcal{N}(\mu,\Sigma)$, and $\mu$ and $\Sigma$ is partitioned according to $\left(X^{(1)}, X^{(2)}\right)$ as
+One particular feature enjoyed by the jointly normally distributed vector $X$ is that, for any partitioning $X = \left(X^{\left(1\right)}, X^{\left(2\right)}\right)$, the conditional distribution of $X^{\left(1\right)}|X^{\left(2\right)} = x^{\left(2\right)}$ is also normal.
+In fact, if $X \sim \mathcal{N}\left(\mu,\Sigma\right)$, and $\mu$ and $\Sigma$ is partitioned according to $\left(X^{\left(1\right)}, X^{\left(2\right)}\right)$ as
 $$\mu = \begin{pmatrix} \mu_1 \\\mu_2 \end{pmatrix} \,\, \textrm{ and } \,\, \Sigma = \begin{pmatrix} \Sigma_{11} & \Sigma_{12} \\ \Sigma_{21} & \Sigma_{22} \end{pmatrix},$$
-then $X^{(1)}|X^{(2)} = x^{(2)} \sim \mathcal{N}(\mu^*, \Sigma^*)$, where
+then $X^{\left(1\right)}|X^{\left(2\right)} = x^{\left(2\right)} \sim \mathcal{N}\left(\mu^*, \Sigma^*\right)$, where
 
 \begin{align}
-\mu^* &= \mu_1 + \Sigma_{12}\Sigma_{22}^{-1}\left(x^{(2)} - \mu_2\right) \label{eq:cond-mu} \\
+\mu^* &= \mu_1 + \Sigma_{12}\Sigma_{22}^{-1}\left(x^{\left(2\right)} - \mu_2\right) \label{eq:cond-mu} \\
 \Sigma^* &= \Sigma_{11} - \Sigma_{12}\Sigma_{22}^{-1}\Sigma_{21}, \label{eq:cond-sig}
@@
 \citet{otne:tjos:2018} demonstrate that this property may be translated into a corresponding local argument without modification.
-That is, if the joint density $f_X(\cdot)$ can be written on a locally Gaussian form
-
-$$f_X(x) = \psi(x, \mu(x), \Sigma(x)),$$
-then the conditional density of $X^{(1)}|X^{(2)} = x^{(2)}$ can also be written on the same locally Gaussian form with local parameters given by equations (\ref{eq:cond-mu}) and (\ref{eq:cond-sig}), except that all quantities are $x$-dependent.
+That is, if the joint density $f_X\left(\cdot\right)$ can be written on a locally Gaussian form
+$$f_X\left(x\right) = \psi\left(x, \mu\left(x\right), \Sigma\left(x\right)\right),$$
+then the conditional density of $X^{\left(1\right)}|X^{\left(2\right)} = x^{\left(2\right)}$ can also be written on the same locally Gaussian form with local parameters given by equations (\ref{eq:cond-mu}) and (\ref{eq:cond-sig}), except that all quantities are $x$-dependent.
 If we use the transformation technique described above together with simplification (\ref{eq:simp}), the local versions of equations (\ref{eq:cond-mu}) and (\ref{eq:cond-sig}) simplify to
@@
 \begin{align}
-\mu^*(z) &= R_{12}(z)R_{22}(z)^{-1}z^{(2)},  \label{eq:cond-mu-trans}\\
-\Sigma^*(z) &= R_{11}(z) - R_{12}(z)R_{22}(z)^{-1}R_{21}(z), \label{eq:cond-sig-trans}
+\mu^*\left(z\right) &= R_{12}\left(z\right)R_{22}\left(z\right)^{-1}z^{\left(2\right)},  \label{eq:cond-mu-trans}\\
+\Sigma^*\left(z\right) &= R_{11}\left(z\right) - R_{12}\left(z\right)R_{22}\left(z\right)^{-1}R_{21}\left(z\right), \label{eq:cond-sig-trans}
 \end{align}
 where we, again, switch to $z$-notation in order to make it clear that these quantities are estimated on the standard normal $z$-scale.
-An estimator $\widehat f_{X^{(1)}|X^{(2)}}(\cdot|\cdot)$ of the conditional density $f_{X^{(1)}|X^{(2)}}(\cdot|\cdot)$ follows immediately from an expression corresponding to (\ref{eq:backtrans}), and the \CRANpkg{lg}-package provides functions for implementing this estimator in R.
+An estimator $\widehat f_{X^{\left(1\right)}|X^{\left(2\right)}}\left(\cdot|\cdot\right)$ of the conditional density $f_{X^{\left(1\right)}|X^{\left(2\right)}}\left(\cdot|\cdot\right)$ follows immediately from an expression corresponding to (\ref{eq:backtrans}), and the \CRANpkg{lg}-package provides functions for implementing this estimator in R.
 We describe the implementation of this functionality in Section \ref{chap:density}.
@@
 Finally, we refer to \citet{otne:tjos:2019} who take the local version of the conditional covariance matrix (\ref{eq:cond-sig}) (or (\ref{eq:cond-sig-trans}) in the transformed case) as a measure for conditional dependence, and thus as an instrument to test for conditional independence.
-Consider the stochastic vector $X = \left(X^{(1)}, X^{(2)}, X^{(3)}\right)$, where $X^{(1)}$ and $X^{(2)}$ are scalars and $X^{(3)}$ may be a vector.
-$X^{(1)}$ is conditionally independent from $X^{(2)}$ given $X^{(3)}$, written $X^{(1)} \perp X^{(2)} \,|\,X^{(3)}$, if the stochastic variables $X^{(1)}\,|\,X^{(3)}$ and $X^{(2)}\,|\,X^{(3)}$ are independent, or, equivalently, if the joint conditional density of $X^{(1)}$ and $X^{(2)}$ given $X^{(3)}$ can be written as the product
+Consider the stochastic vector $X = \left(X^{\left(1\right)}, X^{\left(2\right)}, X^{\left(3\right)}\right)$, where $X^{\left(1\right)}$ and $X^{\left(2\right)}$ are scalars and $X^{\left(3\right)}$ may be a vector.
+$X^{\left(1\right)}$ is conditionally independent from $X^{\left(2\right)}$ given $X^{\left(3\right)}$, written $X^{\left(1\right)} \perp X^{\left(2\right)} \,|\,X^{\left(3\right)}$, if the stochastic variables $X^{\left(1\right)}\,|\,X^{\left(3\right)}$ and $X^{\left(2\right)}\,|\,X^{\left(3\right)}$ are independent, or, equivalently, if the joint conditional density of $X^{\left(1\right)}$ and $X^{\left(2\right)}$ given $X^{\left(3\right)}$ can be written as the product
 
 \begin{equation}
-f_{X^{(1)},X^{(2)}|X^{(3)}}\left(x^{(1)}, x^{(2)}|x^{(3)}\right) = f_{X^{(1)}|X^{(3)}}\left(x^{(1)}|x^{(3)}\right) \times f_{X^{(2)}|X^{(3)}}\left(x^{(2)}|x^{(3)}\right).
+f_{X^{\left(1\right)},X^{\left(2\right)}|X^{\left(3\right)}}\left(x^{\left(1\right)}, x^{\left(2\right)}|x^{\left(3\right)}\right) = f_{X^{\left(1\right)}|X^{\left(3\right)}}\left(x^{\left(1\right)}|x^{\left(3\right)}\right) \times f_{X^{\left(2\right)}|X^{\left(3\right)}}\left(x^{\left(2\right)}|x^{\left(3\right)}\right).
 \label{eq:cond-ind}
 \end{equation}
-In this case, denote by $\alpha(z)$ the off-diagonal element in the $2\times2$ local correlation matrix $R^*(z)$ (which derives directly from $\Sigma^*(z)$ as given in (\ref{eq:cond-sig-trans})).
-If $X$ has a local Gaussian distribution, the conditional independence (\ref{eq:cond-ind}) is equivalent to $\alpha(z) \equiv 0$, and \citet{otne:tjos:2019} take departures from this relation as evidence against the hypothesis of conditional independence between $X^{(1)}$ and $X^{(2)}$ given $X^{(3)}$.
+In this case, denote by $\alpha\left(z\right)$ the off-diagonal element in the $2\times2$ local correlation matrix $R^*\left(z\right)$ (which derives directly from $\Sigma^*\left(z\right)$ as given in (\ref{eq:cond-sig-trans})).
+If $X$ has a local Gaussian distribution, the conditional independence (\ref{eq:cond-ind}) is equivalent to $\alpha\left(z\right) \equiv 0$, and \citet{otne:tjos:2019} take departures from this relation as evidence against the hypothesis of conditional independence between $X^{\left(1\right)}$ and $X^{\left(2\right)}$ given $X^{\left(3\right)}$.
 The natural way to quantify this is the test functional
@@
 \begin{equation}
-T^{CI}_{n, b} = \int h(\widehat\alpha(z)) \, \textrm{d}F_n(z).
+T^{CI}_{n, b} = \int h\left(\widehat\alpha\left(z\right)\right) \, \textrm{d}F_n(z).
 \label{eq:ci-test-statistic}
@@
 \textbf{1.
-A full locally Gaussian fit for bivariate data.} If $p = 2$, we may fit the bivariate normal $\psi(x, \theta)$ locally to $f(x)$, and by a "full local fit" we mean that we jointly estimate the five parameters
-$$\theta(x) = \big(\mu_1(x_1, x_2), \mu_2(X_1, x_2), \sigma_1(x_1, x_2), \sigma_2(x_1,x_2),\rho(x_1, x_2)\big)$$
-by optimizing the local likelihood function (\ref{eq:loclik}) in the grid point $x = (x_1, x_2)$.
+A full locally Gaussian fit for bivariate data.} If $p = 2$, we may fit the bivariate normal $\psi\left(x, \theta\right)$ locally to $f\left(x\right)$, and by a "full local fit" we mean that we jointly estimate the five parameters
+$$\theta\left(x\right) = \big(\mu_1\left(x_1, x_2\right), \mu_2\left(x_1, x_2\right), \sigma_1\left(x_1, x_2\right), \sigma_2\left(x_1,x_2\right),\rho\left(x_1, x_2\right)\big)$$
+by optimizing the local likelihood function (\ref{eq:loclik}) in the grid point $x = \left(x_1, x_2\right)$.
 To use this estimation method in the subsequent analysis, specify \code{est\_method = "5par"} in the call to \code{lg\_main()}:
@@
 \begin{itemize}
-\item[1.] Calculate $\mu_i(x_i)$ and $\sigma_i(x_i)$, $i = 1,\ldots,p$ by fitting the univariate normal distribution locally to each \emph{marginal} density $f_i(x_i)$ of $f(x)$.
-\item[2.] Keep $\widehat\mu_i(x)$ and $\widehat\sigma_i(x_i)$, $i=1,\ldots,p$ from step 1 fixed when estimating $\rho_{ij}(x_i, x_j)$, $1\leq i<j\leq p$ by fitting the bivariate normal distribution to each \emph{pair} of variables $X_i$ and $X_j$.
+\item[1.] Calculate $\mu_i\left(x_i\right)$ and $\sigma_i\left(x_i\right)$, $i = 1,\ldots,p$ by fitting the univariate normal distribution locally to each \emph{marginal} density $f_i\left(x_i\right)$ of $f\left(x\right)$.
+\item[2.] Keep $\widehat\mu_i\left(x\right)$ and $\widehat\sigma_i\left(x_i\right)$, $i=1,\ldots,p$ from step 1 fixed when estimating $\rho_{ij}\left(x_i, x_j\right)$, $1\leq i<j\leq p$ by fitting the bivariate normal distribution to each \emph{pair} of variables $X_i$ and $X_j$.
 \end{itemize}
@@
 \textbf{4.
-A full locally Gaussian fit for trivariate data.} If the number of variables $p$ is equal to $3$ \emph{and} we choose to transform the data to marginal standard normality (see the next sub-section), the transformed density $f_Z(\cdot)$ in (\ref{eq:transformed-density}) may be estimated by jointly estimating the three local correlations $\rho_{12}(z_1, z_2, z_3)$, $\rho_{13}(z_1, z_2, z_3)$ and $\rho_{23}(z_1, z_2, z_3)$.
+A full locally Gaussian fit for trivariate data.} If the number of variables $p$ is equal to $3$ \emph{and} we choose to transform the data to marginal standard normality (see the next sub-section), the transformed density $f_Z\left(\cdot\right)$ in (\ref{eq:transformed-density}) may be estimated by jointly estimating the three local correlations $\rho_{12}\left(z_1, z_2, z_3\right)$, $\rho_{13}\left(z_1, z_2, z_3\right)$ and $\rho_{23}\left(z_1, z_2, z_3\right)$.
 This estimation method was introduced recently by \citet{otne:tjos:2019} in order to increase power of their conditional independence test, but it can be used in any application described in this paper that consider trivariate data.
@@
     \centering
-    \begin{subfigure}[t]{0.48\textwidth}
+    \begin{minipage}[t]{0.48\textwidth}
       \includegraphics[width=\textwidth]{faithful}
-      \caption{The original data.}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.48\textwidth}
+      \subcaption{The original data.}
+    \end{minipage}
+    \begin{minipage}[t]{0.48\textwidth}
       \includegraphics[width=\textwidth]{faithful-transformed}
-      \caption{The same data transformed to marginal standard normality.}
-    \end{subfigure}
+      \subcaption{The same data transformed to marginal standard normality.}
+    \end{minipage}
     \caption{The same data on different scales.}
@@
 Indeed, if we concentrate on the univariate case for the moment and assume that the (single) bandwidth \(b\) is small, we see from the local likelihood function \eqref{eq:loclik} that only the very few observations closest to a fixed grid point \(x_0\) will have significant weight when determining the
-local parameter estimates \(\widehat\theta_0(x)\) at that point.
-Moving on to another nearby point \(x_1\) may then lead to a fairly different estimate \(\widehat\theta(x_1)\) because the set of observations having weight in this point is very different.
-This may again lead to rougher parameter estimates \(\widehat\theta(x)\) and in turn also to rougher density estimates \(f(x,\widehat\theta(x))\).
+local parameter estimates \(\widehat\theta_0\left(x\right)\) at that point.
+Moving on to another nearby point \(x_1\) may then lead to a fairly different estimate \(\widehat\theta\left(x_1\right)\) because the set of observations having weight in this point is very different.
+This may again lead to rougher parameter estimates \(\widehat\theta\left(x\right)\) and in turn also to rougher density estimates \(f\left(x,\widehat\theta\left(x\right)\right)\).
 
-If the bandwidth \(b\) grows large, on the other hand, all observations receive similar weights, and furthermore: the local likelihood function \eqref{eq:loclik} becomes approximately proportional to the ordinary (global) likelihood function \(L_n(\theta) = \sum_{i=1}^n\log f(X_i;\theta)\).
-In other words, the local parameter estimates \(\widehat \theta(x)\) are smoothed towards the constant maximum likelihood estimates \(\widehat\theta_{ML}\), and the estimated density \(f(x;\widehat\theta(x))\) towards the maximum likelihood estimate \(f(x;\widehat\theta_{ML})\).
-This means that the bandwidth may be chosen to reflect the goodness-of-fit of \(f(x;\theta)\) to the true density \(f(x)\).
+If the bandwidth \(b\) grows large, on the other hand, all observations receive similar weights, and furthermore: the local likelihood function \eqref{eq:loclik} becomes approximately proportional to the ordinary (global) likelihood function \(L_n\left(\theta\right) = \sum_{i=1}^n\log f\left(X_i;\theta\right)\).
+In other words, the local parameter estimates \(\widehat \theta\left(x\right)\) are smoothed towards the constant maximum likelihood estimates \(\widehat\theta_{ML}\), and the estimated density \(f\left(x;\widehat\theta\left(x\right)\right)\) towards the maximum likelihood estimate \(f\left(x;\widehat\theta_{ML}\right)\).
+This means that the bandwidth may be chosen to reflect the goodness-of-fit of \(f\left(x;\theta\right)\) to the true density \(f\left(x\right)\).
 
@@
 
-\[CV(b) = -\frac{1}{n}\sum_{i=1}^n \log f\Big(X_i; \widehat\theta^{(-i)}(X_i)\Big),\]
-where \(\widehat\theta^{(-i)}(x)\) is the parameter estimate obtained after deleting observation \(X_i\) from the data, is proportional to a quantity that estimates the Kullback-Leibler distance between \(f(\cdot, \widehat\theta(\cdot))\) and the true density \(f(\cdot)\), see \citet{bere:tjos:2014}.
-The cross-validated bandwidth \(b_{CV}\) is hence given by \[b_{CV} = \arg\min_b \,\, CV(b).\] If we, for example, wish to use the simplified estimation procedure on the transformed data, we need bandwidths for the marginal estimates of the local means and standard deviations, as well as a \(2\times2\) diagonal bandwidth matrix for each pair of variables.
+\[CV\left(b\right) = -\frac{1}{n}\sum_{i=1}^n \log f\Big(X_i; \widehat\theta^{\left(-i\right)}\left(X_i\right)\Big),\]
+where \(\widehat\theta^{\left(-i\right)}\left(x\right)\) is the parameter estimate obtained after deleting observation \(X_i\) from the data, is proportional to a quantity that estimates the Kullback-Leibler distance between \(f\left(\cdot, \widehat\theta\left(\cdot\right)\right)\) and the true density \(f\left(\cdot\right)\), see \citet{bere:tjos:2014}.
+The cross-validated bandwidth \(b_{CV}\) is hence given by \[b_{CV} = \arg\min_b \,\, CV\left(b\right).\] If we, for example, wish to use the simplified estimation procedure on the transformed data, we need bandwidths for the marginal estimates of the local means and standard deviations, as well as a \(2\times2\) diagonal bandwidth matrix for each pair of variables.
 This is accomplished by the following call to \code{lg\_main()}:
@@
 \(p(p-1)/2\) bandwidth matrices, one for each pair of variables (which in this bivariate example just \emph{one} variable pair, (\code{x1,\ x2})).
-The convergence flags stem from the built-in R functions \code{optim()} and \code{optimize()} that we use to obtain the minimizer of \(CV(\cdot)\), and 0 indicates successful convergence.
+The convergence flags stem from the built-in R functions \code{optim()} and \code{optimize()} that we use to obtain the minimizer of \(CV\left(\cdot\right)\), and 0 indicates successful convergence.
 
@@
 
-\citet{otne:tjos:2017} show that the simplified version of the local Gaussian fit have the same convergence rates as the corresponding nonparametric kernel density estimator, for which \citet{silv:1986} derives the plug-in formula \(b = 1.08\cdot sd(x) \cdot n^{-1/5}\).
+\citet{otne:tjos:2017} show that the simplified version of the local Gaussian fit have the same convergence rates as the corresponding nonparametric kernel density estimator, for which \citet{silv:1986} derives the plug-in formula \(b = 1.08\cdot sd\left(x\right) \cdot n^{-1/5}\).
 By
@@
 \code{bw} & The bandwidths to use if already calculated & \code{NULL} \\
-\code{plugin{\textunderscore}constant{\textunderscore}marginal} & Prop. const. in plugin formula for marg. bw. & $1.75$ \\
-\code{plugin{\textunderscore}exponent{\textunderscore}marginal} & Exponent in plugin formula for marg. bw.& $-1/5$ \\
-\code{plugin{\textunderscore}constant{\textunderscore}joint} & Prop. const. in plugin formula for joint bw. & $1.75$ \\
-\code{plugin{\textunderscore}exponent{\textunderscore}joint} & Exponent in plugin formula for joint bw.& $-1/6$ \\
-\code{tol{\textunderscore}marginal} & Abs. tolerance when optimizing $CV(b)$, marg.& $10^{-3}$ \\
-\code{tol{\textunderscore}joint} &Abs. tolerance when optimizing $CV(b)$, joint& $10^{-3}$ \\
+\code{plugin{\textunderscore}constant{\textunderscore}} &  &  \\
+\code{  marginal} & Prop. const. in plugin formula for marg. bw. & $1.75$ \\
+\code{plugin{\textunderscore}exponent{\textunderscore}} & &  \\
+\code{  marginal} & Exponent in plugin formula for marg. bw.& $-1/5$ \\
+\code{plugin{\textunderscore}constant{\textunderscore}} &  &  \\
+\code{joint} & Prop. const. in plugin formula for joint bw. & $1.75$ \\
+\code{plugin{\textunderscore}exponent{\textunderscore}} & &  \\
+\code{  joint} & Exponent in plugin formula for joint bw.& $-1/6$ \\
+\code{tol{\textunderscore}marginal} & Abs. tolerance when optimizing $CV\left(b\right)$, marg.& $10^{-3}$ \\
+\code{tol{\textunderscore}joint} &Abs. tolerance when optimizing $CV\left(b\right)$, joint& $10^{-3}$ \\
 \bottomrule
@@
 \item
-  If \code{bw\_method\ =\ "cv"}, the user may change the numerical tolerance in the optimization of \(CV(b)\). See Table \ref{tab:arguments-lg-main} for the necessary argument names.
+  If \code{bw\_method\ =\ "cv"}, the user may change the numerical tolerance in the optimization of \(CV\left(b\right)\). See Table \ref{tab:arguments-lg-main} for the necessary argument names.
 \end{enumerate}
@@
     \centering
-    \begin{subfigure}[t]{0.48\textwidth}
+    \begin{minipage}[t]{0.48\textwidth}
       \includegraphics[width=\textwidth]{loccor-example}
-      \caption{The estimated pairwise LGC for the stock \newline data along the diagonal indicating that the pairs \\ of stocks are more strongly dependent in the lower \\ part of the distribution than the upper part.}
+      \subcaption{The estimated pairwise LGC for the stock \newline data along the diagonal indicating that the pairs \\ of stocks are more strongly dependent in the lower \\ part of the distribution than the upper part.}
       \label{fig:lgc-plot1}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.48\textwidth}
+    \end{minipage}
+    \begin{minipage}[t]{0.48\textwidth}
       \includegraphics[width=\textwidth]{density-example}
-      \caption{The corresponding joint density estimate of the four return series, plotted along the diagonal in $\mathbb{R}^4$.}
+      \subcaption{The corresponding joint density estimate of the four return series, plotted along the diagonal in $\mathbb{R}^4$.}
       \label{fig:lgc-plot2}
-    \end{subfigure}
-    \caption{Important output from the \code{dlg()-function}.}
+    \end{minipage}
+    \caption{Local correlations and density estimates calculated using the \code{dlg()-function}.}
     \label{fig:lgc-plot}
@@
 The last line of code creates a list containing a number of elements.
-The two most important are \code{\$loc\_cor}, which is a matrix of local correlations having one row per grid point and one column per \emph{pair of variables} (the columns correspond to the rows in \code{density\_object\$pairs}), as well as \code{\$f\_est}, which is a vector containing the estimate \(\widehat f_X(x)\) of the joint density \(f_X(x)\) in the grid points specified in \code{x0}.
+The two most important are \code{\$loc\_cor}, which is a matrix of local correlations having one row per grid point and one column per \emph{pair of variables} (the columns correspond to the rows in \code{density\_object\$pairs}), as well as \code{\$f\_est}, which is a vector containing the estimate \(\widehat f_X\left(x\right)\) of the joint density \(f_X\left(x\right)\) in the grid points specified in \code{x0}.
 The estimated local correlations for this example is plotted in Figure \ref{fig:lgc-plot1}, and the corresponding density estimate is plotted (along the diagonal \(x_1=x_2=x_3=x_4=x\)) in Figure \ref{fig:lgc-plot2}.
@@
 \code{normalization} & The estimated density does not integrate to one by & \\ \code{{  \textunderscore}points} & construction.
-\code{dlg()} will generate the given number   & \\ & of normal variables, having  the same moments as   & \\ &  the data, approximate  $\int \widehat f_X(x)\,\textrm{d}x$ by a Monte Carlo  & \\ & integral, and then normalize the density estimate  & \\ & accordingly & \\ \code{NULL} \\
+\code{dlg()} will generate the given number   & \\ & of normal variables, having  the same moments as   & \\ &  the data, approximate  $\int \widehat f_X\left(x\right)\,\textrm{d}x$ by a Monte Carlo  & \\ & integral, and then normalize the density estimate  & \\ & accordingly & \\ \code{NULL} \\
 \code{bootstrap} & Calculate bootstrapped confidence intervals instead& \\ &  of asymptotic expressions & \code{FALSE} \\
@@
 \item
-  A test for serial independence between \(X_t\) and \(X_{t-k}\) within a time series \(\{X_t\}\), cf. \citet{laca:tjos:2017}. 
+  A test for serial independence between \(X_t\) and \(X_{t-k}\) within a time series \(\left\{X_t\right\}\), cf. \citet{laca:tjos:2017}. 
 \item
-  A test for serial cross-dependence between \(X_t\) and \(Y_{t-k}\) for two time series \(\{X_t\}\) and \(\{Y_t\}\), cf. \citet{laca:tjos:2018}.
+  A test for serial cross-dependence between \(X_t\) and \(Y_{t-k}\) for two time series \(\left\{X_t\right\}\) and \(\left\{Y_t\right\}\), cf. \citet{laca:tjos:2018}.
 \end{enumerate}
@@
 
-The only difference when testing for serial independence within a time series \(\{X_t\}\) is to create a two-column data set consisting of
+The only difference when testing for serial independence within a time series \(\left\{X_t\right\}\) is to create a two-column data set consisting of
 \(X_t\) and \(X_{t-k}\). For example, if we wish to perform this test for \(k=1\) for one of the variables in the stock-exchange series,
@@
 
-Assume that we observe two financial time series \(\{X_t\}\) and \(\{Y_t\}\) at times \(t = 1,\ldots,T\), and that some crisis occurs at time \(T^* < T\).
-As described in Section \ref{chap:background}, \citet{stov:tjos:huft:2014} measure the local correlation between \(\{X_t\}\) and \(\{Y_t\}\) \emph{before} and \emph{after} \(T^*\), and take significant differences between these measurements as evidence against the null hypothesis of no linkage, or contagion, between the time series.
+Assume that we observe two financial time series \(\left\{X_t\right\}\) and \(\left\{Y_t\right\}\) at times \(t = 1,\ldots,T\), and that some crisis occurs at time \(T^* < T\).
+As described in Section \ref{chap:background}, \citet{stov:tjos:huft:2014} measure the local correlation between \(\left\{X_t\right\}\) and \(\left\{Y_t\right\}\) \emph{before} and \emph{after} \(T^*\), and take significant differences between these measurements as evidence against the null hypothesis of no linkage, or contagion, between the time series.
 In order to implement this test using the \CRANpkg{lg}-package, one must create \emph{two} lg-objects: one for the observations covering the non crisis period and one covering the crisis period.
@@
 
-Consider finally the work by \citet{otne:tjos:2019} who take the off-diagonal element in the local correlation matrix corresponding to the local conditional covariance matrix \eqref{eq:cond-sig} or \eqref{eq:cond-sig-trans} as a local measure of conditional dependence between two stochastic variables \(X^{(1)}\) and \(X^{(2)}\) given the stochastic vector \(X^{(3)}\).
+Consider finally the work by \citet{otne:tjos:2019} who take the off-diagonal element in the local correlation matrix corresponding to the local conditional covariance matrix \eqref{eq:cond-sig} or \eqref{eq:cond-sig-trans} as a local measure of conditional dependence between two stochastic variables \(X^{\left(1\right)}\) and \(X^{\left(2\right)}\) given the stochastic vector \(X^{\left(3\right)}\).
 Furthermore, in the case with data having been transformed to marginal standard normality, they take the statistic
 \begin{equation}
-T^{CI}_{n, b} = \int h(\widehat\alpha(z)) \, \textrm{d}F_n(z)
+T^{CI}_{n, b} = \int h\left(\widehat\alpha\left(z\right)\right) \, \textrm{d}F_n\left(z\right)
 \label{eq:condintstat}
@@
 The \CRANpkg{lg}-package provides two key functions in this framework.
-The first, \code{partial\_cor()}, calculates the local partial correlations as well as their estimated standard deviations on a specified grid in the \((x^{(1)}, x^{(2)}, x^{(3)})\)-space, and is essentially a wrapper function for the \code{clg()}-function presented in Section \ref{chap:density}.
+The first, \code{partial\_cor()}, calculates the local partial correlations as well as their estimated standard deviations on a specified grid in the \(\left(x^{\left(1\right)}, x^{\left(2\right)}, x^{\left(3\right)}\right)\)-space, and is essentially a wrapper function for the \code{clg()}-function presented in Section \ref{chap:density}.
 See Table \ref{tab:arguments-partial-cor} for details.
@@
 The critical values of this test are calculated using the bootstrap under the null hypothesis by independently resampling replicates from
-the conditional density estimates \(\widehat f_{X^{(1)|X^{(3)}}}(x^{(1)}|x^{(3)})\) and \(\widehat f_{X^{(2)|X^{(3)}}}(x^{(2)}|x^{(3)})\), as obtained by the \code{clg()}-function, using an approximated accept-reject algorithm.
-In order to avoid excessive optimization of the local likelihood function \eqref{eq:loclik}, we estimate \(f_{X^{(1)|X^{(3)}}}\) and \(f_{X^{(2)|X^{(3)}}}\) on the univariate regular grids \(x_0^{(1)}\) and \(x_0^{(2)}\) respectively (while keeping \(x^{(3)}\) fixed at the observed values of \(X^{(3)}\)), and produce interpolating functions \(\tilde f_{X^{(1)|X^{(3)}}}\) and \(\tilde f_{X^{(2)|X^{(3)}}}\) using cubic splines.
+the conditional density estimates \(\widehat f_{X^{\left(1\right)}|X^{\left(3\right)}}\left(x^{\left(1\right)}|x^{\left(3\right)}\right)\) and \(\widehat f_{X^{\left(2\right)|X^{\left(3\right)}}}\left(x^{\left(2\right)}|x^{\left(3\right)}\right)\), as obtained by the \code{clg()}-function, using an approximated accept-reject algorithm.
+In order to avoid excessive optimization of the local likelihood function \eqref{eq:loclik}, we estimate \(f_{X^{\left(1\right)|X^{\left(3\right)}}}\) and \(f_{X^{\left(2\right)|X^{\left(3\right)}}}\) on the univariate regular grids \(x_0^{\left(1\right)}\) and \(x_0^{\left(2\right)}\) respectively (while keeping \(x^{\left(3\right)}\) fixed at the observed values of \(X^{\left(3\right)}\)), and produce interpolating functions \(\tilde f_{X^{\left(1\right)}|X^{\left(3\right)}}\) and \(\tilde f_{X^{\left(2\right)}|X^{\left(3\right)}}\) using cubic splines.
 It is much less computationally intensive to generate replicates from \(\tilde f\) than directly from \(\widehat f\).
@@
 Some of the options are demonstrated in the code chunk below, in which we, for example, superimpose the observations (by setting \code{plot\_obs\ =\ TRUE)} as well as preventing the the estimated local correlations from being plotted in areas without data.
-The latter option is available through the argument \code{plot\_thres}, which works by calculating a bivariate kernel density estimate \(\tilde f(x_1, x_2)\) for the pair of  variables in question, and only allowing \(\widehat \rho(x_1,x_2)\) to be plotted if \(\tilde f(x_1,x_2)/\max\tilde f(\cdot) >\) \code{plot\_thres}.
+The latter option is available through the argument \code{plot\_thres}, which works by calculating a bivariate kernel density estimate \(\tilde f\left(x_1, x_2\right)\) for the pair of  variables in question, and only allowing \(\widehat \rho\left(x_1,x_2\right)\) to be plotted if \(\tilde f\left(x_1,x_2\right)/\max\tilde f\left(\cdot\right) >\) \code{plot\_thres}.
 Adding layers to a dependence map using the ordinary \CRANpkg{ggplot2} syntax works as well, which we demonstrate in Figure \ref{fig:graphics2} by changing the \CRANpkg{ggplot2} theme.
@@
 
-The author gratefully acknowledges the constructive comments made by an anonymous referee.
+The author gratefully acknowledges the constructive comments made by two anonymous referees.
 
@@
   Norway\\
+  ORCID: 0000-0002-6004-0237\\
   \email{hakon.otneim@nhh.no}}

### 2021-09-10 otneim.tex
Submission to the R Journal after accept

@@
 
-\citet{tjos:huft:2013} propose the \emph{local Gaussian correlation} (LGC) as a new measure of statistical dependence between two stochastic variables $X_1$ and $X_2$, which has the following important property yet unrivaled in the literature: It can separate between positive and negative nonlinear dependence while still reducing to the ordinary Pearson correlation coefficient if $X_1$ and $X_2$ are jointly normally distributed.
-The R-package \CRANpkg{localgauss} \citep{bere:klep:tjos:2014} provides two important functions in this context; one that calculates the sample LGC based on observed values of $\left(X_1,X_2\right)$, and one that uses the estimated LGC to perform a local test of independence between $X_1$ and $X_2$ as described in detail by \citet{bere:tjos:2014}.
+\cite{tjos:huft:2013} propose the \emph{local Gaussian correlation} (LGC) as a new measure of statistical dependence between two stochastic variables $X_1$ and $X_2$, which has the following important property yet unrivaled in the literature: It can separate between positive and negative nonlinear dependence while still reducing to the ordinary Pearson correlation coefficient if $X_1$ and $X_2$ are jointly normally distributed.
+The R-package \CRANpkg{localgauss} \citep{bere:klep:tjos:2014} provides two important functions in this context; one that calculates the sample LGC based on observed values of $\left(X_1,X_2\right)$, and one that uses the estimated LGC to perform a local test of independence between $X_1$ and $X_2$ as described in detail by \cite{bere:tjos:2014}.
 
 We have lately seen a number of new applications of the LGC that the \CRANpkg{localgauss}-package does not support, however.
-\citet{stov:tjos:huft:2014} use the LGC to test for financial contagion across markets during crises.
-\citet{otne:tjos:2017} present a procedure for estimating multivariate density functions via the LGC, which \citet{otne:tjos:2018} modify in order to compute estimates of conditional density functions.
-\citet{laca:tjos:2017} present a test for serial independence within a time series, which \citet{laca:tjos:2018} extend in order to include a test for cross-correlation between two time series.
-Finally, \citet{otne:tjos:2019} develop the local Gaussian \emph{partial} correlation (LGPC) as a measure of conditional dependence, and a corresponding test for conditional independence.
+\cite{stov:tjos:huft:2014} use the LGC to test for financial contagion across markets during crises.
+\cite{otne:tjos:2017} present a procedure for estimating multivariate density functions via the LGC, which \cite{otne:tjos:2018} modify in order to compute estimates of conditional density functions.
+\cite{laca:tjos:2017} present a test for serial independence within a time series, which \cite{laca:tjos:2018} extend in order to include a test for cross-correlation between two time series.
+Finally, \cite{otne:tjos:2019} develop the local Gaussian \emph{partial} correlation (LGPC) as a measure of conditional dependence, and a corresponding test for conditional independence.
 
 This paper describes the \CRANpkg{lg}-package \citep{otne:2019}, which provides a unified framework to implement all these methods, as well as a tool for visualizing the LGC and LGPC as dependence maps.
-\citet{jord:tjos:2017} use the LGC in spectral analysis of time series, but those methods have their own computational ecosystem in the \pkg{localgaussSpec}-package, \citet{jord:2018}.
+\cite{jord:tjos:2017} use the LGC in spectral analysis of time series, but those methods have their own computational ecosystem in the \pkg{localgaussSpec}-package, \cite{jord:2018}.
 
@@
 where $K$ is a symmetric density function (the kernel) and $b$ is a tuning parameter (the bandwidth) that controls the smoothness of the estimate $\widehat f_X\left(\cdot\right)$.
-See \citet{silv:1986} for an introduction to this topic.
+See \cite{silv:1986} for an introduction to this topic.
 There is also a massive statistical literature on density estimation containing extensions and improvements to the classical methods to be used in various practical situations.
 
-\citet{hjor:jone:1996} provide one such idea.
+\cite{hjor:jone:1996} provide one such idea.
 They consider a parametric family $F_{\theta}$, but instead of searching for a single parameter value $\theta_0$, for which $f_X\left(x\right) = f\left(x;\theta_0\right)$ (or approximately so), they rather assert that different members of $F_{\theta}$ may approximate $f_X$ locally in different parts of its domain.
@@
 The second term in the local likelihood function is a penalty that ensures that the estimated density $\widehat f_X\left(x\right) = f\left(x; \widehat\theta\left(x\right)\right)$ converges correctly to the true density function $f_X\left(x\right)$ as the sample size $n$ increases to infinity, and the bandwidth $b$ decreases towards zero.
-See \citet{hjor:jone:1996} for a detailed discussion about this construction.
+See \cite{hjor:jone:1996} for a detailed discussion about this construction.
 
-\citet{tjos:huft:2013} consider the bivariate case $X = \left(X_1, X_2\right)$ and take $F_{\theta}$ to be the family of bivariate normal distributions consisting of densities on the form
+\cite{tjos:huft:2013} consider the bivariate case $X = \left(X_1, X_2\right)$ and take $F_{\theta}$ to be the family of bivariate normal distributions consisting of densities on the form
 
@@
 In Figure \ref{fig:initial-example} we have plotted the estimated LGC for two bivariate data sets on a grid; 1000 simulated observations from a binormal distribution having correlation equal to 0.5, and the daily return on the CAC40 and FTSE100 stock indices on 1000 consecutive trading days starting on May 5th 2014 \citep{datastream}.
-In the first panel, we see that the estimated local correlation coincides with the global correlation, except for the estimation error which is comparable to the uncertainty observed in other non-parametric estimation methods such as the kernel density estimator (see, for instance, \citet{otne:tjos:2017} for a formal asymptotic analysis of relevant convergence rates).
+In the first panel, we see that the estimated local correlation coincides with the global correlation, except for the estimation error which is comparable to the uncertainty observed in other non-parametric estimation methods such as the kernel density estimator (see, for instance, \cite{otne:tjos:2017} for a formal asymptotic analysis of relevant convergence rates).
 In the second panel we see clearly that the local correlation, and thus the dependence, is stronger in the lower left and upper right regions of the distribution than in the central parts.
@@
 Taking the LGC as a measure of dependence opens up for a number of possibilities to construct statistical tests.
-\citet{bere:tjos:2014} show that $\rho\left(x\right) \equiv 0$ implies that $X_1$ and $X_2$ are independent.
+\cite{bere:tjos:2014} show that $\rho\left(x\right) \equiv 0$ implies that $X_1$ and $X_2$ are independent.
 They show further that independence between $X_1$ and $X_2$ implies $\rho\left(x\right) \equiv 0$ if the population values of the local mean and standard deviation functions satisfy the following conditions: $\mu_i\left(x_1, x_2\right) = \mu_i\left(x_i\right)$ and $\sigma_i\left(x_1, x_2\right) = \sigma_i\left(x_i\right)$ for $i=1,2$.
@@
 It follows then that departures from $\widehat\rho\left(x\right) \equiv 0$ may be taken as evidence against the hypothesis that $X_1$ and $X_2$ are statistically independent.
-\citet{bere:tjos:2014} formalize this notion by testing whether $\rho\left(x\right) \equiv 0$ for all $x\in S \subset \mathbb{R}^2$ using the test statistic
+\cite{bere:tjos:2014} formalize this notion by testing whether $\rho\left(x\right) \equiv 0$ for all $x\in S \subset \mathbb{R}^2$ using the test statistic
 
@@
 The autocorrelation function (ACF) $\rho_k = \rho\left(X_t, X_{t-k}\right)$ is a well known concept for describing the serial dependence in the time series, but the ACF is, again, only capable to completely capture \emph{linear} serial dependence.
-\citet{laca:tjos:2017} seek to remedy this by rather calculating the the local correlation between $X_t$ and $X_{t-k}$.
+\cite{laca:tjos:2017} seek to remedy this by rather calculating the the local correlation between $X_t$ and $X_{t-k}$.
 This leads to a test for serial independence in the natural way.
@@
 
-\citet{laca:tjos:2018} extend this problem to test for serial cross-dependence between two time series $\left\{X_t, Y_t\right\}$ by measuring the LGC between $X_t$ and $Y_{t-k}$.
+\cite{laca:tjos:2018} extend this problem to test for serial cross-dependence between two time series $\left\{X_t, Y_t\right\}$ by measuring the LGC between $X_t$ and $Y_{t-k}$.
 Departures from $\widehat\rho\left(x,y\right) \equiv 0$ is again taken as evidence against independence, and the test statistic (\ref{eq:test-statistic1}) provides an aggregate measure of this discrepancy in the specified region $S$.
 In this case, however, we can not obtain replicates of the test statistic under the null hypothesis by simple permutations of the data.
-\citet{laca:tjos:2018} suggest rather two block bootstrap procedures to this end, using fixed and random block sizes respectively.
+\cite{laca:tjos:2018} suggest rather two block bootstrap procedures to this end, using fixed and random block sizes respectively.
 The tests for serial dependence, and serial cross-dependence are both implemented in the \CRANpkg{lg}-package, as we demonstrate in Section \ref{chap:indtests}.
 
-We find another application of the local Gaussian approximation in the work by \citet{stov:tjos:huft:2014}, who measure and test for financial contagion.
+We find another application of the local Gaussian approximation in the work by \cite{stov:tjos:huft:2014}, who measure and test for financial contagion.
 They define contagion by "a significant increase in cross-market linkages after a shock to one country" \citep[p.
@@
 where $w\left(\cdot, \cdot\right)$ is a weight function that serve the same purpose as the integration area $S$ in (\ref{eq:test-statistic1}).
-In this case, \citet{stov:tjos:huft:2014} show that a standard bootstrap will suffice in order to produce approximate replicates of $T_{n,b}^D$ under the null hypothesis of no financial contagion, and we demonstrate the implementation of this test using the \CRANpkg{lg}-package in Section \ref{chap:cont}.
+In this case, \cite{stov:tjos:huft:2014} show that a standard bootstrap will suffice in order to produce approximate replicates of $T_{n,b}^D$ under the null hypothesis of no financial contagion, and we demonstrate the implementation of this test using the \CRANpkg{lg}-package in Section \ref{chap:cont}.
 
-Although the original work by \citet{hjor:jone:1996} provide a general framework for local likelihood density estimation using any $p$-variate parametric family as the local family, it is evident that the method may struggle in multivariate applications much in the same way as the kernel density estimator does.
-This is a consequence of the curse of dimensionality, the effect of which is sought remedied by an algorithm provided by \citet{otne:tjos:2017}.
+Although the original work by \cite{hjor:jone:1996} provide a general framework for local likelihood density estimation using any $p$-variate parametric family as the local family, it is evident that the method may struggle in multivariate applications much in the same way as the kernel density estimator does.
+This is a consequence of the curse of dimensionality, the effect of which is sought remedied by an algorithm provided by \cite{otne:tjos:2017}.
 The idea is to fit the $p$-variate normal distribution $\psi\left(\mu, \Sigma\right)$ locally, where $\mu$ is the vector of $p$ expectations, and $\Sigma$ is the $p\times p$ covariance matrix (to which the correlation matrix $R$ corresponds), but under the following structural simplifications:
@@
 \end{align}
-\citet{otne:tjos:2017} estimate the local parameters above by first obtaining univariate marginal locally Gaussian fits (eqs.
+\cite{otne:tjos:2017} estimate the local parameters above by first obtaining univariate marginal locally Gaussian fits (eqs.
 \ref{eq:simp-mu} and \ref{eq:simp-sig}), and then pairwise bivariate locally Gaussian fits (eq.
@@
 In the second step the estimates $\widehat\mu_i\left(x_i\right)$, $\widehat\mu_j\left(x_j\right)$, $\widehat\sigma_i\left(x_i\right)$ and $\widehat\sigma_j\left(x_j\right)$ are kept fixed in the estimation of the pairwise local correlation.
-They argue further that the following transformation technique will produce better density estimates in many situations.
 The motivation for introducing the simplifications defined in equations \ref{eq:simp-mu}-\ref{eq:simp-rho} can be compared to the practical advantages of estimating additive regression models, where $\textrm{E}\left(Y\right) = f\left(x_1, \ldots, x_p\right) \stackrel{\textrm{def}}{=} f_1\left(x_1\right) + \cdots + f_p\left(x_p\right)$.
+\cite{otne:tjos:2017} argue further that the following transformation technique will produce better density estimates in many situations.
 
@@
 \end{equation}
-where $\mu\left(z\right) = \left\{\mu_i\left(z\right)\right\}$ and $\sigma\left(z\right) = \left\{\sigma_i\left(z\right)\right\}$ for $i=1,\ldots,p$, and then substitute $f_Z$ for $\widehat f_Z$ in the following relation obtained by \citet{otne:tjos:2017} in order to estimate the unknown density $f_X$:
+where $\mu\left(z\right) = \left\{\mu_i\left(z\right)\right\}$ and $\sigma\left(z\right) = \left\{\sigma_i\left(z\right)\right\}$ for $i=1,\ldots,p$, and then substitute $f_Z$ for $\widehat f_Z$ in the following relation obtained by \cite{otne:tjos:2017} in order to estimate the unknown density $f_X$:
 \begin{equation}
@@
 see e.g.
-\citet[chapter 4]{john:wich:2007}.
-\citet{otne:tjos:2018} demonstrate that this property may be translated into a corresponding local argument without modification.
+\cite[chapter 4]{john:wich:2007}.
+\cite{otne:tjos:2018} demonstrate that this property may be translated into a corresponding local argument without modification.
 That is, if the joint density $f_X\left(\cdot\right)$ can be written on a locally Gaussian form
@@
 
-Finally, we refer to \citet{otne:tjos:2019} who take the local version of the conditional covariance matrix (\ref{eq:cond-sig}) (or (\ref{eq:cond-sig-trans}) in the transformed case) as a measure for conditional dependence, and thus as an instrument to test for conditional independence.
+Finally, we refer to \cite{otne:tjos:2019} who take the local version of the conditional covariance matrix (\ref{eq:cond-sig}) (or (\ref{eq:cond-sig-trans}) in the transformed case) as a measure for conditional dependence, and thus as an instrument to test for conditional independence.
 Consider the stochastic vector $X = \left(X^{\left(1\right)}, X^{\left(2\right)}, X^{\left(3\right)}\right)$, where $X^{\left(1\right)}$ and $X^{\left(2\right)}$ are scalars and $X^{\left(3\right)}$ may be a vector.
@@
 In this case, denote by $\alpha\left(z\right)$ the off-diagonal element in the $2\times2$ local correlation matrix $R^*\left(z\right)$ (which derives directly from $\Sigma^*\left(z\right)$ as given in (\ref{eq:cond-sig-trans})).
-If $X$ has a local Gaussian distribution, the conditional independence (\ref{eq:cond-ind}) is equivalent to $\alpha\left(z\right) \equiv 0$, and \citet{otne:tjos:2019} take departures from this relation as evidence against the hypothesis of conditional independence between $X^{\left(1\right)}$ and $X^{\left(2\right)}$ given $X^{\left(3\right)}$.
+If $X$ has a local Gaussian distribution, the conditional independence (\ref{eq:cond-ind}) is equivalent to $\alpha\left(z\right) \equiv 0$, and \cite{otne:tjos:2019} take departures from this relation as evidence against the hypothesis of conditional independence between $X^{\left(1\right)}$ and $X^{\left(2\right)}$ given $X^{\left(3\right)}$.
 The natural way to quantify this is the test functional
@@
 \end{equation}
-\citet{otne:tjos:2019} describe a bootstrap procedure for generating replicates of $T_{n,b}^{CI}$ under the null hypothesis.
+\cite{otne:tjos:2019} describe a bootstrap procedure for generating replicates of $T_{n,b}^{CI}$ under the null hypothesis.
 In Section \ref{chap:partial} we demonstrate how the \CRANpkg{lg}-package may be used to extract estimates of the local partial correlation and perform tests for conditional independence according to this scheme.
@@
 A full locally Gaussian fit for trivariate data.} If the number of variables $p$ is equal to $3$ \emph{and} we choose to transform the data to marginal standard normality (see the next sub-section), the transformed density $f_Z\left(\cdot\right)$ in (\ref{eq:transformed-density}) may be estimated by jointly estimating the three local correlations $\rho_{12}\left(z_1, z_2, z_3\right)$, $\rho_{13}\left(z_1, z_2, z_3\right)$ and $\rho_{23}\left(z_1, z_2, z_3\right)$.
-This estimation method was introduced recently by \citet{otne:tjos:2019} in order to increase power of their conditional independence test, but it can be used in any application described in this paper that consider trivariate data.
+This estimation method was introduced recently by \cite{otne:tjos:2019} in order to increase the power of their conditional independence test, but it can be used in any application described in this paper that consider trivariate data.
 To use this estimation method, create the lg-object by running
@@
 \[CV\left(b\right) = -\frac{1}{n}\sum_{i=1}^n \log f\Big(X_i; \widehat\theta^{\left(-i\right)}\left(X_i\right)\Big),\]
-where \(\widehat\theta^{\left(-i\right)}\left(x\right)\) is the parameter estimate obtained after deleting observation \(X_i\) from the data, is proportional to a quantity that estimates the Kullback-Leibler distance between \(f\left(\cdot, \widehat\theta\left(\cdot\right)\right)\) and the true density \(f\left(\cdot\right)\), see \citet{bere:tjos:2014}.
+where \(\widehat\theta^{\left(-i\right)}\left(x\right)\) is the parameter estimate obtained after deleting observation \(X_i\) from the data, is proportional to a quantity that estimates the Kullback-Leibler distance between \(f\left(\cdot, \widehat\theta\left(\cdot\right)\right)\) and the true density \(f\left(\cdot\right)\), see \cite{bere:tjos:2014}.
 The cross-validated bandwidth \(b_{CV}\) is hence given by \[b_{CV} = \arg\min_b \,\, CV\left(b\right).\] If we, for example, wish to use the simplified estimation procedure on the transformed data, we need bandwidths for the marginal estimates of the local means and standard deviations, as well as a \(2\times2\) diagonal bandwidth matrix for each pair of variables.
@@
 
-\citet{otne:tjos:2017} show that the simplified version of the local Gaussian fit have the same convergence rates as the corresponding nonparametric kernel density estimator, for which \citet{silv:1986} derives the plug-in formula \(b = 1.08\cdot sd\left(x\right) \cdot n^{-1/5}\).
+\cite{otne:tjos:2017} show that the simplified version of the local Gaussian fit have the same convergence rates as the corresponding nonparametric kernel density estimator, for which \cite{silv:1986} derives the plug-in formula \(b = 1.08\cdot sd\left(x\right) \cdot n^{-1/5}\).
 By
@@
 
-\section{Statistical inference using the \texttt{lg}-package}
+\section{Statistical inference using the \CRANpkg{lg}-package}
 \label{chap:inference}
@@
 
-We start by introducing a basic function for estimating the LGC on a grid as described by \citet{otne:tjos:2017}, and thus also a probability density estimate.
+We start by introducing a basic function for estimating the LGC on a grid as described by \cite{otne:tjos:2017}, and thus also a probability density estimate.
 We create a grid, \code{x0}, having the same number of columns as the data in the code below.
@@
 
-Note that the configuration \code{transform\_to\_marginal\_normality\ =\ TRUE} and \code{est\_method\ =\ 5par} in the bivariate case coincides with the situation considered by \citet{tjos:huft:2013}.
+Note that the configuration \code{transform\_to\_marginal\_normality\ =\ TRUE} and \code{est\_method\ =\ 5par} in the bivariate case coincides with the situation considered by \cite{tjos:huft:2013}.
 In that case, \code{dlg()} serves as a wrapper for the function \code{localgauss()} in the \CRANpkg{localgauss}-package \citep{bere:klep:tjos:2014}.
 
-Obtaining the estimate of a conditional density using the \citet{otne:tjos:2018} algorithm described in Section \ref{chap:background} is very similar, but one must take particular care of the \emph{ordering} of the variables in the data set.
+Obtaining the estimate of a conditional density using the \cite{otne:tjos:2018} algorithm described in Section \ref{chap:background} is very similar, but one must take particular care of the \emph{ordering} of the variables in the data set.
 The estimation function, \code{clg()}, will \emph{always} assume that the free variables come first, and that the conditioning variables come last.
@@
 \item
-  A test for independence between the stochastic variables \(X_1\) and \(X_2\) based on iid data, cf. \citet{bere:tjos:2014}.
+  A test for independence between the stochastic variables \(X_1\) and \(X_2\) based on iid data, cf. \cite{bere:tjos:2014}.
 \item
-  A test for serial independence between \(X_t\) and \(X_{t-k}\) within a time series \(\left\{X_t\right\}\), cf. \citet{laca:tjos:2017}. 
+  A test for serial independence between \(X_t\) and \(X_{t-k}\) within a time series \(\left\{X_t\right\}\), cf. \cite{laca:tjos:2017}. 
 \item
-  A test for serial cross-dependence between \(X_t\) and \(Y_{t-k}\) for two time series \(\left\{X_t\right\}\) and \(\left\{Y_t\right\}\), cf. \citet{laca:tjos:2018}.
+  A test for serial cross-dependence between \(X_t\) and \(Y_{t-k}\) for two time series \(\left\{X_t\right\}\) and \(\left\{Y_t\right\}\), cf. \cite{laca:tjos:2018}.
 \end{enumerate}
@@
 In this case, \(X_1\) and \(X_2\) are uncorrelated, but obviously strongly dependent.
-\citet{bere:tjos:2014} considers mainly the full bivariate fit to the raw data, which we easily encode into the lg-object as before.
+\cite{bere:tjos:2014} considers mainly the full bivariate fit to the raw data, which we easily encode into the lg-object as before.
 The implementation of the test using 100 bootstrap replicates is shown below.
@@
 This is implemented in the \code{ind\_test()}-function by setting the \code{bootstrap\_type}-argument to \code{"plain"}, which is the default option.
-When testing for serial cross-dependence we need to use a block-bootstrap procedure, and \citet{laca:tjos:2018} consider two options here: The block bootstrap with either fixed \citep{kuns:1989} or random \citep{poli:roma:1994} block sizes, and this choice is specified by choosing \code{bootstrap\_type\ =\ "block"} or \code{bootstrap\_type\ =\ "stationary"}, respectively, in the call to \code{ind\_test()}.
-\citet{laca:tjos:2018} do not report significant differences in test performance using the different bootstrap types.
+When testing for serial cross-dependence we need to use a block-bootstrap procedure, and \cite{laca:tjos:2018} consider two options here: The block bootstrap with either fixed \citep{kuns:1989} or random \citep{poli:roma:1994} block sizes, and this choice is specified by choosing \code{bootstrap\_type\ =\ "block"} or \code{bootstrap\_type\ =\ "stationary"}, respectively, in the call to \code{ind\_test()}.
+\cite{laca:tjos:2018} do not report significant differences in test performance using the different bootstrap types.
 
@@
 Assume that we observe two financial time series \(\left\{X_t\right\}\) and \(\left\{Y_t\right\}\) at times \(t = 1,\ldots,T\), and that some crisis occurs at time \(T^* < T\).
-As described in Section \ref{chap:background}, \citet{stov:tjos:huft:2014} measure the local correlation between \(\left\{X_t\right\}\) and \(\left\{Y_t\right\}\) \emph{before} and \emph{after} \(T^*\), and take significant differences between these measurements as evidence against the null hypothesis of no linkage, or contagion, between the time series.
+As described in Section \ref{chap:background}, \cite{stov:tjos:huft:2014} measure the local correlation between \(\left\{X_t\right\}\) and \(\left\{Y_t\right\}\) \emph{before} and \emph{after} \(T^*\), and take significant differences between these measurements as evidence against the null hypothesis of no linkage, or contagion, between the time series.
 In order to implement this test using the \CRANpkg{lg}-package, one must create \emph{two} lg-objects: one for the observations covering the non crisis period and one covering the crisis period.
@@
 
-Let us illustrate the implementation of this test by looking at the same financial returns data that we have used in preceding sections, but this time we will, in the spirit of \citet{stov:tjos:huft:2014}, concentrate on \emph{GARCH(1,1)-filtrated daily returns on the S\&P500 and FTSE100 indices from 2 January 1985 to 29 April 1987} in order to test for financial contagion between the US and UK stock markets following the global stock market crash of 19 October 1987 (``Black Monday'').
+Let us illustrate the implementation of this test by looking at the same financial returns data that we have used in preceding sections, but this time we will, in the spirit of \cite{stov:tjos:huft:2014}, concentrate on \emph{GARCH(1,1)-filtrated daily returns on the S\&P500 and FTSE100 indices from 2 January 1985 to 29 April 1987} in order to test for financial contagion between the US and UK stock markets following the global stock market crash of 19 October 1987 (``Black Monday'').
 Assume that these observations are loaded into the R workspace as the \(n_1 \times 2\) data frame \code{x\_nc} containing the observations covering the \(n_1 = 728\) days preceding the crisis , and the \(n_2 \times 2\) data frame \code{x\_c} containing the observations covering the \(n_2 = 140\) consecutive trading days starting on \emph{Black Monday} (see the online code supplement for details concerning the GARCH-filtration and data processing).
-In the code below we construct one lg-object for each of these data frames with configuration matching the setup used by \citet{stov:tjos:huft:2014}, and perform the test by means of the \code{cont\_test()}-function.
+In the code below we construct one lg-object for each of these data frames with configuration matching the setup used by \cite{stov:tjos:huft:2014}, and perform the test by means of the \code{cont\_test()}-function.
 
@@
 
-Consider finally the work by \citet{otne:tjos:2019} who take the off-diagonal element in the local correlation matrix corresponding to the local conditional covariance matrix \eqref{eq:cond-sig} or \eqref{eq:cond-sig-trans} as a local measure of conditional dependence between two stochastic variables \(X^{\left(1\right)}\) and \(X^{\left(2\right)}\) given the stochastic vector \(X^{\left(3\right)}\).
+Consider finally the work by \cite{otne:tjos:2019} who take the off-diagonal element in the local correlation matrix corresponding to the local conditional covariance matrix \eqref{eq:cond-sig} or \eqref{eq:cond-sig-trans} as a local measure of conditional dependence between two stochastic variables \(X^{\left(1\right)}\) and \(X^{\left(2\right)}\) given the stochastic vector \(X^{\left(3\right)}\).
 Furthermore, in the case with data having been transformed to marginal standard normality, they take the statistic
@@
 We conclude this article by describing the \code{corplot()} function for drawing dependence maps such as those displayed in Figure \ref{fig:initial-example}.
-\citet{bere:klep:tjos:2014} report on such capabilities in the \CRANpkg{localgauss}-package, but the possibility of creating dependence maps was unfortunately removed from \CRANpkg{localgauss} in the latest version 0.4.0 due to incompatibilities with the \CRANpkg{ggplot2} \citep{wick:2016} plotting engine.
+\cite{bere:klep:tjos:2014} report on such capabilities in the \CRANpkg{localgauss}-package, but the possibility of creating dependence maps was unfortunately removed from \CRANpkg{localgauss} in the latest version 0.4.0 due to incompatibilities with the \CRANpkg{ggplot2} \citep{wick:2016} plotting engine.
 We make up for this loss by providing \code{corplot()}, a function that plots the estimated local correlations as provided by \code{dlg()}, or the estimated local \emph{partial} correlations as provided by \code{partial\_cor()}.

### 2021-12-14 otneim.tex
Copy editing comments implemented

--- 2021-09-09 otneim.tex
@@
 The package \CRANpkg{lg} for the R programming language provides implementations of recent methodological advances on applications of the local Gaussian correlation.
-This includes estimation of the local Gaussian correlation itself, multivariate density estimation, conditional density estimation, various tests for independence and conditional independence, as well as a graphical module for creating dependence maps.
+This includes the estimation of the local Gaussian correlation itself, multivariate density estimation, conditional density estimation, various tests for independence and conditional independence, as well as a graphical module for creating dependence maps.
 This paper describes the \CRANpkg{lg} package, its principles, and its practical use.
@@
 
-We have lately seen a number of new applications of the LGC that the \CRANpkg{localgauss}-package does not support, however.
+We have lately seen a number of new applications of the LGC that the \CRANpkg{localgauss} package does not support, however.
 \citet{stov:tjos:huft:2014} use the LGC to test for financial contagion across markets during crises.
@@
 \citet{laca:tjos:2017} present a test for serial independence within a time series, which \citet{laca:tjos:2018} extend in order to include a test for cross-correlation between two time series.
-Finally, \citet{otne:tjos:2019} develop the local Gaussian \emph{partial} correlation (LGPC) as a measure of conditional dependence, and a corresponding test for conditional independence.
+Finally, \citet{otne:tjos:2019} develop the local Gaussian \emph{partial} correlation (LGPC) as a measure of conditional dependence and a corresponding test for conditional independence.
 
-This paper describes the \CRANpkg{lg}-package \citep{otne:2019}, which provides a unified framework to implement all these methods, as well as a tool for visualizing the LGC and LGPC as dependence maps.
-\citet{jord:tjos:2017} use the LGC in spectral analysis of time series, but those methods have their own computational ecosystem in the \pkg{localgaussSpec}-package, \citet{jord:2018}.
+This paper describes the \CRANpkg{lg} package \citep{otne:2019}, which provides a unified framework to implement all these methods, as well as a tool for visualizing the LGC and LGPC as dependence maps.
+\citet{jord:tjos:2017} use the LGC in spectral analysis of time series, but those methods have their own computational ecosystem in the \pkg{localgaussSpec} package \citep{jord:2018}.
 
-In Section \ref{chap:background} we provide a brief introduction to the LGC as well as the methods and applications referred to above.
-In Section \ref{chap:lgobject} we describe the core function in the \CRANpkg{lg}-package and move on to demonstrate the implementation of various applications in Section \ref{chap:inference}.
-We conclude this paper in Section \ref{chap:graphics} by demonstrating the graphical capabilities of the \CRANpkg{lg}-package.
+In Section \ref{chap:background}, we provide a brief introduction to the LGC as well as the methods and applications referred to above.
+In Section \ref{chap:lgobject}, we describe the core function in the \CRANpkg{lg} package and move on to demonstrate the implementation of various applications in Section \ref{chap:inference}.
+We conclude this paper in Section \ref{chap:graphics} by demonstrating the graphical capabilities of the \CRANpkg{lg} package.
 
@@
 One may, for example, make the assumption that the unknown density function has a particular parametric form, $f_X \in F_{\theta}$, where $F_{\theta} = \left\{f(x;\theta), \theta \in \Theta\right\}$ is a family of probability density functions indexed by some parameter $\theta$, and where $\Theta$ is the parameter space.
-Under this assumption we will typically produce an estimate of the parameter $\theta$, written $\widehat \theta$, using the maximum likelihood method.
+Under this assumption, we will typically produce an estimate of the parameter $\theta$, written $\widehat \theta$, using the maximum likelihood method.
 The estimated probability density function is then given as $\widehat f_X\left(x\right) = f\left(x; \widehat\theta\right)$.
 
-A different approach is to estimate $f_X\left(\cdot\right)$ without any prior parametric assumptions. The classical method for non-parametric density estimation is the kernel estimator 
+A different approach is to estimate $f_X\left(\cdot\right)$ without any prior parametric assumptions. The classical method for nonparametric density estimation is the kernel estimator 
 $$\widehat f_X\left(x\right) = \frac{1}{nb}\sum_{i=1}^n K\left(\frac{X_i - x}{b}\right),$$
@@
 \citet{hjor:jone:1996} provide one such idea.
-They consider a parametric family $F_{\theta}$, but instead of searching for a single parameter value $\theta_0$, for which $f_X\left(x\right) = f\left(x;\theta_0\right)$ (or approximately so), they rather assert that different members of $F_{\theta}$ may approximate $f_X$ locally in different parts of its domain.
+They consider a parametric family $F_{\theta}$, but instead of searching for a single parameter value $\theta_0$ for which $f_X\left(x\right) = f\left(x;\theta_0\right)$ (or approximately so), they rather assert that different members of $F_{\theta}$ may approximate $f_X$ locally in different parts of its domain.
 In other words, they seek to estimate a parameter \emph{function} $\theta_0\left(x\right)$ for which $f_X\left(x\right) = f\left(x;\theta_0\left(x\right)\right)$ (or approximately so), and do this by maximizing a \emph{local likelihood function} in each point $x$:
@@
 where, again, $K$ is a symmetric density function and $b$ is a bandwidth parameter that controls the smoothness of the estimate.
-The second term in the local likelihood function is a penalty that ensures that the estimated density $\widehat f_X\left(x\right) = f\left(x; \widehat\theta\left(x\right)\right)$ converges correctly to the true density function $f_X\left(x\right)$ as the sample size $n$ increases to infinity, and the bandwidth $b$ decreases towards zero.
+The second term in the local likelihood function is a penalty that ensures that the estimated density $\widehat f_X\left(x\right) = f\left(x; \widehat\theta\left(x\right)\right)$ converges correctly to the true density function $f_X\left(x\right)$ as the sample size $n$ increases to infinity and the bandwidth $b$ decreases towards zero.
 See \citet{hjor:jone:1996} for a detailed discussion about this construction.
@@
 $$\widehat\theta\left(x\right) = \left(\widehat\mu_1\left(x\right), \widehat\mu_2\left(x\right), \widehat\sigma_1\left(x\right), \widehat\sigma_2\left(x\right), \widehat\rho\left(x\right)\right),$$
-and take special interest in the estimated correlation function $\widehat\rho\left(x\right)$ (i.e.
-the LGC) because it serves as an attractive local measure of statistical dependence between $X_1$ and $X_2$.
+and take special interest in the estimated correlation function $\widehat\rho\left(x\right)$ (i.e., the LGC) because it serves as an attractive local measure of statistical dependence between $X_1$ and $X_2$.
 They show that the LGC reveals many types of nonlinear statistical dependence that are not captured by the ordinary (global) Pearson correlation coefficient.
 Furthermore, the LGC distinguishes between positive and negative dependence and reduces to the Pearson $\rho$ if $X_1$ and $X_2$ are jointly normal.
-We refer to \cite{tjos:huft:2013} for a detailed treatment of the theoretical foundations of the LGC as well as several examples, and rather present two simple illustrations at this point in order to demonstrate the concept.
+We refer to \cite{tjos:huft:2013} for a detailed treatment of the theoretical foundations of the LGC as well as several examples and rather present two simple illustrations at this point in order to demonstrate the concept.
 
@@
       \includegraphics[width=\textwidth]{returns-example}
-      \subcaption{Local Gaussian correlation between daily returns from the CAC40 and FTSE100 stock indices, based on 1000 consecutive trading days.}
+      \subcaption{Local Gaussian correlation between daily returns from the CAC40 and FTSE100 stock indices based on 1000 consecutive trading days.}
     \end{minipage}
@@
 
-In Figure \ref{fig:initial-example} we have plotted the estimated LGC for two bivariate data sets on a grid; 1000 simulated observations from a binormal distribution having correlation equal to 0.5, and the daily return on the CAC40 and FTSE100 stock indices on 1000 consecutive trading days starting on May 5th 2014 \citep{datastream}.
-In the first panel, we see that the estimated local correlation coincides with the global correlation, except for the estimation error which is comparable to the uncertainty observed in other non-parametric estimation methods such as the kernel density estimator (see, for instance, \citet{otne:tjos:2017} for a formal asymptotic analysis of relevant convergence rates).
-In the second panel we see clearly that the local correlation, and thus the dependence, is stronger in the lower left and upper right regions of the distribution than in the central parts.
+In Figure \ref{fig:initial-example}, we have plotted the estimated LGC for two bivariate data sets on a grid; 1000 simulated observations from a binormal distribution having correlation equal to 0.5, and the daily return on the CAC40 and FTSE100 stock indices on 1000 consecutive trading days starting on May 5th 2014 \citep{datastream}.
+In the first panel, we see that the estimated local correlation coincides with the global correlation, except for the estimation error which is comparable to the uncertainty observed in other nonparametric estimation methods such as the kernel density estimator (see, for instance, \citet{otne:tjos:2017} for a formal asymptotic analysis of relevant convergence rates).
+In the second panel, we see clearly that the local correlation, and thus the dependence, is stronger in the lower left and upper right regions of the distribution than in the central parts.
 The phenomenon of local dependence is well known in the financial literature, and using the LGC it can be measured, interpreted, and visualized in a natural way.
@@
 
-One may obtain these particular estimates from the older \CRANpkg{localgauss}-package (as well as the \CRANpkg{lg}-package of course), but the plotting routine that was used to produce these figures is included in the \CRANpkg{lg}-package and will be described in more detail in Section \ref{chap:graphics}.
+One may obtain these particular estimates from the older \CRANpkg{localgauss} package (as well as the \CRANpkg{lg} package, of course), but the plotting routine that was used to produce these figures is included in the \CRANpkg{lg} package and will be described in more detail in Section \ref{chap:graphics}.
 
-Taking the LGC as a measure of dependence opens up for a number of possibilities to construct statistical tests.
+Taking the LGC as a measure of dependence opens up a number of possibilities to construct statistical tests.
 \citet{bere:tjos:2014} show that $\rho\left(x\right) \equiv 0$ implies that $X_1$ and $X_2$ are independent.
@@
 for some non-negative function $h$, for example $h\left(x\right) = x^2$ or $h\left(x\right) = |x|$.
-Critical values may be obtained by permutations of the data under the null hypothesis, and we demonstrate the implementation of this test using the \CRANpkg{lg}-package in Section \ref{chap:indtests}.
+Critical values may be obtained by permutations of the data under the null hypothesis, and we demonstrate the implementation of this test using the \CRANpkg{lg} package in Section \ref{chap:indtests}.
 
@@
 The autocorrelation function (ACF) $\rho_k = \rho\left(X_t, X_{t-k}\right)$ is a well known concept for describing the serial dependence in the time series, but the ACF is, again, only capable to completely capture \emph{linear} serial dependence.
-\citet{laca:tjos:2017} seek to remedy this by rather calculating the the local correlation between $X_t$ and $X_{t-k}$.
-This leads to a test for serial independence in the natural way.
+\citet{laca:tjos:2017} seek to remedy this by rather calculating the local correlation between $X_t$ and $X_{t-k}$.
+This leads to a test for serial independence in a natural way.
 In fact, this work is mainly a theoretical exercise in order to accommodate time series dependence.
@@
 \citet{laca:tjos:2018} extend this problem to test for serial cross-dependence between two time series $\left\{X_t, Y_t\right\}$ by measuring the LGC between $X_t$ and $Y_{t-k}$.
-Departures from $\widehat\rho\left(x,y\right) \equiv 0$ is again taken as evidence against independence, and the test statistic (\ref{eq:test-statistic1}) provides an aggregate measure of this discrepancy in the specified region $S$.
+Departures from $\widehat\rho\left(x,y\right) \equiv 0$ are, again, taken as evidence against independence, and the test statistic (\ref{eq:test-statistic1}) provides an aggregate measure of this discrepancy in the specified region $S$.
 In this case, however, we can not obtain replicates of the test statistic under the null hypothesis by simple permutations of the data.
-\citet{laca:tjos:2018} suggest rather two block bootstrap procedures to this end, using fixed and random block sizes respectively.
-The tests for serial dependence, and serial cross-dependence are both implemented in the \CRANpkg{lg}-package, as we demonstrate in Section \ref{chap:indtests}.
+\citet{laca:tjos:2018} suggest two block bootstrap procedures instead to this end, using fixed and random block sizes, respectively.
+The tests for serial dependence and serial cross-dependence are both implemented in the \CRANpkg{lg} package, as we demonstrate in Section \ref{chap:indtests}.
 
-We find another application of the local Gaussian approximation in the work by \citet{stov:tjos:huft:2014}, who measure and test for financial contagion.
-They define contagion by "a significant increase in cross-market linkages after a shock to one country" \citep[p.
+We find another application of the local Gaussian approximation in work by \citet{stov:tjos:huft:2014}, who measure and test for financial contagion.
+They define contagion as "a significant increase in cross-market linkages after a shock to one country" \citep[p.
 2223]{forb:rigo:2002} and employ the LGC to quantify this potential linkage.
-The authors estimate the LGC on a grid $\left\{x_1, x_2\right\}_k$, $k=1,\ldots,K$ along the diagonal $D=\left\{\left(x_1,x_2\right): x_1 = x_2\right\}$ \emph{before} and \emph{after} some critical event in the financial markets, denoted as the crisis (C) and the non-crisis (NC) periods respectively.
-They compare the two estimates using the following test statistic;
+The authors estimate the LGC on a grid $\left\{x_1, x_2\right\}_k$, $k=1,\ldots,K$ along the diagonal $D=\left\{\left(x_1,x_2\right): x_1 = x_2\right\}$ \emph{before} and \emph{after} some critical event in the financial markets, denoted as the crisis (C) and the non-crisis (NC) periods, respectively.
+They compare the two estimates using the following test statistic,
 $$T_{n,b}^D = \sum_{k=1}^K \left\{\widehat\rho_C\left(x_k, x_k\right) - \widehat\rho_{NC}\left(x_k, x_k\right) \right\}w\left(x_k,x_k\right),$$
-
 where $w\left(\cdot, \cdot\right)$ is a weight function that serve the same purpose as the integration area $S$ in (\ref{eq:test-statistic1}).
-In this case, \citet{stov:tjos:huft:2014} show that a standard bootstrap will suffice in order to produce approximate replicates of $T_{n,b}^D$ under the null hypothesis of no financial contagion, and we demonstrate the implementation of this test using the \CRANpkg{lg}-package in Section \ref{chap:cont}.
+In this case, \citet{stov:tjos:huft:2014} show that a standard bootstrap will suffice in order to produce approximate replicates of $T_{n,b}^D$ under the null hypothesis of no financial contagion, and we demonstrate the implementation of this test using the \CRANpkg{lg} package in Section \ref{chap:cont}.
 
@@
 \ref{eq:simp-rho}).
-In the second step the estimates $\widehat\mu_i\left(x_i\right)$, $\widehat\mu_j\left(x_j\right)$, $\widehat\sigma_i\left(x_i\right)$ and $\widehat\sigma_j\left(x_j\right)$ are kept fixed in the estimation of the pairwise local correlation.
+In the second step, the estimates $\widehat\mu_i\left(x_i\right)$, $\widehat\mu_j\left(x_j\right)$, $\widehat\sigma_i\left(x_i\right)$, and $\widehat\sigma_j\left(x_j\right)$ are kept fixed in the estimation of the pairwise local correlation.
 They argue further that the following transformation technique will produce better density estimates in many situations.
@@
 Denote by $F_i\left(x_i\right)$, $i = 1,\ldots, p$ the marginal distribution functions of the stochastic vector $X$, and by $\widehat F_i\left(x_i\right) = n^{-1}\sum_{i=1}^n 1\left(X_i \leq x_i\right)$ their empirical counterparts.
-They then estimate the density $f_Z\left(z\right)$ of the vector $Z = \left\{\Phi^{-1}\left(F_i\left(X_i\right)\right)\right\}_{i=1,\ldots,p}$, in practice approximated by
+They then estimate the density $f_Z\left(z\right)$ of the vector $Z = \left\{\Phi^{-1}\left(F_i\left(X_i\right)\right)\right\}_{i=1,\ldots,p}$.
+In practice it is approximated by
 
@@
 so that the only parameter functions left to estimate are the pairwise local Gaussian correlations $R\left(z\right) = \left\{\rho_{ij}\left(z_i, z_j\right)\right\}_{i<j}$.
-We use the notation $Z$, $z_i$ and $z_j$ to signify that the estimation is performed on the (approximate) standard normal scale, or $z$-scale for short.
+We use the notation $Z$, $z_i$, and $z_j$ to signify that the estimation is performed on the (approximate) standard normal scale or $z$-scale for short.
 We can then estimate the joint density $f_Z\left(z\right)$ of $Z$ as
@@
 \end{equation}
-where $\phi\left(\cdot\right)$ is the standard normal pdf. This estimator is implemented the \CRANpkg{lg}-package as demonstrated in Section \ref{chap:density}.
+where $\phi\left(\cdot\right)$ is the standard normal pdf. This estimator is implemented the \CRANpkg{lg} package as demonstrated in Section \ref{chap:density}.
 
-One particular feature enjoyed by the jointly normally distributed vector $X$ is that, for any partitioning $X = \left(X^{\left(1\right)}, X^{\left(2\right)}\right)$, the conditional distribution of $X^{\left(1\right)}|X^{\left(2\right)} = x^{\left(2\right)}$ is also normal.
+One particular feature enjoyed by the jointly normally distributed vector $X$ is that for any partitioning $X = \left(X^{\left(1\right)}, X^{\left(2\right)}\right)$, the conditional distribution of $X^{\left(1\right)}|X^{\left(2\right)} = x^{\left(2\right)}$ is also normal.
 In fact, if $X \sim \mathcal{N}\left(\mu,\Sigma\right)$, and $\mu$ and $\Sigma$ is partitioned according to $\left(X^{\left(1\right)}, X^{\left(2\right)}\right)$ as
@@
 where we, again, switch to $z$-notation in order to make it clear that these quantities are estimated on the standard normal $z$-scale.
-An estimator $\widehat f_{X^{\left(1\right)}|X^{\left(2\right)}}\left(\cdot|\cdot\right)$ of the conditional density $f_{X^{\left(1\right)}|X^{\left(2\right)}}\left(\cdot|\cdot\right)$ follows immediately from an expression corresponding to (\ref{eq:backtrans}), and the \CRANpkg{lg}-package provides functions for implementing this estimator in R.
+An estimator $\widehat f_{X^{\left(1\right)}|X^{\left(2\right)}}\left(\cdot|\cdot\right)$ of the conditional density $f_{X^{\left(1\right)}|X^{\left(2\right)}}\left(\cdot|\cdot\right)$ follows immediately from an expression corresponding to (\ref{eq:backtrans}), and the \CRANpkg{lg} package provides functions for implementing this estimator in R.
 We describe the implementation of this functionality in Section \ref{chap:density}.
@@
 \citet{otne:tjos:2019} describe a bootstrap procedure for generating replicates of $T_{n,b}^{CI}$ under the null hypothesis.
-In Section \ref{chap:partial} we demonstrate how the \CRANpkg{lg}-package may be used to extract estimates of the local partial correlation and perform tests for conditional independence according to this scheme.
+In Section \ref{chap:partial}, we demonstrate how the \CRANpkg{lg} package may be used to extract estimates of the local partial correlation and perform tests for conditional independence according to this scheme.
 
@@
 The local Gaussian correlation may be used to perform a number of statistical analyses, as is evident from the preceding section.
-The practitioner must first, however, make three quite specific modeling choices; namely (i) to choose an estimation method, i.e.
-the level of simplification in multivariate applications, (ii) to determine whether the data should be transformed towards marginal standard normality before estimating the LGC, and (iii) to choose a set of bandwidths, or at least a method for calculating bandwidths.
-The architecture of the \pkg{lg}-packages requires the user to make these choices \emph{before} endeavoring further into specific applications by imposing a strict, two step procedure:
+The practitioner must first, however, make three quite specific modeling choices; namely (i) to choose an estimation method, i.e., the level of simplification in multivariate applications, (ii) to determine whether the data should be transformed towards marginal standard normality before estimating the LGC, and (iii) to choose a set of bandwidths or at least a method for calculating bandwidths.
+The architecture of the \CRANpkg{lg} package requires the user to make these choices \emph{before} endeavoring further into specific applications by imposing a strict, two-step procedure:
 
@@
 
-In the following we assume that one has a \emph{data set} \code{x} loaded into the R workspace, which must be an $n\times p$ matrix (one column per variable, one row per observation), possibly including \code{NA}s which will be excluded from the analysis, or a data frame having the same dimensions.
-The fundamental syntax for creating an lg-object is \code{lg\_object <- lg\_main(x)}, and we will in this section explain how the modeling decisions (i)-(iii) can be encoded into the lg-object by using appropriate arguments in this function.
+In the following, we assume that one has a \emph{data set} \code{x} loaded into the R workspace, which must be an $n\times p$ matrix (one column per variable, one row per observation), possibly including \code{NA}s which will be excluded from the analysis, or a data frame having the same dimensions.
+The fundamental syntax for creating an lg-object is \code{lg\_object <- lg\_main(x)}, and we will, in this section, explain how the modeling decisions (i)-(iii) can be encoded into the lg-object by using appropriate arguments in this function.
 
@@
 
-Given a data set \code{x} having $n$ rows and $p \geq 2$ columns, the user must choose between four distinct estimation methods, and specify this choice by using the argument \code{est\_method} to the \code{lg\_main()}-function.
-We look at the built-in bivariate data set \code{faithful}, which records the waiting time between eruptions and the duration of the eruption for the Old Faithful geyser in the Yellowstone National Park, USA (see the help file in R for more details: \code{?faithful}), and load the \CRANpkg{lg}-package in order to demonstrate the implementation:
+Given a data set \code{x} having $n$ rows and $p \geq 2$ columns, the user must choose between four distinct estimation methods and specify this choice by using the argument \code{est\_method} to the \code{lg\_main()}-function.
+We look at the built-in bivariate data set \code{faithful}, which records the waiting time between eruptions and the duration of the eruption for the Old Faithful geyser in the Yellowstone National Park, USA (see the help file in R for more details: \code{?faithful}), and load the \CRANpkg{lg} package in order to demonstrate the implementation:
 
@@
 \textbf{1.
-A full locally Gaussian fit for bivariate data.} If $p = 2$, we may fit the bivariate normal $\psi\left(x, \theta\right)$ locally to $f\left(x\right)$, and by a "full local fit" we mean that we jointly estimate the five parameters
+A full locally Gaussian fit for bivariate data.} If $p = 2$, we may fit the bivariate normal $\psi\left(x, \theta\right)$ locally to $f\left(x\right)$, and by a "full local fit", we mean that we jointly estimate the five parameters
 $$\theta\left(x\right) = \big(\mu_1\left(x_1, x_2\right), \mu_2\left(x_1, x_2\right), \sigma_1\left(x_1, x_2\right), \sigma_2\left(x_1,x_2\right),\rho\left(x_1, x_2\right)\big)$$
@@
 \textbf{3.
-A simplified locally Gaussian fit for marginally standard normal data.} This estimation method is applicable for marginally standard normal data, or data that have been transformed to approximate marginal standard normality by e.g.
-the transformation (\ref{eq:trans}).
-In that case, we fix the marginal expectation functions and standard deviation functions to the constant values $0$ and $1$ respectively, and estimate only the pairwise local Gaussian correlations as in (\ref{eq:transformed-density}).
+A simplified locally Gaussian fit for marginally standard normal data.} This estimation method is applicable for marginally standard normal data, or data that have been transformed to approximate marginal standard normality by, e.g., the transformation (\ref{eq:trans}).
+In that case, we fix the marginal expectation functions and standard deviation functions to the constant values $0$ and $1$, respectively, and estimate only the pairwise local Gaussian correlations as in (\ref{eq:transformed-density}).
 To use this estimation method, create the lg-object by running
@@
 \textbf{4.
-A full locally Gaussian fit for trivariate data.} If the number of variables $p$ is equal to $3$ \emph{and} we choose to transform the data to marginal standard normality (see the next sub-section), the transformed density $f_Z\left(\cdot\right)$ in (\ref{eq:transformed-density}) may be estimated by jointly estimating the three local correlations $\rho_{12}\left(z_1, z_2, z_3\right)$, $\rho_{13}\left(z_1, z_2, z_3\right)$ and $\rho_{23}\left(z_1, z_2, z_3\right)$.
+A full locally Gaussian fit for trivariate data.} If the number of variables $p$ is equal to $3$, and we choose to transform the data to marginal standard normality (see the next sub-section), the transformed density $f_Z\left(\cdot\right)$ in (\ref{eq:transformed-density}) may be estimated by jointly estimating the three local correlations $\rho_{12}\left(z_1, z_2, z_3\right)$, $\rho_{13}\left(z_1, z_2, z_3\right)$, and $\rho_{23}\left(z_1, z_2, z_3\right)$.
 This estimation method was introduced recently by \citet{otne:tjos:2019} in order to increase power of their conditional independence test, but it can be used in any application described in this paper that consider trivariate data.
@@
 
-The resulting \code{lg\_object} now includes the element \code{transform\_to\_marginal\_normality} set according to the input, and if this is \code{TRUE}, it also includes the \code{transformed\_data} and a function \code{trans\_new()} that may be used later to apply the same transformation to e.g.
-grid points.
+The resulting \code{lg\_object} now includes the element \code{transform\_to\_marginal\_normality} set according to the input, and if this is \code{TRUE}, it also includes the \code{transformed\_data} and a function \code{trans\_new()} that may be used later to apply the same transformation to, e.g., grid points.
 If the transformation option is set to \code{FALSE}, the \code{transformed\_data} element contains the input data \code{x}, and \code{trans\_new()} is nothing more than the identity mapping for points in \(\mathbb{R}^p\).
@@
 
-Finally, the user must specify a set of bandwidths, or a method for calculating them.
-Given that the different estimation methods described in Section \ref{chap:est-method} require different sets of bandwidths (i.e joint, marginal and/or pairwise), the easiest approach for the user is to leave the selection and formatting of the bandwidths to the \code{lg\_main()}-function.
+Finally, the user must specify a set of bandwidths or a method for calculating them.
+Given that the different estimation methods described in Section \ref{chap:est-method} require different sets of bandwidths (i.e, joint, marginal, and/or pairwise), the easiest approach for the user is to leave the selection and formatting of the bandwidths to the \code{lg\_main()}-function.
 
@@
 It controls the \emph{level of localization} and thus only indirectly the \emph{smoothness} of the estimates.
-Indeed, if we concentrate on the univariate case for the moment and assume that the (single) bandwidth \(b\) is small, we see from the local likelihood function \eqref{eq:loclik} that only the very few observations closest to a fixed grid point \(x_0\) will have significant weight when determining the
+Indeed, suppose we concentrate on the univariate case for the moment and assume that the (single) bandwidth \(b\) is small.
+In that case, we see from the local likelihood function \eqref{eq:loclik} that only the very few observations closest to a fixed grid point \(x_0\) will have significant weight when determining the
 local parameter estimates \(\widehat\theta_0\left(x\right)\) at that point.
-Moving on to another nearby point \(x_1\) may then lead to a fairly different estimate \(\widehat\theta\left(x_1\right)\) because the set of observations having weight in this point is very different.
-This may again lead to rougher parameter estimates \(\widehat\theta\left(x\right)\) and in turn also to rougher density estimates \(f\left(x,\widehat\theta\left(x\right)\right)\).
+Moving on to another nearby point, \(x_1\) may then lead to a fairly different estimate \(\widehat\theta\left(x_1\right)\) because the set of observations having weight in this point is very different.
+This may, again, lead to rougher parameter estimates \(\widehat\theta\left(x\right)\) and in turn also to rougher density estimates \(f\left(x,\widehat\theta\left(x\right)\right)\).
 
@@
 We have in practice seen two automatic bandwidth selectors employed in the applications referred to in Section \ref{chap:background}: a
-cross-validation procedure, that is fairly slow to compute, but accurate with respect to density estimation, and a plug-in bandwidth that is much quicker to calculate, but less accurate with respect to density estimation.
+cross-validation procedure that is fairly slow to compute but accurate with respect to density estimation, and a plug-in bandwidth that is much quicker to calculate but less accurate with respect to density estimation.
 We use the argument
@@
 \[CV\left(b\right) = -\frac{1}{n}\sum_{i=1}^n \log f\Big(X_i; \widehat\theta^{\left(-i\right)}\left(X_i\right)\Big),\]
-where \(\widehat\theta^{\left(-i\right)}\left(x\right)\) is the parameter estimate obtained after deleting observation \(X_i\) from the data, is proportional to a quantity that estimates the Kullback-Leibler distance between \(f\left(\cdot, \widehat\theta\left(\cdot\right)\right)\) and the true density \(f\left(\cdot\right)\), see \citet{bere:tjos:2014}.
+where \(\widehat\theta^{\left(-i\right)}\left(x\right)\) is the parameter estimate obtained after deleting observation \(X_i\) from the data, is proportional to a quantity that estimates the Kullback-Leibler distance between \(f\left(\cdot, \widehat\theta\left(\cdot\right)\right)\) and the true density \(f\left(\cdot\right)\); see \citet{bere:tjos:2014}.
 The cross-validated bandwidth \(b_{CV}\) is hence given by \[b_{CV} = \arg\min_b \,\, CV\left(b\right).\] If we, for example, wish to use the simplified estimation procedure on the transformed data, we need bandwidths for the marginal estimates of the local means and standard deviations, as well as a \(2\times2\) diagonal bandwidth matrix for each pair of variables.
@@
 Using plug-in bandwidths.} Obtaining cross-validated bandwidths is unfortunately fairly slow on a standard computer.
-For sample sizes in the 500-1000 range, the process may take several minutes, which is unfeasible when embarking on analyses that require e.g.
-resampling.
-We have therefore implemented a quick plug-in bandwidth selector as well, that may suffice in many practical situations, especially at the initial or exploratory stage.
+For sample sizes in the 500-1000 range, the process may take several minutes, which is unfeasible when embarking on analyses that require, e.g., resampling.
+We have, therefore, implemented a quick plug-in bandwidth selector as well that may suffice in many practical situations, especially at the initial or exploratory stage.
 
-\citet{otne:tjos:2017} show that the simplified version of the local Gaussian fit have the same convergence rates as the corresponding nonparametric kernel density estimator, for which \citet{silv:1986} derives the plug-in formula \(b = 1.08\cdot sd\left(x\right) \cdot n^{-1/5}\).
+\citet{otne:tjos:2017} show that the simplified version of the local Gaussian fit have the same convergence rates as the corresponding nonparametric kernel density estimator for which \citet{silv:1986} derives the plug-in formula \(b = 1.08\cdot sd\left(x\right) \cdot n^{-1/5}\).
 By
@@
 The latter number is the result of regressing \(b_{CV}\) on \(n^{-1/6}\) in a large simulation experiment covering various data generating processes \citep{otneim2016multivariate}.
-We see
-the effect of switching to plugin bandwidths in the code below:
+We see the effect of switching to plug-in bandwidths in the code below:
 
@@
 \code{transform{\textunderscore}to{\textunderscore}} & & \\
-\code{  marginal{\textunderscore}normality} & Transform the data? & \code{TRUE} \\
+\code{  marginal{\textunderscore}normality} & Transform the data & \code{TRUE} \\
 \code{bw} & The bandwidths to use if already calculated & \code{NULL} \\
@@
 
-In the sub-section above we present the three most important arguments to \code{lg\_main()}.
+In the sub-section above, we present the three most important arguments to \code{lg\_main()}.
 Each of them allows the user to configure one of the three crucial modeling choices.
@@
   The user may supply the bandwidths directly to \code{lg\_main()} by passing them to the \code{bw}-argument.
-They have to be in the   correct format though, which is a list containing the vector \code{\$marginal} if \code{est\_method\ =\ "5par\_marginals\_fixed"}, and always a data frame \code{\$joint} specifying all variable pairs in the \code{x1} and \code{x2} columns and the corresponding bandwidths
+They have to be in the   correct format, though, which is a list containing the vector \code{\$marginal} if \code{est\_method\ =\ "5par\_marginals\_fixed"}, and always a data frame \code{\$joint} specifying all variable pairs in the \code{x1} and \code{x2} columns and the corresponding bandwidths
   in the \code{bw1} and \code{bw2} columns. The function \code{bw\_simple()} will assist in creating bandwidth objects.
@@
 
-\section{Statistical inference using the \texttt{lg}-package}
+\section{Statistical inference using the \texttt{lg} package}
 \label{chap:inference}
@@
 We proceed in this section to demonstrate how to implement each of the tasks that we discussed in Section \ref{chap:background}.
-The general pattern is to pass the lg-object to one of the estimation or test functions provided in the \CRANpkg{lg}-package.
-We will look at some financial data in the examples: the monthly returns on the S\&P500, FTSE100, DAX30 and TOPIX stock indices from January 1985 to March 2018 \citep{datastream}.
+The general pattern is to pass the lg-object to one of the estimation or test functions provided in the \CRANpkg{lg} package.
+We will look at some financial data in the examples: the monthly returns on the S\&P500, FTSE100, DAX30, and TOPIX stock indices from January 1985 to March 2018 \citep{datastream}.
 
@@
 We create a grid, \code{x0}, having the same number of columns as the data in the code below.
-Note that we use the pipe operator \code{\%\textgreater{}\%} from the \CRANpkg{magrittr}-package \citep{bach:wick:2014} as well as functions from the \CRANpkg{dplyr}-package \citep{wick:fran:mull:2018} for easy manipulation of data frames.
+Note that we use the pipe operator \code{\%\textgreater{}\%} from the \CRANpkg{magrittr} package \citep{bach:wick:2014} as well as functions from the \CRANpkg{dplyr} package \citep{wick:fran:mull:2018} for easy manipulation of data frames.
 We then pass the grid and the lg-object containing our modeling choices to the \code{dlg()}-function in order to do the estimation.
@@
 The last line of code creates a list containing a number of elements.
-The two most important are \code{\$loc\_cor}, which is a matrix of local correlations having one row per grid point and one column per \emph{pair of variables} (the columns correspond to the rows in \code{density\_object\$pairs}), as well as \code{\$f\_est}, which is a vector containing the estimate \(\widehat f_X\left(x\right)\) of the joint density \(f_X\left(x\right)\) in the grid points specified in \code{x0}.
+The two most important are \code{\$loc\_cor}, which is a matrix of local correlations having one row per grid point and one column per \emph{pair of variables} (the columns correspond to the rows in \code{density\_object\$pairs}), and \code{\$f\_est}, which is a vector containing the estimate \(\widehat f_X\left(x\right)\) of the joint density \(f_X\left(x\right)\) in the grid points specified in \code{x0}.
 The estimated local correlations for this example is plotted in Figure \ref{fig:lgc-plot1}, and the corresponding density estimate is plotted (along the diagonal \(x_1=x_2=x_3=x_4=x\)) in Figure \ref{fig:lgc-plot2}.
@@
 
-Obtaining the estimate of a conditional density using the \citet{otne:tjos:2018} algorithm described in Section \ref{chap:background} is very similar, but one must take particular care of the \emph{ordering} of the variables in the data set.
-The estimation function, \code{clg()}, will \emph{always} assume that the free variables come first, and that the conditioning variables come last.
+Obtaining the estimate of a conditional density using the \citet{otne:tjos:2018} algorithm described in Section \ref{chap:background} is very similar.
+However, one must take particular care of the \emph{ordering} of the variables in the data set.
+The estimation function, \code{clg()}, will \emph{always} assume that the free variables come first and the conditioning variables last.
 Let us illustrate this in the following code chunk by estimating the joint conditional density of S\&P500 and FTSE100, given that DAX30 = TOPIX = 0.
@@
 The key argument in the call to \code{clg()} above is \code{cond\ =\ c(0,\ 0)}.
-This means that the last two variables are conditioning variables (and hence that the first \(4-2=2\) variables are free), and that the value of the conditioning variables are fixed at DAX30 = 0 and TOPIX = 0 respectively.
+This means that the last two variables are conditioning variables (and hence, that the first \(4-2=2\) variables are free).
+The value of the conditioning variables are fixed at DAX30 = 0 and TOPIX = 0, respectively.
 This also means that the number of columns in the grid \code{x0} plus the number of elements in \code{cond} must equal the number of variables \(p\) in the data set, and the call to \code{clg()} will result in an error message if this requirement is not fulfilled.
@@
 \end{enumerate}
-As we noted in Section \ref{chap:background}, their practical implementations are very similar, and the \CRANpkg{lg}-package provides the function \code{ind\_test()} to perform the tests.
-Let us first consider the iid case, and generate 500 observations \code{test\_x} from the well known parabola model \(X_2 = X_1^2 + \varepsilon\), where both \(X_1\) and \(\varepsilon\) are independent and standard normal.
-In this case, \(X_1\) and \(X_2\) are uncorrelated, but obviously strongly dependent.
+As we noted in Section \ref{chap:background}, their practical implementations are very similar, and the \CRANpkg{lg} package provides the function \code{ind\_test()} to perform the tests.
+Let us first consider the i.i.d.
+case, and generate 500 observations \code{test\_x} from the well known parabola model \(X_2 = X_1^2 + \varepsilon\), where both \(X_1\) and \(\varepsilon\) are independent and standard normal.
+In this case, \(X_1\) and \(X_2\) are uncorrelated but obviously strongly dependent.
 \citet{bere:tjos:2014} considers mainly the full bivariate fit to the raw data, which we easily encode into the lg-object as before.
@@
 \end{example}
-This may take a few minutes to run on a desktop computer due to the bootstrapping.
-The small $p$-value indicate that we reject the null-hypothesis of independence between $X_1$ and $X_2$ in the parabola model defined above.
+This may take a few minutes to run on a desktop computer due to bootstrapping.
+The small $p$-value indicates that we reject the null hypothesis of independence between $X_1$ and $X_2$ in the parabola model defined above.
 We can further specify the function \(h\) and the integration area \(S\) in the test
-statistic \eqref{eq:test-statistic1}, see Table \ref{tab:indtest} for details.
+statistic \eqref{eq:test-statistic1}; see Table \ref{tab:indtest} for details.
 
@@
 \(X_t\) and \(X_{t-k}\). For example, if we wish to perform this test for \(k=1\) for one of the variables in the stock-exchange series,
-create the matrix of observations as below, and proceed exactly as in the iid case.
+create the matrix of observations as below, and proceed exactly as in the i.i.d. case.
 
@@
 This is implemented in the \code{ind\_test()}-function by setting the \code{bootstrap\_type}-argument to \code{"plain"}, which is the default option.
-When testing for serial cross-dependence we need to use a block-bootstrap procedure, and \citet{laca:tjos:2018} consider two options here: The block bootstrap with either fixed \citep{kuns:1989} or random \citep{poli:roma:1994} block sizes, and this choice is specified by choosing \code{bootstrap\_type\ =\ "block"} or \code{bootstrap\_type\ =\ "stationary"}, respectively, in the call to \code{ind\_test()}.
+When testing for serial cross-dependence, we need to use a block-bootstrap procedure, and \citet{laca:tjos:2018} consider two options here: The block bootstrap with either fixed \citep{kuns:1989} or random \citep{poli:roma:1994} block sizes.
+This choice is specified by choosing \code{bootstrap\_type\ =\ "block"} or \code{bootstrap\_type\ =\ "stationary"}, respectively, in the call to \code{ind\_test()}.
 \citet{laca:tjos:2018} do not report significant differences in test performance using the different bootstrap types.
@@
 As described in Section \ref{chap:background}, \citet{stov:tjos:huft:2014} measure the local correlation between \(\left\{X_t\right\}\) and \(\left\{Y_t\right\}\) \emph{before} and \emph{after} \(T^*\), and take significant differences between these measurements as evidence against the null hypothesis of no linkage, or contagion, between the time series.
-In order to implement this test using the \CRANpkg{lg}-package, one must create \emph{two} lg-objects: one for the observations covering the non crisis period and one covering the crisis period.
+In order to implement this test using the \CRANpkg{lg} package, one must create \emph{two} lg-objects: one for the observations covering the non-crisis period and one covering the crisis period.
 We do not enter a discussion here how to empirically identify such time periods; this is a job that must be done by the practitioner before performing the statistical test.
 
-Let us illustrate the implementation of this test by looking at the same financial returns data that we have used in preceding sections, but this time we will, in the spirit of \citet{stov:tjos:huft:2014}, concentrate on \emph{GARCH(1,1)-filtrated daily returns on the S\&P500 and FTSE100 indices from 2 January 1985 to 29 April 1987} in order to test for financial contagion between the US and UK stock markets following the global stock market crash of 19 October 1987 (``Black Monday'').
-Assume that these observations are loaded into the R workspace as the \(n_1 \times 2\) data frame \code{x\_nc} containing the observations covering the \(n_1 = 728\) days preceding the crisis , and the \(n_2 \times 2\) data frame \code{x\_c} containing the observations covering the \(n_2 = 140\) consecutive trading days starting on \emph{Black Monday} (see the online code supplement for details concerning the GARCH-filtration and data processing).
-In the code below we construct one lg-object for each of these data frames with configuration matching the setup used by \citet{stov:tjos:huft:2014}, and perform the test by means of the \code{cont\_test()}-function.
+Let us illustrate the implementation of this test by looking at the same financial returns data that we have used in preceding sections.
+However, this time we will, in the spirit of \citet{stov:tjos:huft:2014}, concentrate on \emph{GARCH(1,1)-filtrated daily returns on the S\&P500 and FTSE100 indices from 2 January 1985 to 29 April 1987} in order to test for financial contagion between the US and UK stock markets following the global stock market crash of 19 October 1987 (``Black Monday'').
+Assume that these observations are loaded into the R workspace as the \(n_1 \times 2\) data frame \code{x\_nc} containing the observations covering the \(n_1 = 728\) days preceding the crisis and the \(n_2 \times 2\) data frame \code{x\_c} containing the observations covering the \(n_2 = 140\) consecutive trading days starting on \emph{Black Monday} (see the online code supplement for details concerning the GARCH-filtration and data processing).
+In the code below, we construct one lg-object for each of these data frames with configuration matching the setup used by \citet{stov:tjos:huft:2014} and perform the test by means of the \code{cont\_test()}-function.
 
@@
 
-Consider finally the work by \citet{otne:tjos:2019} who take the off-diagonal element in the local correlation matrix corresponding to the local conditional covariance matrix \eqref{eq:cond-sig} or \eqref{eq:cond-sig-trans} as a local measure of conditional dependence between two stochastic variables \(X^{\left(1\right)}\) and \(X^{\left(2\right)}\) given the stochastic vector \(X^{\left(3\right)}\).
+Consider the work finally by \citet{otne:tjos:2019}, who take the off-diagonal element in the local correlation matrix corresponding to the local conditional covariance matrix \eqref{eq:cond-sig} or \eqref{eq:cond-sig-trans} as a local measure of conditional dependence between two stochastic variables \(X^{\left(1\right)}\) and \(X^{\left(2\right)}\) given the stochastic vector \(X^{\left(3\right)}\).
 Furthermore, in the case with data having been transformed to marginal standard normality, they take the statistic
@@
 as a measure of \emph{global} conditional dependence.
-The \CRANpkg{lg}-package provides two key functions in this framework.
+The \CRANpkg{lg} package provides two key functions in this framework.
 The first, \code{partial\_cor()}, calculates the local partial correlations as well as their estimated standard deviations on a specified grid in the \(\left(x^{\left(1\right)}, x^{\left(2\right)}, x^{\left(3\right)}\right)\)-space, and is essentially a wrapper function for the \code{clg()}-function presented in Section \ref{chap:density}.
@@
 in the period starting in January 2009, the converse of which is a sufficient, but not necessary, condition for \(R_{\textrm{SP500}, t}\) Granger causing \(R_{\textrm{FTSE100}, t}\).
-We perform the test by running the code below, where \code{x} is a data frame having the following columns strictly ordered as \(R_{\textrm{FTSE100}, t}\), \(R_{\textrm{SP500}, t-1}\) and \(R_{\textrm{FTSE100}, t-1}\) (see the online supplement for the pre-processing of data).
+We perform the test by running the code below, where \code{x} is a data frame having the following columns strictly ordered as \(R_{\textrm{FTSE100}, t}\), \(R_{\textrm{SP500}, t-1}\), and \(R_{\textrm{FTSE100}, t-1}\) (see the online supplement for the pre-processing of data).
 
@@
 the conditional density estimates \(\widehat f_{X^{\left(1\right)}|X^{\left(3\right)}}\left(x^{\left(1\right)}|x^{\left(3\right)}\right)\) and \(\widehat f_{X^{\left(2\right)|X^{\left(3\right)}}}\left(x^{\left(2\right)}|x^{\left(3\right)}\right)\), as obtained by the \code{clg()}-function, using an approximated accept-reject algorithm.
-In order to avoid excessive optimization of the local likelihood function \eqref{eq:loclik}, we estimate \(f_{X^{\left(1\right)|X^{\left(3\right)}}}\) and \(f_{X^{\left(2\right)|X^{\left(3\right)}}}\) on the univariate regular grids \(x_0^{\left(1\right)}\) and \(x_0^{\left(2\right)}\) respectively (while keeping \(x^{\left(3\right)}\) fixed at the observed values of \(X^{\left(3\right)}\)), and produce interpolating functions \(\tilde f_{X^{\left(1\right)}|X^{\left(3\right)}}\) and \(\tilde f_{X^{\left(2\right)}|X^{\left(3\right)}}\) using cubic splines.
+In order to avoid excessive optimization of the local likelihood function \eqref{eq:loclik}, we estimate \(f_{X^{\left(1\right)|X^{\left(3\right)}}}\) and \(f_{X^{\left(2\right)|X^{\left(3\right)}}}\) on the univariate regular grids \(x_0^{\left(1\right)}\) and \(x_0^{\left(2\right)}\), respectively (while keeping \(x^{\left(3\right)}\) fixed at the observed values of \(X^{\left(3\right)}\)), and produce interpolating functions \(\tilde f_{X^{\left(1\right)}|X^{\left(3\right)}}\) and \(\tilde f_{X^{\left(2\right)}|X^{\left(3\right)}}\) using cubic splines.
 It is much less computationally intensive to generate replicates from \(\tilde f\) than directly from \(\widehat f\).
 
-We refer to the documentation of the \CRANpkg{lg}-package for details on how to finely tune the behavior of the bootstrapping algorithm by altering the arguments of the \code{ci\_test()}-function, and limit our treatment to describing the arguments most suitable for modifications by the user in Table \ref{tab:arguments-ci-test}.
+We refer to the documentation of the \CRANpkg{lg} package for details on how to finely tune the behavior of the bootstrapping algorithm by altering the arguments of the \code{ci\_test()}-function and limiting our treatment to describing the arguments most suitable for modifications by the user in Table \ref{tab:arguments-ci-test}.
 
@@
 \end{example}
-The conditional independence test does not provide evidence against the null hypothesis \eqref{eq:cond-ind-test}.
+The conditional independence test does not provide evidence against the null-hypothesis \eqref{eq:cond-ind-test}.
 
@@
 We conclude this article by describing the \code{corplot()} function for drawing dependence maps such as those displayed in Figure \ref{fig:initial-example}.
-\citet{bere:klep:tjos:2014} report on such capabilities in the \CRANpkg{localgauss}-package, but the possibility of creating dependence maps was unfortunately removed from \CRANpkg{localgauss} in the latest version 0.4.0 due to incompatibilities with the \CRANpkg{ggplot2} \citep{wick:2016} plotting engine.
+\citet{bere:klep:tjos:2014} report on such capabilities in the \CRANpkg{localgauss} package, but the possibility of creating dependence maps was unfortunately removed from \CRANpkg{localgauss} in the latest version 0.4.0 due to incompatibilities with the \CRANpkg{ggplot2} \citep{wick:2016} plotting engine.
 We make up for this loss by providing \code{corplot()}, a function that plots the estimated local correlations as provided by \code{dlg()}, or the estimated local \emph{partial} correlations as provided by \code{partial\_cor()}.
@@
 The plotting function is highly customizable and provides a number of options covering most basic graphical options.
-Users well versed in the \CRANpkg{ggplot2}-package may also modify the graphical object returned by \code{corplot()} in the standard way by adding layers as demonstrated in the example below.
+Users well versed in the \CRANpkg{ggplot2} package may also modify the graphical object returned by \code{corplot()} in the standard way by adding layers as demonstrated in the example below.
 
-In the first example we generate a set of bivariate normally distributed data using the \CRANpkg{mvtnorm}-package \citep{gentz:etal:2018} and estimate the local Gaussian correlation on a regular \code{grid} using the \code{dlg()}-function.
+In the first example, we generate a set of bivariate normally distributed data using the \CRANpkg{mvtnorm} package \citep{gentz:etal:2018} and estimate the local Gaussian correlation on a regular \code{grid} using the \code{dlg()}-function.
 Passing the resulting \code{dlg\_object} to \code{corplot()} without further arguments results in Figure \ref{fig:graphics1}.
@@
 We may tweak the appearance of our dependence map by passing further arguments to \code{corplot()}.
-Some of the options are demonstrated in the code chunk below, in which we, for example, superimpose the observations (by setting \code{plot\_obs\ =\ TRUE)} as well as preventing the the estimated local correlations from being plotted in areas without data.
-The latter option is available through the argument \code{plot\_thres}, which works by calculating a bivariate kernel density estimate \(\tilde f\left(x_1, x_2\right)\) for the pair of  variables in question, and only allowing \(\widehat \rho\left(x_1,x_2\right)\) to be plotted if \(\tilde f\left(x_1,x_2\right)/\max\tilde f\left(\cdot\right) >\) \code{plot\_thres}.
+Some of the options are demonstrated in the code chunk below, in which we, for example, superimpose the observations (by setting \code{plot\_obs\ =\ TRUE)} and preventing the estimated local correlations from being plotted in areas without data.
+The latter option is available through the argument \code{plot\_thres}, which works by calculating a bivariate kernel density estimate \(\tilde f\left(x_1, x_2\right)\) for the pair of  variables in question and only allowing \(\widehat \rho\left(x_1,x_2\right)\) to be plotted if \(\tilde f\left(x_1,x_2\right)/\max\tilde f\left(\cdot\right) >\) \code{plot\_thres}.
 Adding layers to a dependence map using the ordinary \CRANpkg{ggplot2} syntax works as well, which we demonstrate in Figure \ref{fig:graphics2} by changing the \CRANpkg{ggplot2} theme.
@@
 The statistical literature has seen a number of applications of local Gaussian approximations in the last decade, covering several topics in dependence modeling and inference, as well as the estimation of multivariate density and conditional density functions.
-In this paper, we demonstrate the implementation of these methods i the R programming language using the \CRANpkg{lg}-package, as well as the graphical representation of the estimated local Gaussian correlation.
-The package is complete, in the sense that all major methods that have been published within this framework is now easily accessible to the practitioner.
+In this paper, we demonstrate the implementation of these methods in the R programming language using the \CRANpkg{lg} package, as well as the graphical representation of the estimated local Gaussian correlation.
+The package is complete in the sense that all major methods that have been published within this framework is now easily accessible to the practitioner.
 The package is also designed with a modular infrastructure that allows future methodological developments using local Gaussian approximations to be easily added to the package.
